---
title: "Comparison_Teichman"
author: "Gabriele Lubatti"
date: "3/31/2021"
output: html_document
---
```{r }
gc()
rm(list=ls())
library(ggplot2)
library(ComplexHeatmap)
library(circlize)
library(Seurat)
library(DescTools)
library(rlist)
library(devtools)
library(clustree)
library(Biobase)
library(CellSIUS)
library('FiRE')
library(RaceID)
library(singleCellHaystack)
library(mcclust)
```

```{r }
Function_fisher=function(a,b,c){
matrice=matrix(c(sum(c[c%in%a]%in%c[c%in%b]),sum(c[c%in%a]%in%c[!c%in%b]),sum(c[!c%in%a]%in%c[c%in%b]),sum(c[!c%in%a]%in%c[!c%in%b])),nrow=2,ncol=2,byrow = T)
#Fisher test
fisher.test(matrice)
}
```

```{r }
Test_Fisher_fire=function(entropy_cluster,method_cluster){
level=levels(as.factor(entropy_cluster))
p_value=rep(0,length(level))
odds_ratio=rep(0,length(level))
for (i in 1:length(level)){
  set_2=names(entropy_cluster)[entropy_cluster==level[i]]
set_all=colnames(raw_elmir_5_30)
show=Function_fisher(set_1,set_2,set_all)

p_value[i]=as.numeric(show$p.value)
names(p_value)[i]=level[i]
odds_ratio[i]=as.numeric(show$estimate)
names(odds_ratio)[i]=level[i]
}
return(list(p_value,odds_ratio))
}
```



```{r}
white_black_marker=function(cluster_race_entropy,single_cluster,norm_race,marker_list,soglia=1){
  
  white_black_7=apply(norm_race[marker_list[names(marker_list)==single_cluster],],1,function(x){
  mean_one=median(x[cluster_race_entropy==single_cluster])
  mean_different=rep(0,length(levels(factor(cluster_race_entropy[cluster_race_entropy!=single_cluster]))))
  for ( i in 1:length(levels(factor(cluster_race_entropy[cluster_race_entropy!=single_cluster])))){
  mean_different[i]=median(x[cluster_race_entropy==levels(factor(cluster_race_entropy[cluster_race_entropy!=single_cluster]))[i]])}
  
  if (mean_one>soglia&sum(mean_different)==0){
    return(TRUE)
  }
  else{return(FALSE)}
})
  return(white_black_7)
}
```
```{r }
find_jacardi=function(cluster_new,cluster_old,norm_matrix,single_new,single_old){
unione=union(colnames(norm_matrix)[cluster_new==single_new],colnames(norm_matrix)[cluster_old==single_old])
intersezione=intersect(colnames(norm_matrix)[cluster_new==single_new],colnames(norm_matrix)[cluster_old==single_old])

jacardi=length(intersezione)/length(unione)
print(jacardi)
return(jacardi)
}
```

```{r}

barplot_cell_cycle=function(cell_cycle,cluster,title,fraction=TRUE){
dfForPlot <- data.frame(Phases =cell_cycle
                                ,clu =cluster)

if(fraction==TRUE){
pnm <-ggplot(dfForPlot, aes(y=cell_cycle,x=cluster))+geom_bar( aes(fill = cell_cycle,y = (..count..)/sum(..count..)),position="fill")+ ggtitle(title)+ylab("Fraction of cells")+xlab("Cluster")+labs(fill="Elmir annotation")+theme(text = element_text(size=10),
        axis.text.x = element_text(angle=90, hjust=1,size=6)) 
pnm}

if(fraction!=TRUE){
  pnm <-ggplot(dfForPlot, aes(y=cell_cycle,x=cluster))+geom_bar( aes(fill = cell_cycle,y = (..count..),position="fill"))+ ggtitle(title)+ylab("Number of cells")+xlab("Cluster")+labs(fill="Elmir annotation")+theme(text = element_text(size=10),
        axis.text.x = element_text(angle=90, hjust=1,size=6)) 
pnm


}
return(list(pnm))}
```

```{r }
#install_github("ScialdoneLab/EntropyMixingR",auth_token="8f6bf418b6ea0274f5f832b3a96de9b63be0997a",ref="master")


```

```{r }
library(IDENTOM)
```






```{r }
load(system.file("extdata", "raw_elmir_5_30.Rda", package ="IDENTOM"))
load(system.file("extdata", "CELLSIUS_elmir.Rda", package ="EntropyMixingR"))
CELLSIUS_elmir=result

umap_elmir=readRDS(system.file("extdata", "annot_umap.rds", package ="IDENTOM"))

vedo=load(system.file("extdata", "experiment_name_StatisticsTable_afterLOESS_elmir.RData", package ="EntropyMixingR"))

vedo=load(system.file("extdata", "elmir_entropy_5_30.Rda", package ="EntropyMixingR"))

vedo=load(system.file("extdata", "experiment_name_FinalClustering_elmir.RData", package ="EntropyMixingR"))



setwd("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/elmir_data")
load(file="final_cluster_version_sub.Rda")


```

```{r}

cordinate_umap=umap_elmir[,2:3]

general_cluster_version=as.vector(umap_elmir$cluster_id)

```



## Giniclust2 
```{r }
setwd("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/GiniClust2-master/risultato/results")
final_geni=load("experiment_name_GeneList_final.RData")
#GeneList.final

```


```{r }
setwd("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/elmir_data")
load("norm_elmir_5_30.Rda")
```

```{r }


Plot_gene(norm_elmir_5_30,cordinate_umap,"NANOS3","NANOS3")
Plot_gene(norm_elmir_5_30,cordinate_umap,"DPPA5","DPPA5")
Plot_gene(norm_elmir_5_30,cordinate_umap,"NANOG","NANOG")

```

```{r }
find_positive_cells=function(norm_elmir_5_30,geni_marker,threshold){

logic_cell=apply(norm_elmir_5_30[geni_marker,],2,function(x){
  
  if (sum(x>threshold)==length(x))
  {return(TRUE)}
  else{return(FALSE)}
})
positive_cells=colnames(norm_elmir_5_30)[logic_cell]
return(positive_cells)
}



```

```{r }
find_jacardi_new=function(cluster_new,norm_matrix,single_new,cells_positive){
unione=union(colnames(norm_matrix)[cluster_new==single_new],cells_positive)
intersezione=intersect(colnames(norm_matrix)[cluster_new==single_new],cells_positive)

jacardi=length(intersezione)/length(unione)
print(jacardi)
return(jacardi)
}
```

```{r }
find_top_marker=function(marker_elmir_all,norm_elmir_5_30,name_cluster,fraction){
geni_7=marker_elmir_all[names(marker_elmir_all)==name_cluster]
logic_7=apply(norm_elmir_5_30[geni_7,final_cluster_version_sub==name_cluster],1,function(x){
  if (length(x[x>1])>=(length(x)*fraction)){
    return(TRUE)
  }
  else{return(FALSE)}
})


geni_7_top=geni_7[logic_7]
names(geni_7_top)=rep(name_cluster,length(geni_7_top))

return(geni_7_top)
}
```


##For PGC definition
```{r }
geni_pgc=c("NANOS3","NANOG","DPPA5")
positive_cells=find_positive_cells(norm_elmir_5_30,threshold=0,geni_pgc)
```

```{r }
for (i in 1:length(names(table(as.character(finalCluster))))){

bo=find_jacardi_new(as.character(finalCluster),norm_elmir_5_30,i,positive_cells)
  if (bo>=0.50){
    print(paste0("yes we have-",i))
  }
}

```

```{r}
setwd("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/elmir_data")
load(file="markers_first_elmir_final.Rda")
marker_elmir_all=markers_first_elmir_final[[3]]


load("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/elmir_data/final_cluster_version_sub.Rda")
```



## Yes first approach


## Top ExE meso
```{r}



top_endo=white_black_marker(final_cluster_version_sub,"ExE Mesoderm_1",norm_elmir_5_30,marker_elmir_all,0)
top_endo=names(top_endo)[top_endo]


mean_top_endo=apply(norm_elmir_5_30[top_endo,final_cluster_version_sub=="ExE Mesoderm_1"],1,mean)
mean_top_endo=sort(mean_top_endo,decreasing = T)

top_endo=names(mean_top_endo)[1:4]
names(top_endo)=rep("ExE Mesoderm_1",length(top_endo))

Plot_balons(norm_elmir_5_30,final_cluster_version_sub,top_endo,length(top_endo),max_size=5,text_size=10)

geni_marker=top_endo
positive_cells=find_positive_cells(norm_elmir_5_30,threshold=0,geni_marker)

for (i in 1:length(names(table(as.character(finalCluster))))){

bo=find_jacardi_new(as.character(finalCluster),norm_elmir_5_30,i,positive_cells)
  if (bo>=0.50){
    print(paste0("yes we have-",i))
  }
}


```

## Top endo
```{r}



top_endo=white_black_marker(final_cluster_version_sub,"Endoderm_2",norm_elmir_5_30,marker_elmir_all,1)
top_endo=names(top_endo)[top_endo]


mean_top_endo=apply(norm_elmir_5_30[top_endo,final_cluster_version_sub=="Endoderm_2"],1,mean)
mean_top_endo=sort(mean_top_endo,decreasing = T)

top_endo=names(mean_top_endo)[1:4]
names(top_endo)=rep("Endoderm_2",length(top_endo))

Plot_balons(norm_elmir_5_30,final_cluster_version_sub,top_endo,length(top_endo),max_size=5,text_size=10)

geni_marker=top_endo
positive_cells=find_positive_cells(norm_elmir_5_30,threshold=0,geni_marker)

for (i in 1:length(names(table(as.character(finalCluster))))){

bo=find_jacardi_new(as.character(finalCluster),norm_elmir_5_30,i,positive_cells)
  if (bo>=0.50){
    print(paste0("yes we have-",i))
  }
}


```

## Top hemo
```{r}



top_hemo=white_black_marker(final_cluster_version_sub,"Hemogenic Endothelial Progenitors_4",norm_elmir_5_30,marker_elmir_all,1)
top_hemo=names(top_hemo)[top_hemo]


mean_top_hemo=apply(norm_elmir_5_30[top_hemo,final_cluster_version_sub=="Hemogenic Endothelial Progenitors_4"],1,mean)
mean_top_hemo=sort(mean_top_hemo,decreasing = T)

top_hemo=names(mean_top_hemo)[1:4]
names(top_hemo)=rep("Hemogenic Endothelial Progenitors_4",length(top_hemo))

Plot_balons(norm_elmir_5_30,final_cluster_version_sub,top_hemo,length(top_hemo),max_size=5,text_size=4)

geni_marker=top_hemo
positive_cells=find_positive_cells(norm_elmir_5_30,threshold=0,geni_marker)

for (i in 1:length(names(table(as.character(finalCluster))))){

bo=find_jacardi_new(as.character(finalCluster),norm_elmir_5_30,i,positive_cells)
  if (bo>=0.50){
    print(paste0("yes we have-",i))
  }
}


```


## GiniClust2


```{r }
Plot_umap(cordinate_umap,as.character(finalCluster))
```

```{r}
table(final_cluster_version_sub,finalCluster)
```

```{r }



for (i in levels(factor(finalCluster))){
jacardi=find_jacardi(final_cluster_version_sub,finalCluster,raw_elmir_5_30,"Hemogenic Endothelial Progenitors_4",i)
  print(paste0(i,"-",jacardi))
}

for (i in levels(factor(finalCluster))){
jacardi=find_jacardi(final_cluster_version_sub,finalCluster,raw_elmir_5_30,"Endoderm_2",i)
  print(paste0(i,"-",jacardi))
}


for (i in levels(factor(finalCluster))){
jacardi=find_jacardi(final_cluster_version_sub,finalCluster,raw_elmir_5_30,"ExE Mesoderm_1",i)
  print(paste0(i,"-",jacardi))
}

for (i in levels(factor(finalCluster))){
jacardi=find_jacardi(final_cluster_version_sub,finalCluster,raw_elmir_5_30,"PGC",i)
  print(paste0(i,"-",jacardi))
}

```





## Comparison with higi gini index

```{r }
setwd("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/elmir_data")
vedo=load("elmir_hyper_old.Rda")

```

```{r }
entropy_genes=row.names(result)[result[,2]<1]
geni_top=row.names(result)[result[,1]<0.0001&result[,2]==0]
```



```{r }
background=row.names(result)
```




```{r }
library(UpSetR)
listInput <- list(entropy_genes=entropy_genes,gini_genes=GeneList.final)

upset(fromList(listInput), order.by = "freq")

listInput <- list(entropy_genes=geni_top,gini_genes=GeneList.final)

upset(fromList(listInput), order.by = "freq")


```

```{r }
Function_fisher=function(a,b,c){
matrice=matrix(c(sum(c[c%in%a]%in%c[c%in%b]),sum(c[c%in%a]%in%c[!c%in%b]),sum(c[!c%in%a]%in%c[c%in%b]),sum(c[!c%in%a]%in%c[!c%in%b])),nrow=2,ncol=2,byrow = T)
#Fisher test
fisher.test(matrice)
}
```

```{r}
Test_Fisher=function(entropy_cluster,markers_cluster,marker_top){
level=levels(as.factor(entropy_cluster))
p_value=rep(0,length(level))
odds_ratio=rep(0,length(level))
set_1=marker_top
for (i in 1:length(level)){
  set_2=markers_cluster[names(markers_cluster)==level[i]]
set_all=markers_cluster
show=Function_fisher(set_1,set_2,set_all)

p_value[i]=as.numeric(show$p.value)
names(p_value)[i]=level[i]
odds_ratio[i]=as.numeric(show$estimate)
names(odds_ratio)[i]=level[i]
}
return(list(p_value,odds_ratio))
}

```

```{r}


setwd("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/elmir_data")
load("norm_elmir_5_30.Rda")
elmir_seurat_5=Cluster_analysis_integrate_rare(raw_elmir_5_30,"Elmir_data",0.1,5,30)
library(Biobase)
DefaultAssay(elmir_seurat_5) <- "RNA"
markers_first_elmir=markers_first_method(elmir_seurat_5,finalCluster,names(elmir_seurat_5$RNA_snn_res.0.1),20)
#markers_first_elmir=markers_first_method(elmir_seurat_5[,final_gini%in%seq(0,29)],final_gini[final_gini%in%seq(0,29)],names(elmir_seurat_5$RNA_snn_res.0.1)[final_gini%in%seq(0,29)],20)
```


```{r}
library(ComplexHeatmap)
library(circlize)
marker_plot_day_24_first=markers_first_elmir[[1]]

marker_plot_plot_day_24_first=markers_first_elmir[[2]]
marker_all_day_24_first=markers_first_elmir[[3]]



```

```{r }
test=Test_Fisher(finalCluster,marker_all_day_24_first,union(row.names(ExprM.Stat2),background))
test[[2]][test[[1]]<0.05&test[[2]]>2]
```

```{r }
Function_fisher(entropy_genes,GeneList.final,union(row.names(ExprM.Stat2),background))

"NANOS3"
"NANOG"
"SOX17"
"DPPA5"

```


## Comparison for PGC: gini index vs entropy of mixing

```{r}
vedo=load(system.file("extdata", "experiment_name_StatisticsTable_afterLOESS_elmir.RData", package ="EntropyMixingR"))
#"ExprM.Stat2"   

setwd("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/elmir_data")
vedo=load("elmir_hyper_old.Rda")


entropy_genes=row.names(result)[result[,2]<1]

max_gini=ExprM.Stat2$log2.Maxs
gini_value=ExprM.Stat2$Gini
#max_gini=gini_index$log2..Max.Expression.Level..


label_all=c(rep("No Gini gene",length(max_gini)))
#label_all[ExprM.Stat2$Gini.pvalue<0.0001]="Gini gene"
label_all[row.names(ExprM.Stat2)%in%Genelist.top_pvalue]="Gini gene"

mtcars=data.frame(label_all,gini_value,max_gini)
ggplot(mtcars, aes(x=max_gini, y=gini_value, color=label_all)) +
  geom_point()+xlab("log2 (Max Expression Level)")+ylab("Gini")+labs(color="Label")



label_all=c(rep("No Entropy gene",length(max_gini)))
label_all[row.names(ExprM.Stat2)%in%entropy_genes]="Entropy gene"

mtcars=data.frame(label_all,gini_value,max_gini)
row.names(mtcars)=row.names(ExprM.Stat2)
mtcars=mtcars[c(row.names(ExprM.Stat2)[!row.names(ExprM.Stat2)%in%entropy_genes],row.names(ExprM.Stat2)[row.names(ExprM.Stat2)%in%entropy_genes]),]
ggplot(mtcars, aes(x=max_gini, y=gini_value, color=label_all)) +
  geom_point()+xlab("log2 (Max Expression Level)")+ylab("Gini")+labs(color="Label")


```

## End GiniClust2 

## Giniclust3 from python script giniclust3

```{r }
setwd("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/elmir_data")
final_gini=read.table("all_gini.txt")
final_gini=final_gini$V1
setwd("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/elmir_data")
only_gini=read.table("only_gini.txt")
only_gini=only_gini$V1
Plot_umap(cordinate_umap,final_gini)
Plot_umap(cordinate_umap,only_gini)
```
## New approach with positive cells
```{r }




geni_marker=geni_pgc
positive_cells=find_positive_cells(norm_elmir_5_30,threshold=0,geni_marker)

final_gini=as.character(final_gini)

for (i in levels(factor(final_gini))){
print(i)
bo=find_jacardi_new((final_gini),norm_elmir_5_30,i,positive_cells)
  if (bo>=0.50){
    print(paste0("yes we have-",i))
  }
}


geni_marker=top_endo
positive_cells=find_positive_cells(norm_elmir_5_30,threshold=0,geni_marker)

final_gini=as.character(final_gini)

for (i in levels(factor(final_gini))){
print(i)
bo=find_jacardi_new((final_gini),norm_elmir_5_30,i,positive_cells)
  if (bo>=0.50){
    print(paste0("yes we have-",i))
  }
}

geni_marker=top_hemo
positive_cells=find_positive_cells(norm_elmir_5_30,threshold=0,geni_marker)

for (i in levels(factor(final_gini))){
print(i)
bo=find_jacardi_new((final_gini),norm_elmir_5_30,i,positive_cells)
  if (bo>=0.50){
    print(paste0("yes we have-",i))
  }
}
```

## Old approach with intersection between cluster

```{r }
for (i in levels(factor(final_gini))){
jacardi=find_jacardi(final_cluster_version_sub,final_gini,raw_elmir_5_30,"Endoderm_2",i)
  print(paste0(i,"-",jacardi))
}


for (i in levels(factor(final_gini))){
jacardi=find_jacardi(final_cluster_version_sub,final_gini,raw_elmir_5_30,"Hemogenic Endothelial Progenitors_4",i)
  print(paste0(i,"-",jacardi))
}

for (i in levels(factor(final_gini))){
jacardi=find_jacardi(final_cluster_version_sub,final_gini,raw_elmir_5_30,"PGC",i)
  print(paste0(i,"-",jacardi))
}

for (i in levels(factor(final_gini))){
jacardi=find_jacardi(final_cluster_version_sub,final_gini,raw_elmir_5_30,"ExE Mesoderm_1",i)
  print(paste0(i,"-",jacardi))
}



```




## Comparison with CellSIUS

## run on server hpc-build.scidom.de at path /home/ies/gabriele.lubatti/entropy_mixing/cellsius/ane
## Rscript Cellsius_elmir_new.R 

```{r}
library(CellSIUS)

#CELLSIUS_elmir=result

#Below the changed version of cellsius for detecting PGC (in this way we don't detect any more sub endo 2 and hemo_4)
 #setwd("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/elmir_data")
 #vedo=load("CELLSIUS_elmir_new.Rda")
 cluster_id=general_cluster_version
names(cluster_id)=colnames(raw_elmir_5_30)
 Final_Clusters = CellSIUS_final_cluster_assignment(CellSIUS.out=CELLSIUS_elmir, group_id=cluster_id, min_n_genes = 3)

```

```{r }
Plot_umap(cordinate_umap,Final_Clusters)
Plot_umap(cordinate_umap,cluster_id)
```

```{r }


geni_marker=geni_pgc
positive_cells=find_positive_cells(norm_elmir_5_30,threshold=0,geni_marker)
Final_Clusters=as.vector(Final_Clusters)

for (i in levels(factor(Final_Clusters))){
  print(i)
bo=find_jacardi_new(Final_Clusters,norm_elmir_5_30,i,positive_cells)
  if (bo>=0.50){
    print(paste0("yes we have-",i))
  }
}


geni_marker=top_endo
positive_cells=find_positive_cells(norm_elmir_5_30,threshold=0,geni_marker)
Final_Clusters=as.vector(Final_Clusters)

for (i in levels(factor(Final_Clusters))){
  print(i)
bo=find_jacardi_new(Final_Clusters,norm_elmir_5_30,i,positive_cells)
  if (bo>=0.50){
    print(paste0("yes we have-",i))
  }
}


geni_marker=top_hemo
positive_cells=find_positive_cells(norm_elmir_5_30,threshold=0,geni_marker)
Final_Clusters=as.vector(Final_Clusters)

for (i in levels(factor(Final_Clusters))){
  print(i)
bo=find_jacardi_new(Final_Clusters,norm_elmir_5_30,i,positive_cells)
  if (bo>=0.50){
    print(paste0("yes we have-",i))
  }
}


```

```{r }
table(final_cluster_version_sub,Final_Clusters)
```

```{r }

for (i in levels(factor(Final_Clusters))){
jacardi=find_jacardi(final_cluster_version_sub,Final_Clusters,raw_elmir_5_30,"Hemogenic Endothelial Progenitors_4",i)
  print(paste0(i,"-",jacardi))
}








for (i in levels(factor(Final_Clusters))){
jacardi=find_jacardi(final_cluster_version_sub,Final_Clusters,raw_elmir_5_30,"Endoderm_2",i)
  print(paste0(i,"-",jacardi))
}


for (i in levels(factor(Final_Clusters))){
jacardi=find_jacardi(final_cluster_version_sub,Final_Clusters,raw_elmir_5_30,"ExE Mesoderm_1",i)
  print(paste0(i,"-",jacardi))
}

for (i in levels(factor(Final_Clusters))){
jacardi=find_jacardi(final_cluster_version_sub,Final_Clusters,raw_elmir_5_30,"PGC",i)
  print(paste0(i,"-",jacardi))
}




```




```{r }

Final_Clusters_plot=Final_Clusters
Final_Clusters_plot[Final_Clusters=="Hemogenic Endothelial Progenitors_1"]="Hemogenic Endothelial Progenitors_sub_1"
Final_Clusters_plot[Final_Clusters=="Hemogenic Endothelial Progenitors_2"]="Hemogenic Endothelial Progenitors_sub_2"
Final_Clusters_plot[Final_Clusters=="Hemogenic Endothelial Progenitors_3"]="Hemogenic Endothelial Progenitors_sub_3"
Final_Clusters_plot[Final_Clusters=="Hemogenic Endothelial Progenitors_3-Hemogenic Endothelial Progenitors_1 "]="Hemogenic Endothelial Progenitors_sub_1"
Final_Clusters_plot[Final_Clusters=="Hemogenic Endothelial Progenitors_4"]="Hemogenic Endothelial Progenitors_sub_4"
Final_Clusters_plot[Final_Clusters=="Hemogenic Endothelial Progenitors"]="Hemogenic Endothelial Progenitors_sub_0"
Final_Clusters_plot[Final_Clusters=="Hemogenic Endothelial Progenitors_3-Hemogenic Endothelial Progenitors_1"]="Ambigous"
Final_Clusters_plot[Final_Clusters=="Endoderm"]="Endoderm_sub_0"
Final_Clusters_plot[Final_Clusters=="Endoderm_1"]="Endoderm_sub_1"
Final_Clusters_plot[Final_Clusters=="ExE Mesoderm"]="ExE Mesoderm_sub_0"
Final_Clusters_plot[Final_Clusters=="ExE Mesoderm_1"]="ExE Mesoderm_sub_1"



```






```{r }
Plot_umap(cordinate_umap,Final_Clusters)
```



## FiRE
```{r }

#setwd("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/FIRE")
#source('preprocess.R')

source(system.file("extdata", "preprocess.R", package ="EntropyMixingR"))

```

```{r }
data <- t(raw_elmir_5_30)
genes <- colnames(data)#It can be replaced with original gene names

data_mat <- list(mat=data, gene_symbols=genes)

preprocessedList <- ranger_preprocess(data_mat)
preprocessedData <- as.matrix(preprocessedList$preprocessedData)

model <- new(FiRE::FiRE, 100, 50, 1017881, 5489, 0)

model$fit(preprocessedData)

score <- model$score(preprocessedData)
q3 <- quantile(score, 0.75)
iqr <- IQR(score)
th <- q3 + (1.5*iqr)
th_new=q3 + (0.5*iqr)
indIqr <- which(score >= th_new)
predictions <- integer(dim(data)[1])
predictions[indIqr] <- 1 

#model$b
```


```{r }
cluster_plot=predictions
cluster_plot[predictions==1]="Rare cells"
cluster_plot[predictions==0]="NO Rare cells"
Plot_umap(cordinate_umap,cluster_plot)
```

```{r }
table(cluster_plot)
```










```{r }
predictions <- integer(dim(data)[1])
predictions[indIqr] <- 1 
names(predictions)=colnames(raw_elmir_5_30)
set_1=names(predictions)[predictions==1]
final_cluster_version_sub_new=final_cluster_version_sub
names(final_cluster_version_sub_new)=colnames(raw_elmir_5_30)


test=Test_Fisher_fire(final_cluster_version_sub_new,predictions)
```

```{r }
test[[2]][test[[1]]<=0.05&test[[2]]>2]
```


##New part







```{r }
#top_hemo
#top_endo
#positive_cells_pgc=find_positive_cells(norm_elmir_5_30,threshold=0,top_endo)
#positive_cells_pgc=find_positive_cells(norm_elmir_5_30,threshold=0,top_hemo)
positive_cells_pgc=find_positive_cells(norm_elmir_5_30,threshold=0,geni_pgc)
positive_cells_other=colnames(norm_elmir_5_30)[!colnames(norm_elmir_5_30)%in%positive_cells_pgc]

predictions_small_pgc=predictions[names(predictions)%in%positive_cells_pgc]
predictions_small_other=predictions[names(predictions)%in%positive_cells_other]

predictions_small_all=c(predictions_small_pgc,predictions_small_other)

cluster_pgc=rep("Positive PGC",length(positive_cells_pgc))
cluster_other=rep("OTHER",length(positive_cells_other))
cluster_all=c(cluster_pgc,cluster_other)


names(cluster_all)=c(positive_cells_pgc,positive_cells_other)


test=Test_Fisher_fire(cluster_all,predictions_small_all)
```

```{r }
test[[2]][test[[1]]<=0.05&test[[2]]>2]
```



## Race ID 3
```{r }

sc <- SCseq(raw_elmir_5_30)
sc <- filterdata(sc,mintotal=2000)
fdata <- getfdata(sc)

sc <- compdist(sc,metric="pearson")

sc <- clustexp(sc,clustnr =30)
```



```{r }
pgc_cells=colnames(raw_elmir_5_30)[umap_elmir$sub_cluster=="PGC"]
sc@cpart[names(sc@cpart)%in%pgc_cells]

sc <- clustexp(sc,cln=15,sat=FALSE)
```

```{r }
sc <- findoutliers(sc)
```








```{r }
Plot_umap(cordinate_umap,sc@cpart)
```


```{r }
Final_Clusters=as.vector(sc@cpart)

geni_marker=top_endo
positive_cells=find_positive_cells(norm_elmir_5_30,threshold=0,geni_marker)

table(Final_Clusters[colnames(raw_elmir_5_30)%in%positive_cells])

for (i in levels(factor(Final_Clusters))){
  print(i)
bo=find_jacardi_new(Final_Clusters,norm_elmir_5_30,i,positive_cells)
  if (bo>=0.50){
    print(paste0("yes we have-",Final_Clusters[i]))
  }
}

geni_marker=top_hemo
positive_cells=find_positive_cells(norm_elmir_5_30,threshold=0,geni_marker)

table(Final_Clusters[colnames(raw_elmir_5_30)%in%positive_cells])

for (i in levels(factor(Final_Clusters))){
  print(i)
bo=find_jacardi_new(Final_Clusters,norm_elmir_5_30,i,positive_cells)
  if (bo>=0.50){
    print(paste0("yes we have-",Final_Clusters[i]))
  }
}

geni_marker=geni_pgc
positive_cells=find_positive_cells(norm_elmir_5_30,threshold=0,geni_marker)

table(Final_Clusters[colnames(raw_elmir_5_30)%in%positive_cells])

for (i in levels(factor(Final_Clusters))){
  print(i)
bo=find_jacardi_new(Final_Clusters,norm_elmir_5_30,i,positive_cells)
  if (bo>=0.50){
    print(paste0("yes we have-",Final_Clusters[i]))
  }
}
```





```{r }
table(sc@cpart)
```
```{r }
#outliers cells
out_cells=sc@out$out
length(out_cells)
table(general_cluster_version[colnames(raw_elmir_5_30)%in%out_cells])
```
```{r }
outlier_cluster=general_cluster_version
outlier_cluster[colnames(raw_elmir_5_30)%in%out_cells]="Outliers"
outlier_cluster[!colnames(raw_elmir_5_30)%in%out_cells]="NO Outliers"


Plot_umap(cordinate_umap,outlier_cluster)
```



```{r}
for (i in levels(factor(sc@cpart))){
jacardi=find_jacardi(final_cluster_version_sub,sc@cpart,raw_elmir_5_30,"Endoderm_2",i)
  print(paste0(i,"-",jacardi))
}

for (i in levels(factor(sc@cpart))){
jacardi=find_jacardi(final_cluster_version_sub,sc@cpart,raw_elmir_5_30,"Hemogenic Endothelial Progenitors_4",i)
  print(paste0(i,"-",jacardi))
}

for (i in levels(factor(sc@cpart))){
jacardi=find_jacardi(final_cluster_version_sub,sc@cpart,raw_elmir_5_30,"ExE Mesoderm_1",i)
  print(paste0(i,"-",jacardi))
}

```




## singleCellHaystack

# Standard parameter

```{r}
#elmir_seurat_5=Cluster_analysis_integrate_rare(raw_elmir_5_30,"Elmir_data",0.1,5,30)
#norm_elmir_5_30=as.matrix(GetAssayData(elmir_seurat_5, slot = "data",assay="RNA"))
#norm_elmir=norm_elmir_5_30

```


```{r}
library(singleCellHaystack)
set.seed(123)
dat.pca=as.data.frame(Embeddings(elmir_seurat_5, reduction = "pca"))
median.per.gene <- apply(norm_elmir_5_30,1,median) # get the median read count per gene
dat.detection <- norm_elmir_5_30 > median.per.gene
res.pc30 <- haystack(x = dat.pca, detection = dat.detection , method = "highD")


show_result_haystack(res.haystack = res.pc30, n = 5)


top_values=row.names(res.pc30$results)[order(res.pc30$results$log.p.adj,decreasing = F)]
# plotting detection of Trim10

plot_gene_haystack(x = cordinate_umap, gene = top_values[1], expression = dat.detection)
plot_gene_haystack(x = cordinate_umap, gene = top_values[2], expression = dat.detection)
plot_gene_haystack(x = cordinate_umap, gene = top_values[3], expression = dat.detection)
plot_gene_haystack(x = cordinate_umap, gene = top_values[4], expression = dat.detection)

```



```{r}
setwd("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/elmir_data")
load(file="markers_first_elmir_final.Rda")
marker_elmir_all=markers_first_elmir_final[[3]]
```



```{r}
Test_Fisher=function(entropy_cluster,markers_cluster,marker_top){
level=levels(as.factor(entropy_cluster))
p_value=rep(0,length(level))
odds_ratio=rep(0,length(level))
set_1=marker_top
for (i in 1:length(level)){
  set_2=markers_cluster[names(markers_cluster)==level[i]]
set_all=markers_cluster
show=Function_fisher(set_1,set_2,set_all)

p_value[i]=as.numeric(show$p.value)
names(p_value)[i]=level[i]
odds_ratio[i]=as.numeric(show$estimate)
names(odds_ratio)[i]=level[i]
}
return(list(p_value,odds_ratio))
}
test=Test_Fisher(final_cluster_version_sub,marker_elmir_all,top_values[1:2000])

```

```{r}
test[[2]][test[[1]]<0.05&test[[2]]>2]
```

```{r}
grep("NANOS3",top_values)
which(top_values=="NANOG")
grep("SOX17",top_values)
which(top_values=="DPPA5")

which(top_values%in%top_endo)
which(top_values%in%top_hemo)
```

```{r}

res.top <- show_result_haystack(res.haystack = res.pc30, n = 1000)
# cluster DEGs by their expression pattern in the 2D plot
genes.top <- row.names(res.top)
res.hc <- hclust_haystack(x = cordinate_umap, genes = genes.top, detection = dat.detection)
res.hc.clusters <- cutree(res.hc, k=5)
table(res.hc.clusters)

pl <- lapply(1:5, function(cluster) {
  gene.set <- names(res.hc.clusters)[res.hc.clusters==cluster]
  plot.title <- paste0("Cluster ", cluster)
  p <- plot_gene_set_haystack(x = cordinate_umap, genes = gene.set, detection = dat.detection)
  p + ggtitle(plot.title) + theme(legend.title = element_text(size = 8))
})
pl

```

```{r }
library(UpSetR)
load(system.file("extdata", "elmir_entropy_5_30.Rda", package ="EntropyMixingR"))

listInput <- list(EntropyMixing=(entropy_genes),singleCellHaystack=genes.top)

upset(fromList(listInput), order.by = "freq")




```




```{r}


p=list()
for(i in genes.top[1:3]){
  q=Plot_gene(norm_elmir_5_30,cordinate_umap,i,paste(i,which(genes.top==i)))
  p=list(p,q)
}
p
```

##SAM application 


```{r}
setwd("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/elmir_data")
load(file="markers_first_elmir_final.Rda")
marker_elmir_all=markers_first_elmir_final[[3]]
```

```{r}
sam_genes=read.csv("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/published_data/sam_output/sorted_sam_genes.csv")
sam_genes=sam_genes$X0
```



```{r}
Test_Fisher=function(entropy_cluster,markers_cluster,marker_top){
level=levels(as.factor(entropy_cluster))
p_value=rep(0,length(level))
odds_ratio=rep(0,length(level))
set_1=marker_top
for (i in 1:length(level)){
  set_2=markers_cluster[names(markers_cluster)==level[i]]
set_all=markers_cluster
show=Function_fisher(set_1,set_2,set_all)

p_value[i]=as.numeric(show$p.value)
names(p_value)[i]=level[i]
odds_ratio[i]=as.numeric(show$estimate)
names(odds_ratio)[i]=level[i]
}
return(list(p_value,odds_ratio))
}
test=Test_Fisher(final_cluster_version_sub,marker_elmir_all,sam_genes[1:2000])

```

```{r}
test[[2]][test[[1]]<0.05&test[[2]]>2]
```

```{r}
grep("NANOS3",sam_genes)
which(sam_genes=="NANOG")
grep("SOX17",sam_genes)
which(sam_genes=="DPPA5")


```



```{r }
#load(system.file("extdata", "elmir_hyper_old.Rda", package ="EntropyMixingRnew"))
load(system.file("extdata", "elmir_december_2021.Rda", package ="EntropyMixingRnew"))

```



```{r }
entropy_genes=row.names(result)[result[,2]<1]
geni_top=row.names(result)[result[,2]<0.0001&result[,1]==0]
geni_top_top=row.names(result)[order(as.numeric(result[,2]))]
```


```{r}
grep("NANOS3",geni_top_top)
which(geni_top_top=="NANOG")
grep("SOX17",geni_top_top)
which(geni_top_top=="DPPA5")


```


