---
title: "Other dataset"
author: "Gabriele Lubatti"
date: "10/22/2021"
output: html_document
---
```{r }
gc()
rm(list=ls())
library(ggplot2)
library(ComplexHeatmap)
library(circlize)
library(Seurat)
library(DescTools)
library(rlist)
library(devtools)
library(clustree)
library(Biobase)
library(CellSIUS)
library('FiRE')
library(RaceID)
library(singleCellHaystack)
library(mcclust)
```



## Function Definition
```{r }
find_jacardi=function(cluster_new,cluster_old,norm_matrix,single_new,single_old){
unione=union(colnames(norm_matrix)[cluster_new==single_new],colnames(norm_matrix)[cluster_old==single_old])
intersezione=intersect(colnames(norm_matrix)[cluster_new==single_new],colnames(norm_matrix)[cluster_old==single_old])

jacardi=length(intersezione)/length(unione)
print(jacardi)
return(jacardi)
}


white_black_marker=function(cluster_race_entropy,single_cluster,norm_race,marker_list,soglia=1){
  
  white_black_7=apply(norm_race[marker_list[names(marker_list)==single_cluster],],1,function(x){
  mean_one=median(x[cluster_race_entropy==single_cluster])
  mean_different=rep(0,length(levels(factor(cluster_race_entropy[cluster_race_entropy!=single_cluster]))))
  for ( i in 1:length(levels(factor(cluster_race_entropy[cluster_race_entropy!=single_cluster])))){
  mean_different[i]=median(x[cluster_race_entropy==levels(factor(cluster_race_entropy[cluster_race_entropy!=single_cluster]))[i]])}
  
  if (mean_one>soglia&sum(mean_different)==0){
    return(TRUE)
  }
  else{return(FALSE)}
})
  return(white_black_7)
}


find_jacardi_genes=function(entropy_mixing_old,entropy_mixing_new){
unione=union(entropy_mixing_old,entropy_mixing_new)
intersezione=intersect(entropy_mixing_old,entropy_mixing_new)

jacardi=length(intersezione)/length(unione)
print(jacardi)
return(jacardi)
}

```

```{r }
Plot_balons=function(norm_counts_ane_24,final_cluster_version_sub,marker_plot_day_24_first_plot,max_number,total_cells,max_size=5,text_size=7){
  #norm_counts_ane_24=norm_rabbit_es_all_new
  #final_cluster_version_sub=cluster_plot
  #marker_plot_day_24_first_plot=final_markers_human[names(final_markers_human)==9]
  #max_number=10
  #total_cells=length(cluster_plot)
  #max_size=5
livelli=levels(factor(final_cluster_version_sub))
all_markers_plot=rep(list(0),length(livelli))
all_markers=rep(list(0),length(livelli))
for ( i in 1:length(livelli)){
  genes_4_0=marker_plot_day_24_first_plot[(names(marker_plot_day_24_first_plot)==livelli[i])]
  if(length(genes_4_0)>0){
  genes_4_0=genes_4_0[1:max_number]
  all_markers[[i]]=genes_4_0
genes_4_0_plot=paste(livelli[i],genes_4_0,sep="_")
all_markers_plot[[i]]=genes_4_0_plot
  }
  }

all_markers=lapply(all_markers, function(x) {x[x!=0]})
all_markers_plot=lapply(all_markers_plot, function(x) {x[x!=0]})

all_markers_values=rep(list(0),length(livelli))
for ( i in 1:length(livelli)){
  gene_values_0=apply(norm_counts_ane_24[unlist(all_markers, use.names=FALSE),final_cluster_version_sub==livelli[i]],1,mean)
  all_markers_values[[i]]=gene_values_0
}
#values=rep(list(0),length(livelli))
#for ( i in 1:length(livelli)){
  #valore=sum(final_cluster_version_sub==livelli[i])/total_cells
  #values[[i]]=rep(valore,length(unlist(all_markers)))
#}

values=rep(list(0),length(livelli))
for ( i in 1:length(livelli)){
  valore=apply(norm_counts_ane_24[unlist(all_markers, use.names=FALSE),final_cluster_version_sub==livelli[i]],1,function(x){sum(x>1)/length(x)})
  values[[i]]=valore
}

cluster_label_all=rep(list(0),length(livelli))
for ( i in 1:length(livelli)){
  cluster_label_all[[i]]=rep(livelli[i],length(unlist(all_markers)))
}


all_markers_plot=unlist(all_markers_plot)
cluster_label_all=unlist(cluster_label_all)
values=unlist(values)
all_markers_values=unlist(all_markers_values)
balloon_melted=data.frame(all_markers_plot,cluster_label_all,values)
colnames(balloon_melted)=c("genes_plot","cluster_label_all","values")

p <- ggplot(balloon_melted, aes(x =cluster_label_all, y = as.character(genes_plot))) 
p+geom_point( aes(size=values,colour=as.numeric(all_markers_values)))+theme(panel.background=element_blank(), panel.border = element_rect(colour = "blue", fill=NA, size=3))+scale_size_area(max_size=max_size)+theme(text = element_text(size=text_size))+ scale_colour_gradient(low = "yellow", high = "black", na.value = NA)+labs(colour="Mean log2 norm expression",size="Value (fraction cells above 1 log2 norm counts)")+xlab("Cluster")+ylab("Markers")
}
```

```{r }
find_positive_cells=function(norm_elmir_5_30,geni_marker,threshold){

logic_cell=apply(norm_elmir_5_30[geni_marker,],2,function(x){
  
  if (sum(x>threshold)==length(x))
  {return(TRUE)}
  else{return(FALSE)}
})
positive_cells=colnames(norm_elmir_5_30)[logic_cell]
return(positive_cells)
}



```

```{r }
find_jacardi_new=function(cluster_new,norm_matrix,single_new,cells_positive){
unione=union(colnames(norm_matrix)[cluster_new==single_new],cells_positive)
intersezione=intersect(colnames(norm_matrix)[cluster_new==single_new],cells_positive)

jacardi=length(intersezione)/length(unione)
print(jacardi)
return(jacardi)
}
```

## From Gini paper


## We focused on a subset of 2509 cells obtained from the Day 0 stage, where the cells remained undifferentiated. On average, about 13,000 unique molecular identifiers (UMIs) were detected in each cell, corresponding to nearly 6000 genes.
## Raw count matrix from Klein AM, Mazutis L, Akartuna I, Tallapragada N, Veres A, Li V, et al. Droplet barcoding for single-cell transcriptomics applied to embryonic stem cells. Cell. 2015;161:1187–201.

## 2clc cluster made up only by 3 cells from 2509 cells more or less
## Fig.3b 

## Pre processing
```{r }
#2485 cells and  22,830 genes
#feeders=read.csv2("/Users/gabriele.lubatti/Desktop/Phd/Early_embryonic_like_cells/Monkey/10x_data/Output_script/GSE114952_RNA-Seq_FPKM.csv")
#test_data_0=read.csv("/Users/gabriele.lubatti/Downloads/GSE65525_RAW/GSM1599494_ES_d0_main.csv")
#Used in the paper

test_data_1=read.csv("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/published_data/GSM1599495_ES_d0_biorep_techrep1.csv")
row.names(test_data_1)=test_data_1[,1]
test_data_1=test_data_1[,-1]
#test_data_2=read.csv("/Users/gabriele.lubatti/Downloads/GSE65525_RAW/GSM1599496_ES_d0_biorep_techrep2.csv")


```


```{r }
filter_gene=apply(test_data_1,1,function(x){
  x=x[x>0]
  if(length(x)<=3){
    return (FALSE)
  }
  else {
    return(TRUE)
  }
})
```

```{r }
filter_cells=apply(test_data_1,2,function(x){
  x=x[x>0]
  if(length(x)<2000){
    return (FALSE)
  }
  else {
    return(TRUE)
  }
})
```

```{r }
test_data_1=test_data_1[,filter_cells]
```

```{r }
giniclust_seurat=Cluster_analysis_integrate_rare(test_data_1,"GiniClust",0.1,5,30)
```

```{r }
norm_gini=as.matrix(GetAssayData(giniclust_seurat, slot = "data",assay="RNA"))
norm_gini=norm_gini[,1:2484]

general_gini=giniclust_seurat$RNA_snn_res.0.1
general_gini=general_gini[1:2484]

knn_matrix_gini_5_30=as.matrix(giniclust_seurat@graphs$RNA_nn)
knn_matrix_gini_5_30=knn_matrix_gini_5_30[1:2484,1:2484]

cordinate_gini=as.data.frame(Embeddings(giniclust_seurat, reduction = "umap")[, 1:2])
cordinate_gini=cordinate_gini[1:2484,]
```

```{r }
Plot_umap(cordinate_gini,general_gini)
```
```{r }
Plot_gene(norm_gini,cordinate_gini,"Zscan4c","Zscan4c")
Plot_gene(norm_gini,cordinate_gini,"Zscan4d","Zscan4d")
Plot_gene(norm_gini,cordinate_gini,"Tmem92","Tmem92")
```

```{r }
gini_cluster=read.csv2("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/published_data/13059_2016_1010_MOESM7_ESM.csv")
gini_name=gini_cluster$Cell.ID
gini_cluster=gini_cluster$Cluster.ID
Plot_umap(cordinate_gini,gini_cluster)
```

```{r }
setwd("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/published_data")
save(test_data_1,file="test_data_1.Rda")
save(norm_gini,file="norm_gini.Rda")
save(knn_matrix_gini_5_30,file="knn_matrix_gini_5_30.Rda")
save(cordinate_gini,file="cordinate_gini.Rda")
save(test_data_1,file="test_data_1.Rda")

```

## Start with entropy of mixing analysis
#Rscript --vanilla /home/ies/gabriele.lubatti/entropy_mixing/published_data/Entropy_hyper_server_all_cells.R -n norm_gini -k knn_matrix_gini_5_30 -o gini_hyper_old.Rda -r 1 -l 3 -h 20 -t 1 -p 0.001
#Time difference of 7.247149 mins
#cells 2484

##Optimize version for speed
#Rscript --vanilla /home/ies/gabriele.lubatti/entropy_mixing/published_data/Entropy_hyper_server_blood_new.R -n norm_gini -k knn_matrix_gini_5_30 -o gini_hyper_new.Rda -r 1 -l 3 -h 20 -t 1 -p 0.001
#Time difference of 1.864445 mins
#cells 2484

##New
```{r }
##LAST VERSION (setting srun -p normal_q -c 10  --mem=100G -t 1:00:00 --pty bash)
#Rscript --vanilla /home/ies/gabriele.lubatti/entropy_mixing/published_data/Entropy_server_december_2021_easy.R -n norm_gini -k knn_matrix_gini_5_30 -o gini_december_2021.Rda -r 1 -l 3 -h 20 -t 1 -p 0.001 -c 10 -a FALSE

#Time difference Time difference of 9.34686 mins


#Rscript --vanilla /home/ies/gabriele.lubatti/entropy_mixing/published_data/Entropy_server_december_2021_easy.R -n norm_gini -k knn_matrix_gini_5_30 -o gini_december_2021.Rda -r 1 -l 3 -h 20 -t 1 -p 0.001 -c 10 -a TRUE

#Time difference of 1.323001 mins



#Execution Time CIARA Approx: 42.54s
#Execution Time CIARA Full: 203.07s
```

##Old

```{r }

##LAST VERSION (setting srun -p normal_q -c 20  --mem=100G -t 1:00:00 --pty bash)
#Rscript --vanilla /home/ies/gabriele.lubatti/entropy_mixing/published_data/Entropy_server_december_2021_new.R -n norm_gini -k knn_matrix_gini_5_30 -o gini_december_2021.Rda -r 1 -l 3 -h 20 -t 1 -p 0.001 -c 10 -a FALSE

#Time difference of 12.84244 min

#Rscript --vanilla /home/ies/gabriele.lubatti/entropy_mixing/published_data/Entropy_server_december_2021_new.R -n norm_gini -k knn_matrix_gini_5_30 -o gini_december_2021_a.Rda -r 1 -l 3 -h 20 -t 1 -p 0.001 -c 10 -a TRUE

#Time difference of 1.565962 mins
```

```{r }
#No approximation
#is equal to gini hyper old
#setwd("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/published_data")
#load("gini_hyper.Rda")

```

```{r }
# With approximation
#setwd("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/published_data")
#vedo=load("gini_hyper_new.Rda")
#result_new=result
```

```{r }
#No approximation
setwd("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/published_data")
load("gini_hyper_old.Rda")

```



```{r }
# With approximation
setwd("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/published_data")
vedo=load("gini_hyper_new.Rda")
result_new=result
```



```{r }
setwd("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/published_data")
load("test_data_1.Rda")
load("norm_gini.Rda")
load(file="cordinate_gini.Rda")
load(file="knn_matrix_gini_5_30.Rda")
```

```{r }
setwd("/Users/gabriele.lubatti/Downloads")
write.csv(norm_gini,file="norm_gini.csv")
write.csv(knn_matrix_gini_5_30,file="knn_matrix_gini_5_30.csv")
```

```{r }
entropy_genes_new=row.names(result_new)[result_new[,1]<1]
#geni_top_new=row.names(result_new)[result_new[,2]<0.00001&result_new[,1]==0]
```

```{r }
entropy_genes=row.names(result)[result[,2]<1]
geni_top=row.names(result)[order(as.numeric(result[,1]))]


```

```{r }
background=row.names(result)
background_new=row.names(result_new)
```

```{r }


jacardi_entropy=find_jacardi_genes(entropy_genes,entropy_genes_new)
jacardi_entropy=find_jacardi_genes(background,background_new)
```




```{r }
library(ggplot2)
p=list()
for(i in (geni_top[1:10])){
  q=Plot_gene(norm_gini,cordinate_gini,i,i)
  p=list(p,q)
}
p
```



```{r }
Plot_genes_sum(cordinate_gini,norm_gini,(geni_top),"Sum from top entropy genes")

```


## Comparison for PGC: gini index vs entropy of mixing

```{r}
vedo=load(system.file("extdata", "experiment_name_StatisticsTable_afterLOESS_elmir.RData", package ="EntropyMixingR"))
#"ExprM.Stat2"   

setwd("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/elmir_data")
vedo=load("elmir_hyper_old.Rda")


entropy_genes=row.names(result)[result[,2]<1]

max_gini=ExprM.Stat2$log2.Maxs
gini_value=ExprM.Stat2$Gini
#max_gini=gini_index$log2..Max.Expression.Level..


label_all=c(rep("No Gini gene",length(max_gini)))
#label_all[ExprM.Stat2$Gini.pvalue<0.0001]="Gini gene"
label_all[row.names(ExprM.Stat2)%in%Genelist.top_pvalue]="Gini gene"

mtcars=data.frame(label_all,gini_value,max_gini)
pdf("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/github_january_2022/figures_paper/sup_fig_1_c.pdf")
ggplot(mtcars, aes(x=max_gini, y=gini_value, color=label_all)) +
  geom_point()+xlab("log2 (Max Expression Level)")+ylab("Gini")+labs(color="Label")

dev.off()

label_all=c(rep("No CIARA gene",length(max_gini)))
label_all[row.names(ExprM.Stat2)%in%entropy_genes]="CIARA gene"

mtcars=data.frame(label_all,gini_value,max_gini)
row.names(mtcars)=row.names(ExprM.Stat2)
mtcars=mtcars[c(row.names(ExprM.Stat2)[!row.names(ExprM.Stat2)%in%entropy_genes],row.names(ExprM.Stat2)[row.names(ExprM.Stat2)%in%entropy_genes]),]
pdf("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/github_january_2022/figures_paper/sup_fig_1_d.pdf")
ggplot(mtcars, aes(x=max_gini, y=gini_value, color=label_all)) +
  geom_point()+xlab("log2 (Max Expression Level)")+ylab("Gini")+labs(color="Label")
dev.off()

```



```{r }
library(CIARA)
#mayra_seurat_entropy_24=Cluster_analysis_integrate_rare(mayra_dati_raw_24,"24h",0.01,5,30,names(entro#py_genes))
gini_entropy_24=Cluster_analysis_integrate_rare(test_data_1,"24h",0.1,3,30,(entropy_genes))

#mayra_seurat_entropy_24=FindClusters(mayra_seurat_entropy_24,resolution = 0.1)

```




```{r }
library(clustree)
find_resolution(gini_entropy_24,seq(0.01,0.1,0.01))
```



```{r }

norm_gini_entropy=as.matrix(GetAssayData(gini_entropy_24, slot = "data",assay="RNA"))

norm_gini_entropy=norm_gini_entropy[,1:2484]

cordinate_gini_entropy=as.data.frame(Embeddings(gini_entropy_24, reduction = "umap")[, 1:2])

cordinate_gini_entropy=cordinate_gini_entropy[1:2484,]

cluster_entropy=as.vector(gini_entropy_24$RNA_snn_res.0.1)
cluster_entropy=cluster_entropy[1:2484]
```


```{r }
Plot_umap(cordinate_gini_entropy,cluster_entropy)

Plot_gene(norm_gini_entropy,cordinate_gini_entropy,"Zscan4d","Zscan4d")
```



```{r }
#Plot_umap(cordinate_umap_24_update,mayra_seurat_entropy_24$RNA_snn_res.0.1)


cluster_vedo=rep("Others",length(cluster_entropy))
cluster_vedo[cluster_entropy==6]="2CLC"


Plot_umap(cordinate_gini,cluster_vedo)
```

```{r }
gini_cluster=read.csv2("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/published_data/13059_2016_1010_MOESM7_ESM.csv")
gini_name=gini_cluster$Cell.ID
gini_cluster=gini_cluster$Cluster.ID
Plot_umap(cordinate_gini,gini_cluster)



```

```{r }
Plot_gene(norm_gini_entropy,cordinate_gini,"Tcstv1","Tcstv1")
```

## Gene used in the paper
```{r }
norm_gini_entropy=as.matrix(GetAssayData(gini_entropy_24, slot = "data",assay="RNA"))
geni_marker=c("Tcstv1","Dcdc2c","Zscan4f","Zscan4d")
positive_cells=find_positive_cells(norm_gini_entropy,geni_marker,threshold = 0)
```

```{r }

names(gini_entropy_24$RNA_snn_res.0.1)[gini_entropy_24$RNA_snn_res.0.1==6]
positive_cells




find_jacardi_new(as.vector(gini_entropy_24$RNA_snn_res.0.1),norm_gini_entropy,6,positive_cells)

find_jacardi_new(gini_cluster,norm_gini_entropy,"Cluster_2",positive_cells)


```




```{r }
#Singleton is labelled as others in entropy of mixing
which(cluster_vedo=="2CLC")
which(gini_cluster=="Cluster_2")
```





##Start third dataset GiniClust


## 576 cells
## rare cluster 9 cells 
## In the paper is figure 5b




## Pre processing

```{r}
#15,153 genes and 2545 cells.

test_data_3=read.csv("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/published_data/DATA_GBM.csv")
```

```{r }
filter_cells=apply(test_data_3,2,function(x){
  x=x[x>0]
  if(length(x)<2000){
    return (FALSE)
  }
  else {
    return(TRUE)
  }
})
```

```{r }
test_data_3=test_data_3[,filter_cells]
test_data_3=test_data_3[,1:472]
```








```{r }
giniclust_seurat_3=Cluster_analysis_integrate_rare(test_data_3,"GiniClust",0.1,5,30)
```


```{r }
norm_gini_3=as.matrix(GetAssayData(giniclust_seurat_3, slot = "data",assay="RNA"))
#norm_gini=norm_gini[,1:2484]

general_gini_3=giniclust_seurat_3$RNA_snn_res.0.1
#general_gini=general_gini[1:2484]

knn_matrix_gini_5_30_3=as.matrix(giniclust_seurat_3@graphs$RNA_nn)
#knn_matrix_gini_5_30=knn_matrix_gini_5_30[1:2484,1:2484]

cordinate_gini_3=as.data.frame(Embeddings(giniclust_seurat_3, reduction = "umap")[, 1:2])
#cordinate_gini=cordinate_gini[1:2484,]
```

```{r }
Plot_umap(cordinate_gini_3,general_gini_3)
Plot_umap(cordinate_gini_3,gini_cluster_3)
```

```{r }
gini_cluster_3=read.csv2("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/published_data/13059_2016_1010_MOESM11_ESM.csv")

gini_name_3=gini_cluster_3$Cell.ID
gini_cluster_3=gini_cluster_3$Cluster.ID
ordine=data.frame(gini_name_3,gini_cluster_3)
row.names(ordine)=gini_name_3
gini_cluster_3=ordine[colnames(test_data_3),2]
Plot_umap(cordinate_gini_3,gini_cluster_3)
```


```{r }
Plot_gene(norm_gini_3,cordinate_gini_3,"PLP1","PLP1")
```


```{r }
setwd("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/published_data")
save(test_data_3,file="test_data_3.Rda")
save(norm_gini_3,file="norm_gini_3.Rda")
save(knn_matrix_gini_5_30_3,file="knn_matrix_gini_5_30_3.Rda")
save(cordinate_gini_3,file="cordinate_gini_3.Rda")

```


## Entropy of mixing

#Rscript --vanilla /home/ies/gabriele.lubatti/entropy_mixing/published_data/Entropy_hyper_server.R -n norm_gini_3 -k knn_matrix_gini_5_30_3 -o gini_hyper_3.Rda -r 1 -l 3 -h 20 -t 1 -p 0.001



#Rscript --vanilla /home/ies/gabriele.lubatti/entropy_mixing/published_data/Entropy_hyper_server_all_cells.R -n norm_gini_3 -k knn_matrix_gini_5_30_3 -o gini_hyper_3_old.Rda -r 1 -l 3 -h 20 -t 1 -p 0.001
#Time difference of 22.9829 secs
#ncells: 472
#Rscript --vanilla /home/ies/gabriele.lubatti/entropy_mixing/published_data/Entropy_hyper_server_blood_new.R -n norm_gini_3 -k knn_matrix_gini_5_30_3 -o gini_hyper_3_new.Rda -r 1 -l 3 -h 20 -t 1 -p 0.001
#ncells: 472
#Time difference of 22.38227 secs

##New
```{r }
##LAST VERSION (setting srun -p normal_q -c 10  --mem=100G -t 1:00:00 --pty bash)



#Rscript --vanilla /home/ies/gabriele.lubatti/entropy_mixing/published_data/Entropy_server_december_2021_easy.R -n norm_gini_3 -k knn_matrix_gini_5_30_3 -o gini_3_december_2021.Rda -r 1 -l 3 -h 20 -t 1 -p 0.001 -c 10 -a FALSE

#Time difference of 26.66831 secs 



#Rscript --vanilla /home/ies/gabriele.lubatti/entropy_mixing/published_data/Entropy_server_december_2021_easy.R -n norm_gini_3 -k knn_matrix_gini_5_30_3 -o gini_3_december_2021_a.Rda -r 1 -l 3 -h 20 -t 1 -p 0.001 -c 10 -a TRUE
#Time difference of 14.02788 secs

#Execution Time CIARA Approx: 11.96s
#Execution Time CIARA Full: 22.26s

```

#Old
```{r }
##LAST VERSION (setting srun -p normal_q -c 20  --mem=100G -t 1:00:00 --pty bash)
#Rscript --vanilla /home/ies/gabriele.lubatti/entropy_mixing/published_data/Entropy_server_december_2021_new.R -n norm_gini_3 -k knn_matrix_gini_5_30_3 -o gini_3_december_2021.Rda -r 1 -l 3 -h 20 -t 1 -p 0.001 -c 10 -a FALSE

# Time difference of 27.92614 secs 

#Rscript --vanilla /home/ies/gabriele.lubatti/entropy_mixing/published_data/Entropy_server_december_2021_new.R -n norm_gini_3 -k knn_matrix_gini_5_30_3 -o gini_3_december_2021_a.Rda -r 1 -l 3 -h 20 -t 1 -p 0.001 -c 10 -a TRUE

#Time difference of 14.38749 secs

```

```{r }
setwd("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/published_data")
load("test_data_3.Rda")
load(file="norm_gini_3.Rda")
load(file="knn_matrix_gini_5_30_3.Rda")
load(file="cordinate_gini_3.Rda")

gini_cluster_3=read.csv2("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/published_data/13059_2016_1010_MOESM11_ESM.csv")

gini_name_3=gini_cluster_3$Cell.ID
gini_cluster_3=gini_cluster_3$Cluster.ID
ordine=data.frame(gini_name_3,gini_cluster_3)
row.names(ordine)=gini_name_3
gini_cluster_3=ordine[colnames(test_data_3),2]
```


```{r }
setwd("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/published_data")
load("gini_hyper_3_new.Rda")
result_new=result
```

```{r }
entropy_genes_new=row.names(result_new)[result_new[,1]<1]
#geni_top_new=row.names(result_new)[result_new[,2]<0.00001&result_new[,1]==0]
```

```{r }
setwd("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/published_data")
load("gini_hyper_3_old.Rda")
result_old=result
```

```{r }

entropy_genes_old=row.names(result_old)[result_old[,2]<1]
geni_top_old=row.names(result_old)[result_old[,1]<0.00001&result_old[,2]==0]

```

```{r }
jacardi_entropy=find_jacardi_genes(entropy_genes_old,entropy_genes_new)


```



```{r }
setwd("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/published_data")
load("gini_hyper_3.Rda")

```


```{r }
entropy_genes=row.names(result)[result[,2]<1]
geni_top=row.names(result)[result[,1]<0.00001&result[,2]==0]
```



```{r }
background=row.names(result)
#background=c(names(entropy_genes),names(entropy_genes_excluded))
```

```{r }
library(ggplot2)
p=list()
for(i in (geni_top)[1:10]){
  q=Plot_gene(norm_gini_3,cordinate_gini_3,i,i)
  p=list(p,q)
}
p
```


```{r }
Plot_genes_sum(cordinate_gini_3,norm_gini_3,(geni_top),"Sum from top entropy genes")
```


```{r }
#mayra_seurat_entropy_24=Cluster_analysis_integrate_rare(mayra_dati_raw_24,"24h",0.01,5,30,names(entro#py_genes))
gini_entropy_24_3=Cluster_analysis_integrate_rare(test_data_3,"24h",0.1,3,30,(entropy_genes))

#mayra_seurat_entropy_24=FindClusters(mayra_seurat_entropy_24,resolution = 0.1)

```




```{r }
library(clustree)
find_resolution(gini_entropy_24_3,seq(0.01,0.1,0.01))
```



```{r }

norm_gini_entropy_3=as.matrix(GetAssayData(gini_entropy_24_3, slot = "data",assay="RNA"))



cordinate_gini_entropy_3=as.data.frame(Embeddings(gini_entropy_24_3, reduction = "umap")[, 1:2])



cluster_entropy_3=as.vector(gini_entropy_24_3$RNA_snn_res.0.1)

```


```{r }
Plot_umap(cordinate_gini_entropy_3,cluster_entropy_3)
Plot_umap(cordinate_gini_3,cluster_entropy_3)

cluster_entropy_new=rep("other",length(cluster_entropy_3))
cluster_entropy_new[cluster_entropy_3==10]="Rare cluster"
Plot_umap(cordinate_gini_3,cluster_entropy_new)

```


```{r }

cluster_entropy_new=rep("other",length(cluster_entropy_3))
cluster_entropy_new[cluster_entropy_3==6]="Rare cluster2"
Plot_umap(cordinate_gini_3,cluster_entropy_new)

cluster_entropy_new=rep("other",length(cluster_entropy_3))
cluster_entropy_new[cluster_entropy_3==10]="Rare cluster1"
Plot_umap(cordinate_gini_3,cluster_entropy_new)




```



##Gene used in the paper
```{r }
norm_gini_entropy_3=as.matrix(GetAssayData(gini_entropy_24_3, slot = "data",assay="RNA"))
geni_marker=c("CLDN11","MBP","PLP1","KLK6")
positive_cells=find_positive_cells(norm_gini_entropy_3,geni_marker,threshold = 0)
```

```{r }

names(gini_entropy_24_3$RNA_snn_res.0.1)[gini_entropy_24_3$RNA_snn_res.0.1==10]
positive_cells




find_jacardi_new(as.vector(gini_entropy_24_3$RNA_snn_res.0.1),norm_gini_entropy_3,10,positive_cells)


find_jacardi_new(gini_cluster_3,norm_gini_entropy_3,"Cluster_3",positive_cells)

```



## End third dataset from GiniClust paper 

## Start dataset from FiRE paper fig. 3B

## Pre processing
```{r }
setwd("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/published_data")
label_fire=read.table("labels_jurkat_two_species_1580.txt")
fire_data=read.table("preprocessedData_jurkat_two_species_1580.txt")
label_fire=read.table("labels_jurkat_two_species_1580.txt")
```


```{r }
data <- read.table(gzfile('/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/published_data/FiRE-master/data/jurkat_two_species_1580.txt.gz'))
labels <- read.table('/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/published_data/FiRE-master/data/labels_jurkat_two_species_1580.txt') #Cells with label '1' represent abundant, while cells with label '2' represent rare.

#Genes
genes <- c(1:dim(data)[2])
#data <- t(data) #Samples * Features

colnames(data)=genes



for ( i in 1:length(row.names(data))){
  
row.names(data)[i]=paste0("cell",i)
}
```

```{r }
fire_seurat=Cluster_analysis_integrate_rare(data,"Fire",0.1,5,30)
```


```{r }
norm_fire=as.matrix(GetAssayData(fire_seurat, slot = "data",assay="RNA"))
#norm_gini=norm_gini[,1:2484]

general_fire=fire_seurat$RNA_snn_res.0.1
#general_gini=general_gini[1:2484]

knn_matrix_fire=as.matrix(fire_seurat@graphs$RNA_nn)
#knn_matrix_gini_5_30=knn_matrix_gini_5_30[1:2484,1:2484]

cordinate_fire=as.data.frame(Embeddings(fire_seurat, reduction = "umap")[, 1:2])
#cordinate_gini=cordinate_gini[1:2484,]
```

```{r }
labels=labels$V1
Plot_umap(cordinate_fire,general_fire)
Plot_umap(cordinate_fire,labels)
```

```{r }
setwd("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/published_data")
save(data,file="data.Rda")
save(norm_fire,file="norm_fire.Rda")
save(knn_matrix_fire,file="knn_matrix_fire.Rda")
save(cordinate_fire,file="cordinate_fire.Rda")
```


## Start Entropy of mixing

#Rscript --vanilla /home/ies/gabriele.lubatti/entropy_mixing/published_data/Entropy_hyper_server.R -n norm_fire -k knn_matrix_fire -o fire_hyper.Rda -r 1 -l 3 -h 20 -t 1 -p 0.001

#Rscript --vanilla /home/ies/gabriele.lubatti/entropy_mixing/published_data/Entropy_hyper_server_all_cells.R -n norm_fire -k knn_matrix_fire -o fire_hyper_old.Rda -r 1 -l 3 -h 20 -t 1 -p 0.001
#ncells:1580
#Time difference of 5.27283 mins

#Rscript --vanilla /home/ies/gabriele.lubatti/entropy_mixing/published_data/Entropy_hyper_server_blood_new.R -n norm_fire -k knn_matrix_fire -o fire_hyper_new.Rda -r 1 -l 3 -h 20 -t 1 -p 0.001
#ncells:1580
#Time difference of 2.156226 mins


##new
##New
```{r }
##LAST VERSION (setting srun -p normal_q -c 10  --mem=100G -t 1:00:00 --pty bash)

#Rscript --vanilla /home/ies/gabriele.lubatti/entropy_mixing/published_data/Entropy_server_december_2021_easy.R -n norm_fire -k knn_matrix_fire -o fire_december_2021.Rda -r 1 -l 3 -h 20 -t 1 -p 0.001 -c 10 -a FALSE

#Time difference of 5.016574 mins


#Rscript --vanilla /home/ies/gabriele.lubatti/entropy_mixing/published_data/Entropy_server_december_2021_easy.R -n norm_fire -k knn_matrix_fire -o fire_december_2021_a.Rda -r 1 -l 3 -h 20 -t 1 -p 0.001 -c 10 -a TRUE

#Time difference of 1.20237 mins


#Execution Time CIARA Approx: 51.29s
#Execution Time CIARA Full: 176.15s
```



##Old
```{r }
##LAST VERSION (setting srun -p normal_q -c 20  --mem=100G -t 1:00:00 --pty bash)
#Rscript --vanilla /home/ies/gabriele.lubatti/entropy_mixing/published_data/Entropy_server_december_2021_new.R -n norm_fire -k knn_matrix_fire -o fire_december_2021.Rda -r 1 -l 3 -h 20 -t 1 -p 0.001 -c 10 -a FALSE

# Time difference of 5.809975 mins


#Rscript --vanilla /home/ies/gabriele.lubatti/entropy_mixing/published_data/Entropy_server_december_2021_new.R -n norm_fire -k knn_matrix_fire -o fire_december_2021_a.Rda -r 1 -l 3 -h 20 -t 1 -p 0.001 -c 10 -a TRUE


#Time difference of 1.389862 mins

```


```{r }
setwd("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/published_data")
load(file="data.Rda")
load(file="norm_fire.Rda")
load(file="knn_matrix_fire.Rda")
load(file="cordinate_fire.Rda")
setwd("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/published_data")
label_fire=read.table("labels_jurkat_two_species_1580.txt")
labels <- read.table('/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/published_data/FiRE-master/data/labels_jurkat_two_species_1580.txt')
labels=labels$V1
```

```{r }
setwd("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/published_data")
load("fire_hyper_new.Rda")
result_new=result
```

```{r }
entropy_genes_new=row.names(result_new)[result_new[,1]<1]
#geni_top_new=row.names(result_new)[result_new[,2]<0.00001&result_new[,1]==0]
```



```{r }
setwd("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/published_data")
load("fire_hyper_old.Rda")
result_old=result
```

```{r }
entropy_genes_old=row.names(result_old)[result_old[,2]<1]
#geni_top_old=row.names(result_old)[result_old[,1]<0.00001&result_old[,2]==0]
```

```{r }
jacardi_entropy=find_jacardi_genes(entropy_genes_old,entropy_genes_new)
#jacardi_entropy=find_jacardi_genes(background_old,background_new)

```

```{r }
setwd("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/published_data")
load("fire_hyper.Rda")

```


```{r }
entropy_genes=row.names(result)[result[,2]<1]
geni_top=row.names(result)[result[,1]<0.00001&result[,2]==0]
```



```{r }
background=row.names(result)
#background=c(names(entropy_genes),names(entropy_genes_excluded))
```

```{r }
library(ggplot2)
p=list()
for(i in (geni_top)[1:10]){
  q=Plot_gene(norm_fire,cordinate_fire,i,i)
  p=list(p,q)
}
p
```


```{r}
library(IDENTOM)
Plot_genes_sum(cordinate_fire,norm_fire,(geni_top),"Sum from top entropy genes")

```


```{r }
#mayra_seurat_entropy_24=Cluster_analysis_integrate_rare(mayra_dati_raw_24,"24h",0.01,5,30,names(entro#py_genes))
fire_entropy=Cluster_analysis_integrate_rare(data,"24h",0.1,3,30,(entropy_genes))

#mayra_seurat_entropy_24=FindClusters(mayra_seurat_entropy_24,resolution = 0.1)

```




```{r }
library(clustree)
find_resolution(fire_entropy,seq(0.1,0.3,0.1))
```

```{r }

norm_fire_entropy=as.matrix(GetAssayData(fire_entropy, slot = "data",assay="RNA"))



cordinate_fire_entropy=as.data.frame(Embeddings(fire_entropy, reduction = "umap")[, 1:2])



cluster_fire_entropy=as.vector(fire_entropy$RNA_snn_res.0.1)

```


```{r }
Plot_umap(cordinate_fire_entropy,cluster_fire_entropy)

pdf("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/github_january_2022/figures_paper/figure_2/fig_2_c.pdf")
Plot_umap(cordinate_fire,cluster_fire_entropy)
dev.off()

Plot_umap(cordinate_fire,labels)


cluster_fire_entropy_plot=labels
cluster_fire_entropy_plot[labels==1]="Abundant (293T)"
cluster_fire_entropy_plot[labels==2]="Rare (Jurkat)"
pdf("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/github_january_2022/figures_paper/figure_2/fig_2_b.pdf")
Plot_umap(cordinate_fire,cluster_fire_entropy_plot)
dev.off()

cluster_fire_entropy_plot=labels
cluster_fire_entropy_plot[labels==1]="Abundant (293T)"
cluster_fire_entropy_plot[labels==2]="Rare (Jurkat)"
Plot_umap(cordinate_fire,cluster_fire_entropy_plot)
```







```{r }
table(cluster_fire_entropy,labels)

find_jacardi(cluster_fire_entropy,labels,norm_fire,1,2)
```




```{r }
tsne_paper=read.table('/Users/gabriele.lubatti/Downloads/FiRE-master/data/tsne_jurkat_1k_outlierRemoved.txt') 

```


##Comparison with FiRE classification

```{r }
source(system.file("extdata", "preprocess.R", package ="EntropyMixingR"))
library(FiRE)
data_example <- t(data)
genes <- colnames(data_example)#It can be replaced with original gene names

data_mat <- list(mat=data_example, gene_symbols=genes)

preprocessedList <- ranger_preprocess(data_mat)
preprocessedData <- as.matrix(preprocessedList$preprocessedData)

model <- new(FiRE::FiRE, 100, 50, 1017881, 5489, 0)

model$fit(preprocessedData)

score <- model$score(preprocessedData)
q3 <- quantile(score, 0.75)
iqr <- IQR(score)
th <- q3 + (1.5*iqr)
#th_new=q3 + (0.5*iqr)
indIqr <- which(score >= th)
predictions <- integer(dim(data_example)[1])
predictions[indIqr] <- 1 

#model$b
```

```{r }
library("mltools")
#mcc()
#Matthews correlation coefficient 
#methods for quantifying agreement between binary classification (from GiniClust2 paper) (values from -1 to +1)
#preds <- c(1,1,1,0,1,1,0,0)
#actuals <- c(1,1,1,1,0,0,0,0)
preds=as.numeric(cluster_fire_entropy)
actuals=as.numeric(labels)
actuals_new=actuals
actuals_new[actuals==2]=1
actuals_new[actuals==1]=0

#with entropy of mixing
mcc(preds, actuals_new)


#with FiRE

mcc(predictions, actuals_new)
```






## End FiRE

## Start with new dataset from CellSius 

## Pre processing
```{r }
#setwd("/Users/gabriele.lubatti/Downloads/CellSIUS-master/data")
#bo_3=load("my_tsne.RData")
#bo_2=load("clusters.RData")
#bo=load("norm.counts.RData")
#bo_2=load("clusters.RData")
#Plot_umap(as.data.frame(my_tsne),as.vector(clusters))

```

```{r }
setwd("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/published_data")
vedo=load("Supplementary_File_8_sce_raw.RData")
```


```{r }
setwd("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/published_data")
vedo=load("Supplementary_File_10_subset_2_idx.RData")
```
#First approach (SKIP)
```{r }
raw_sce=sce@assayData$counts
cell_cycle=sce$cell_cycle_phase
cell_line=sce$cell_line

raw_sce_small=raw_sce[,cell_cycle=="G1"]
cell_line=cell_line[cell_cycle=="G1"]
hek293=colnames(raw_sce_small)[cell_line=="HEK293"]
ramos=colnames(raw_sce_small)[cell_line=="Ramos"]
jurkat=colnames(raw_sce_small)[cell_line=="Jurkat"]
set.seed(1)
hek293_small=sample(hek293,100)
ramos_small=sample(ramos,100)
jurkat_small=sample(jurkat,2)

cell_subset=c(hek293_small,ramos_small,jurkat_small)
cell_line_1=rep("hek293",(100))
cell_line_2=rep("ramos",(100))
cell_line_3=rep("jurkat",(2))

cell_line_all=c(cell_line_1,cell_line_2,cell_line_3)

raw_sce_small=raw_sce_small[,cell_subset]



```

##Second approach (Yes)
```{r }
raw_sce=sce@assayData$counts
raw_sce_small=raw_sce[,idx]
cell_line_all=sce$cell_line[idx]
```

```{r }
cellsius_seurat=Cluster_analysis_integrate_rare(raw_sce_small,"GiniClust",0.1,5,30)
```


```{r }
norm_cellsius=as.matrix(GetAssayData(cellsius_seurat, slot = "data",assay="RNA"))
#norm_gini=norm_gini[,1:2484]

general_cellsius=cellsius_seurat$RNA_snn_res.0.1
#general_gini=general_gini[1:2484]

knn_matrix_cellsius=as.matrix(cellsius_seurat@graphs$RNA_nn)
#knn_matrix_gini_5_30=knn_matrix_gini_5_30[1:2484,1:2484]

cordinate_cellsius=as.data.frame(Embeddings(cellsius_seurat, reduction = "umap")[, 1:2])
#cordinate_gini=cordinate_gini[1:2484,]
```


```{r }
Plot_umap(cordinate_cellsius,general_cellsius)
Plot_umap(cordinate_cellsius,cell_line_all)
```

```{r }

setwd("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/published_data")
save(raw_sce_small,file="raw_sce_small.Rda")
save(cordinate_cellsius,file="cordinate_cellsius.Rda")
save(cell_line_all,file="cell_line_all.Rda")
save(norm_cellsius,file="norm_cellsius.Rda")
save(knn_matrix_cellsius,file="knn_matrix_cellsius.Rda")

```



```{r}
library(Biobase)
DefaultAssay(cellsius_seurat) <- "RNA"
markers_first_gini=markers_first_method(cellsius_seurat,cell_line_all,names(cellsius_seurat$RNA_snn_res.0.1),20)
```

```{r}
markers_first_elmir=markers_first_gini
norm_mayra_entropy_24=norm_gini
final_cluster_version_sub=cluster_vedo
library(ComplexHeatmap)
library(circlize)
marker_plot_day_24_first=markers_first_elmir[[1]]

marker_plot_plot_day_24_first=markers_first_elmir[[2]]
marker_all_day_24_first=markers_first_elmir[[3]]
condition_day_24=rep("24h",length(colnames(norm_mayra_entropy_24)))
toMatch=seq(0,11,1)
final_cluster=final_cluster_version_sub
new_cluster_plot=final_cluster[grep(paste(toMatch,collapse="|"),final_cluster)]
norm_elmir_plot=norm_mayra_entropy_24[,grep(paste(toMatch,collapse="|"),final_cluster)]
condition_day_24=rep("24h",length(colnames(norm_elmir_plot)))
marker_plot_day_24_first_plot=marker_plot_day_24_first[grep(paste(toMatch,collapse="|"),names(marker_plot_day_24_first))]
marker_plot_plot_day_24_first_plot=marker_plot_plot_day_24_first[grep(paste(toMatch,collapse="|"),marker_plot_plot_day_24_first)]
#heatmap_marker_all_new(marker_plot_day_24_first_plot,marker_plot_plot_day_24_first_plot,new_cluster_p#lot,condition_day_24,norm_elmir_plot,3)


```

## Start with entropy of mixing


#Rscript --vanilla /home/ies/gabriele.lubatti/entropy_mixing/published_data/Entropy_hyper_server.R -n norm_cellsius -k knn_matrix_cellsius -o cellsius_hyper.Rda -r 1 -l 3 -h 20 -t 1 -p 0.001

#Rscript --vanilla /home/ies/gabriele.lubatti/entropy_mixing/published_data/Entropy_hyper_server_all_cells.R -n norm_cellsius -k knn_matrix_cellsius -o cellsius_hyper_old.Rda -r 1 -l 3 -h 20 -t 1 -p 0.001

#ncells:3984
#Time difference of 12.28497 mins

#Rscript --vanilla /home/ies/gabriele.lubatti/entropy_mixing/published_data/Entropy_hyper_server_blood_new.R -n norm_cellsius -k knn_matrix_cellsius -o cellsius_hyper_new.Rda -r 1 -l 3 -h 20 -t 1 -p 0.001
#ncells:3984
# Time difference of 3.798013 mins

##New
```{r }
##LAST VERSION (setting srun -p normal_q -c 10  --mem=100G -t 1:00:00 --pty bash)


#Rscript --vanilla /home/ies/gabriele.lubatti/entropy_mixing/published_data/Entropy_server_december_2021_easy.R -n norm_cellsius -k knn_matrix_cellsius -o cellsius_december_2021.Rda -r 1 -l 3 -h 20 -t 1 -p 0.001 -c 10 -a FALSE

#Time difference of 17.8528 min

#Rscript --vanilla /home/ies/gabriele.lubatti/entropy_mixing/published_data/Entropy_server_december_2021_easy.R -n norm_cellsius -k knn_matrix_cellsius -o cellsius_december_2021_a.Rda -r 1 -l 3 -h 20 -t 1 -p 0.001 -c 10 -a TRUE

#Time difference of 2.947906 mins


#Execution Time CIARA Approx: 119.78s
#Execution Time CIARA Full: 462.7s
```

#Old
```{r }
##LAST VERSION (setting srun -p normal_q -c 20  --mem=100G -t 1:00:00 --pty bash)
#Rscript --vanilla /home/ies/gabriele.lubatti/entropy_mixing/published_data/Entropy_server_december_2021_new.R -n norm_cellsius -k knn_matrix_cellsius -o cellsius_december_2021.Rda -r 1 -l 3 -h 20 -t 1 -p 0.001 -c 10 -a FALSE

# Time difference of 21.26795 mins 

#Rscript --vanilla /home/ies/gabriele.lubatti/entropy_mixing/published_data/Entropy_server_december_2021_new.R -n norm_cellsius -k knn_matrix_cellsius -o cellsius_december_2021_a.Rda -r 1 -l 3 -h 20 -t 1 -p 0.001 -c 10 -a TRUE


#Time difference of 3.951076 mins


```

```{r }
setwd("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/published_data")
load(file="raw_sce_small.Rda")
load(file="cordinate_cellsius.Rda")
load(file="cell_line_all.Rda")
load(file="norm_cellsius.Rda")
load(file="knn_matrix_cellsius.Rda")

```

```{r }
setwd("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/published_data")
load("cellsius_hyper_new.Rda")
result_new=result
```


```{r }
entropy_genes_new=row.names(result_new)[result_new[,1]<1]
#geni_top_new=row.names(result_new)[result_new[,2]<0.00001&result_new[,1]==0]
```



```{r }
setwd("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/published_data")
load("cellsius_hyper_old.Rda")
result_old=result
```

```{r }
entropy_genes_old=row.names(result_old)[result_old[,2]<1]
#geni_top_old=row.names(result_old)[result_old[,1]<0.00001&result_old[,2]==0]
```

```{r }
jacardi_entropy=find_jacardi_genes(entropy_genes_old,entropy_genes_new)
#jacardi_entropy=find_jacardi_genes(background_old,background_new)

```

```{r }
setwd("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/published_data")
load("cellsius_hyper.Rda")

```




```{r }
entropy_genes=row.names(result)[result[,2]<1]
geni_top=row.names(result)[result[,1]<0.00001&result[,2]==0]

#geni_top=row.names(result)[result[,1]<0.0001]
```



```{r }
background=row.names(result)
```

```{r }
library(ggplot2)
p=list()
for(i in (geni_top)[1:10]){
  q=Plot_gene(norm_cellsius,cordinate_cellsius,i,i)
  p=list(p,q)
}
p
```


```{r }
Plot_genes_sum(cordinate_cellsius,norm_cellsius,(geni_top),"Sum from top entropy genes")

```


```{r }
cellsius_normal=Cluster_analysis_integrate_rare(raw_sce_small,"24h",0.1,3,30)

```


```{r }
#mayra_seurat_entropy_24=Cluster_analysis_integrate_rare(mayra_dati_raw_24,"24h",0.01,5,30,names(entro#py_genes))
cellsius_entropy=Cluster_analysis_integrate_rare(raw_sce_small,"24h",0.1,3,30,(entropy_genes))

#mayra_seurat_entropy_24=FindClusters(mayra_seurat_entropy_24,resolution = 0.1)

```




```{r }
library(clustree)
find_resolution(cellsius_entropy,seq(0.1,1,0.1))
```

```{r }

norm_cellsius_entropy=as.matrix(GetAssayData(cellsius_entropy, slot = "data",assay="RNA"))



cordinate_cellsius_entropy=as.data.frame(Embeddings(cellsius_entropy, reduction = "umap")[, 1:2])

#cellsius_entropy=FindClusters(cellsius_entropy,resolution = 0.6)

cluster_cellsius_entropy=as.vector(cellsius_entropy$RNA_snn_res.0.1)

```


```{r }
Plot_umap(cordinate_cellsius_entropy,cluster_cellsius_entropy)


Plot_umap(cordinate_cellsius,cluster_cellsius_entropy)

#Plot_umap(cordinate_cellsius_entropy,cell_line_all)

pdf("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/github_january_2022/figures_paper/figure_2/fig_2_e.pdf")
Plot_umap(cordinate_cellsius,cell_line_all)
dev.off()
pdf("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/github_january_2022/figures_paper/figure_2/fig_2_f.pdf")
Plot_umap(cordinate_cellsius,cluster_cellsius_entropy)
dev.off()


```



```{r }
table(cluster_cellsius_entropy,cell_line_all)

find_jacardi(cluster_cellsius_entropy,cell_line_all,norm_cellsius,8,"H1437")

find_jacardi(cluster_cellsius_entropy,cell_line_all,norm_cellsius,7,"Jurkat")
```







```{r }
library("mltools")
#mcc()
#Matthews correlation coefficient 
#methods for quantifying agreement between binary classification (from GiniClust2 paper) (values from -1 to +1)
#preds <- c(1,1,1,0,1,1,0,0)
#actuals <- c(1,1,1,1,0,0,0,0)
preds=as.numeric(cluster_cellsius_entropy)[cluster_cellsius_entropy==7|cluster_cellsius_entropy==8]
actuals=as.numeric(cell_line_all)[cluster_cellsius_entropy==7|cluster_cellsius_entropy==8]

actuals_new=actuals
actuals_new[actuals==8]=7
actuals_new[actuals==5]=8

#with entropy of mixing
mcc(preds, actuals_new)





```






## End Cellsius 
## Start Race ID



## Pre processing
```{r }
library(RaceID)
sc <- SCseq(intestinalData)
sc <- filterdata(sc,mintotal=2000)
fdata <- getfdata(sc)
sc <- compdist(sc,metric="pearson")
sc <- clustexp(sc,cln=7,sat=FALSE)
sc <- findoutliers(sc)
sc <- compumap(sc)
plotmap(sc,um=TRUE)

cluster_race_id=(sc@cpart)
```


```{r }
race_id_raw=sc@expdata
race_id_raw=race_id_raw[,names(cluster_race_id)]
```

```{r }
race_seurat=Cluster_analysis_integrate_rare(race_id_raw,"GiniClust",0.1,5,30)
```


```{r }
norm_race=as.matrix(GetAssayData(race_seurat, slot = "data",assay="RNA"))
#norm_gini=norm_gini[,1:2484]

general_race=race_seurat$RNA_snn_res.0.1
#general_gini=general_gini[1:2484]

knn_matrix_race=as.matrix(race_seurat@graphs$RNA_nn)
#knn_matrix_gini_5_30=knn_matrix_gini_5_30[1:2484,1:2484]

cordinate_race=as.data.frame(Embeddings(race_seurat, reduction = "umap")[, 1:2])
#cordinate_gini=cordinate_gini[1:2484,]
```

```{r }
Plot_umap(cordinate_race,general_race)
Plot_umap(cordinate_race,cluster_race_id)
```

```{r }

setwd("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/published_data")
save(race_id_raw,file="race_id_raw.Rda")
save(cordinate_race,file="cordinate_race.Rda")
save(cluster_race_id,file="cluster_race_id.Rda")
save(general_race,file="general_race.Rda")
save(norm_race,file="norm_race.Rda")
save(knn_matrix_race,file="knn_matrix_race.Rda")


```



## Start Entropy of mixing

#Rscript --vanilla /home/ies/gabriele.lubatti/entropy_mixing/published_data/Entropy_hyper_server.R -n norm_race -k knn_matrix_race -o race_hyper.Rda -r 1 -l 3 -h 20 -t 1 -p 0.001


#Rscript --vanilla /home/ies/gabriele.lubatti/entropy_mixing/published_data/Entropy_hyper_server_all_cells.R -n norm_race -k knn_matrix_race -o race_hyper_old.Rda -r 1 -l 3 -h 20 -t 1 -p 0.001

#ncells:317
#Time difference of 45.14149 secs

#Rscript --vanilla /home/ies/gabriele.lubatti/entropy_mixing/published_data/Entropy_hyper_server_blood_new.R -n norm_race -k knn_matrix_race -o race_hyper_new.Rda -r 1 -l 3 -h 20 -t 1 -p 0.001
#ncells:317
#Time difference of 18.67952 secs



##New
```{r }
##LAST VERSION (setting srun -p normal_q -c 10  --mem=100G -t 1:00:00 --pty bash)

#Rscript --vanilla /home/ies/gabriele.lubatti/entropy_mixing/published_data/Entropy_server_december_2021_easy.R -n norm_race -k knn_matrix_race -o race_december_2021.Rda -r 1 -l 3 -h 20 -t 1 -p 0.001 -c 10 -a FALSE
#Time difference of 38.31422 secs


#Rscript --vanilla /home/ies/gabriele.lubatti/entropy_mixing/published_data/Entropy_server_december_2021_easy.R -n norm_race -k knn_matrix_race -o race_december_2021_a.Rda -r 1 -l 3 -h 20 -t 1 -p 0.001 -c 10 -a TRUE

#Time difference of 10.31625 secs


#from pyhton
#Execution Time CIARA Approx: 4.31s
#Execution Time CIARA Full: 15.93s

```


##Old

```{r }
##LAST VERSION (setting srun -p normal_q -c 20  --mem=100G -t 1:00:00 --pty bash)
#Rscript --vanilla /home/ies/gabriele.lubatti/entropy_mixing/published_data/Entropy_server_december_2021_new.R -n norm_race -k knn_matrix_race -o race_december_2021.Rda -r 1 -l 3 -h 20 -t 1 -p 0.001 -c 10 -a FALSE

# Time difference of 1.190571 mins 

#Rscript --vanilla /home/ies/gabriele.lubatti/entropy_mixing/published_data/Entropy_server_december_2021_new.R -n norm_race -k knn_matrix_race -o race_december_2021_a.Rda -r 1 -l 3 -h 20 -t 1 -p 0.001 -c 10 -a TRUE


#Time difference of 19.85399 secs


```


```{r }
setwd("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/published_data")
load(file="race_id_raw.Rda")
load(file="norm_race.Rda")
load(file="knn_matrix_race.Rda")
load(file="cordinate_race.Rda")
load(file="cluster_race_id.Rda")
load(file="general_race.Rda")
```



```{r }
setwd("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/published_data")
load("race_hyper_new.Rda")
result_new=result
```
```{r }
entropy_genes_new=row.names(result_new)[result_new[,1]<1]
#geni_top_new=row.names(result_new)[result_new[,2]<0.00001&result_new[,1]==0]
```



```{r }
setwd("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/published_data")
load("race_hyper_old.Rda")
result_old=result
```

```{r }
entropy_genes_old=row.names(result_old)[result_old[,2]<1]
#geni_top_old=row.names(result_old)[result_old[,1]<0.00001&result_old[,2]==0]
```

```{r }
jacardi_entropy=find_jacardi_genes(entropy_genes_old,entropy_genes_new)
#jacardi_entropy=find_jacardi_genes(background_old,background_new)

```




```{r }
setwd("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/published_data")
load("race_hyper.Rda")

```


```{r }
entropy_genes=row.names(result)[result[,2]<1]
geni_top=row.names(result)[result[,1]<0.00001&result[,2]==0]

#geni_top=row.names(result)[result[,1]<0.0001]
```



```{r }
background=row.names(result)
#background=c(names(entropy_genes),names(entropy_genes_excluded))
```

```{r }
library(ggplot2)
p=list()
for(i in (geni_top)[1:10]){
  q=Plot_gene(norm_race,cordinate_race,i,i)
  p=list(p,q)
}
p
```


```{r }
Plot_genes_sum(cordinate_race,norm_race,(geni_top),"Sum from top entropy genes")


```

```{r }
#mayra_seurat_entropy_24=Cluster_analysis_integrate_rare(mayra_dati_raw_24,"24h",0.01,5,30,names(entro#py_genes))
race_entropy_1=Cluster_analysis_integrate_rare(race_id_raw,"24h",0.2,3,20,(entropy_genes))
#race_entropy_2=Cluster_analysis_integrate_rare(race_id_raw,"24h",0.2,3,30,(entropy_genes))

#mayra_seurat_entropy_24=FindClusters(mayra_seurat_entropy_24,resolution = 0.1)

```

```{r }
#table(race_entropy_1$RNA_snn_res.0.2,race_entropy_2$RNA_snn_res.0.2)
```



```{r }
library(clustree)
find_resolution(race_entropy_1,seq(0.1,1,0.1))
```

```{r }
race_entropy=race_entropy_1
norm_race_entropy=as.matrix(GetAssayData(race_entropy, slot = "data",assay="RNA"))



cordinate_race_entropy=as.data.frame(Embeddings(race_entropy, reduction = "umap")[, 1:2])

#cellsius_entropy=FindClusters(cellsius_entropy,resolution = 0.6)

cluster_race_entropy=as.vector(race_entropy$RNA_snn_res.0.2)

```


```{r }
Plot_umap(cordinate_race_entropy,cluster_race_entropy)
Plot_umap(cordinate_race_entropy,cluster_race_id)

#Plot_umap(cordinate_cellsius_entropy,cell_line_all)

Plot_umap(cordinate_race,cluster_race_id)
Plot_umap(cordinate_race,cluster_race_entropy)



```

```{r }

table(cluster_race_entropy,cluster_race_id)

```

```{r }


Plot_gene(norm_race_entropy,cordinate_race,"Apoa1","Apoa1")
Plot_gene(norm_race_entropy,cordinate_race,"Apoa4","Apoa4")
Plot_gene(norm_race_entropy,cordinate_race,"Fabp1","Fabp1")
Plot_gene(norm_race_entropy,cordinate_race,"Fabp2","Fabp2")
Plot_gene(norm_race_entropy,cordinate_race,"Fabp2","Fabp2")


Plot_gene(norm_race_entropy,cordinate_race,"Chga","Chga")
Plot_gene(norm_race_entropy,cordinate_race,"Chgb","Chgb")
Plot_gene(norm_race_entropy,cordinate_race,"Clca3","Clca3")
Plot_gene(norm_race_entropy,cordinate_race,"Clca4","Clca4")
Plot_gene(norm_race_entropy,cordinate_race,"Defa24","Defa24") 

Plot_gene(norm_race_entropy,cordinate_race,"Lgr5","Lgr5")

Plot_gene(norm_race_entropy,cordinate_race,"Ascl2","Ascl2")

Plot_gene(norm_race_entropy,cordinate_race,"Dclk1","Dclk1")
Plot_gene(norm_race_entropy,cordinate_race,"Nol3","Nol3")
Plot_gene(norm_race_entropy,cordinate_race,"Lpl","Lpl")
Plot_gene(norm_race_entropy,cordinate_race,"Emx2os","Emx2os")



```

RaceID2 recovered a larger group of Lgr5+ stem cells (cluster 2) and early progeny (clusters 1 and 8) as well as the major mature cell types; i.e., enterocytes (cluster 3), goblet (clusters 4 and 19), Paneth (clusters 5 and 6), and enteroendocrine cells (cluster 7) (Figures 1C and 1D)


Apoe1: enterocytes ; Chga: mature enteroendocrine cells; Chgb: early and mature enteroendocrine cells; Clca3: Goblet cells; Clca4: crypt bottom columnar cells; Defa24: Paneth cells (also Lgr5)

intestinal stem cell markers Lgr5, Clca4, and Ascl2

##Original cluster classfication (from paper is not the same as cluster annotation provided in raceid R library)
enterocytes (cluster 3), goblet (clusters 4 and 19), Paneth (clusters 5 and 6), and enteroendocrine cells (cluster 7) 


##From race id vignette by eyes
enterocytes (cluster 4), goblet (cluster 6,10,8) Paneth (cluster 7) and enteroendocrine cells (cluster 5)




```{r}
library(Biobase)
DefaultAssay(race_entropy) <- "RNA"
markers_first_gini=markers_first_method(race_entropy,cluster_race_entropy,names(race_entropy$RNA_snn_res.0.2),30)
```

```{r}
markers_first_elmir=markers_first_gini

marker_plot_day_24_first=markers_first_elmir[[1]]

marker_plot_plot_day_24_first=markers_first_elmir[[2]]
marker_all_day_24_first=markers_first_elmir[[3]]
#heatmap_marker_all_new(marker_plot_day_24_first_plot,marker_plot_plot_day_24_first_plot,new_cluster_p#lot,condition_day_24,norm_elmir_plot,3)


```

##Cluster 2 is Goblet cells

```{r}
marker_all_day_24_first[marker_all_day_24_first=="Clca3"]
```


##Cluster 3 is enterocytes
```{r}
marker_all_day_24_first[marker_all_day_24_first=="Apoa1"]
```

##Cluster 4 is Paneth cells
```{r}
marker_all_day_24_first[marker_all_day_24_first=="Defa24"]
```

##Cluster 5 is enteroendocrine cells
```{r}
marker_all_day_24_first[marker_all_day_24_first=="Chgb"]
```
##Cluster 6 is well define (4 cells)

```{r}
geni_7=marker_all_day_24_first[names(marker_all_day_24_first)==6]
logic_7=apply(norm_race[geni_7,cluster_race_entropy==6],1,function(x){
  if (length(x[x>1])==length(x)){
    return(TRUE)
  }
  else{return(FALSE)}
})

geni_7_top=geni_7[logic_7]
names(geni_7_top)=rep("6",length(geni_7_top))
```




```{r}
#Plot_balons(norm_race,cluster_race_entropy,marker_all_day_24_first[names(marker_all_day_24_first)==7],30,length(cluster_race_entropy),max_size=5,text_size=10)

IDENTOM::Plot_balons(norm_race,cluster_race_entropy,geni_7_top,length(geni_7_top),max_size=5,text_size=10)
```

##Cluster 7 is well define (3 cells)
```{r}
geni_7=marker_all_day_24_first[names(marker_all_day_24_first)==7]
logic_7=apply(norm_race[geni_7,cluster_race_entropy==7],1,function(x){
  if (length(x[x>1])==length(x)){
    return(TRUE)
  }
  else{return(FALSE)}
})

geni_7_top=geni_7[logic_7]
names(geni_7_top)=rep("7",length(geni_7_top))
```

```{r}
#Plot_balons(norm_race,cluster_race_entropy,marker_all_day_24_first[names(marker_all_day_24_first)==7],30,length(cluster_race_entropy),max_size=5,text_size=10)

IDENTOM::Plot_balons(norm_race,cluster_race_entropy,marker_all_day_24_first[names(marker_all_day_24_first)==7],30,max_size=5,text_size=10)

IDENTOM::Plot_balons(norm_race,cluster_race_entropy,geni_7_top,length(geni_7_top),max_size=5,text_size=10)
```

```{r}


white_black_7=white_black_marker(cluster_race_entropy,7,norm_race,marker_all_day_24_first,0)
sum(white_black_7)
white_black_6=white_black_marker(cluster_race_entropy,6,norm_race,marker_all_day_24_first,0)
sum(white_black_6)


white_black_5=white_black_marker(cluster_race_entropy,5,norm_race,marker_all_day_24_first,0)
sum(white_black_5)



white_black_4=white_black_marker(cluster_race_entropy,4,norm_race,marker_all_day_24_first,0)
sum(white_black_4)

white_black_3=white_black_marker(cluster_race_entropy,3,norm_race,marker_all_day_24_first,0)
sum(white_black_3)

white_black_2=white_black_marker(cluster_race_entropy,2,norm_race,marker_all_day_24_first,0)
sum(white_black_2)



#minimo 26 markers black and white
```

```{r}
entropy_3=marker_all_day_24_first[names(marker_all_day_24_first)==3]
entropy_5=marker_all_day_24_first[names(marker_all_day_24_first)==5]
entropy_4=marker_all_day_24_first[names(marker_all_day_24_first)==4]

```





```{r}
library(Biobase)
DefaultAssay(race_entropy) <- "RNA"
markers_first_gini_2=markers_first_method(race_entropy,cluster_race_id,names(race_entropy$RNA_snn_res.0.2),30)


```

```{r}
markers_first_elmir=markers_first_gini_2

marker_plot_day_24_first=markers_first_elmir[[1]]

marker_plot_plot_day_24_first=markers_first_elmir[[2]]
marker_all_day_24_first=markers_first_elmir[[3]]



```



```{r}


white_black_11=white_black_marker(cluster_race_id,11,norm_race,marker_all_day_24_first,0)
sum(white_black_11)

white_black_10=white_black_marker(cluster_race_id,10,norm_race,marker_all_day_24_first,0)
sum(white_black_10)

white_black_9=white_black_marker(cluster_race_id,9,norm_race,marker_all_day_24_first,0)
sum(white_black_9)


white_black_5=white_black_marker(cluster_race_id,5,norm_race,marker_all_day_24_first,0)
sum(white_black_5)

white_black_8=white_black_marker(cluster_race_id,8,norm_race,marker_all_day_24_first,0)
sum(white_black_8)

white_black_4=white_black_marker(cluster_race_id,4,norm_race,marker_all_day_24_first,0)
sum(white_black_4)

white_black_6=white_black_marker(cluster_race_id,6,norm_race,marker_all_day_24_first,0)
sum(white_black_6)

white_black_7=white_black_marker(cluster_race_id,7,norm_race,marker_all_day_24_first,0)
sum(white_black_7)




```
##Only cluster 4 and 5 from race id are well defined according to markers

```{r}

race_4=marker_all_day_24_first[names(marker_all_day_24_first)==4]
race_5=marker_all_day_24_first[names(marker_all_day_24_first)==5]

#race_7=marker_all_day_24_first[names(marker_all_day_24_first)==7]


```

```{r}
table(cluster_race_id,cluster_race_entropy)
```

```{r}
find_jacardi(cluster_race_entropy,cluster_race_id,norm_race,3,4)
find_jacardi(cluster_race_entropy,cluster_race_id,norm_race,5,5)
```

```{r}
find_jacardi_genes(entropy_5,race_5)
find_jacardi_genes(entropy_3,race_4)
```

```{r}
intersect(race_5,entropy_5)

intersect(race_4,entropy_3)

#intersect(race_7,entropy_4)
```


```{r }
library(UpSetR)
listInput <- list(entropy_mixing_cluster_5=race_5,race_id_cluster_5=entropy_5)

upset(fromList(listInput), order.by = "freq")

listInput <- list(entropy_mixing_cluster_4=race_4,race_id_cluster_3=entropy_3)

upset(fromList(listInput), order.by = "freq")






```

```{r }
Function_fisher=function(a,b,c){
matrice=matrix(c(sum(c[c%in%a]%in%c[c%in%b]),sum(c[c%in%a]%in%c[!c%in%b]),sum(c[!c%in%a]%in%c[c%in%b]),sum(c[!c%in%a]%in%c[!c%in%b])),nrow=2,ncol=2,byrow = T)
#Fisher test
fisher.test(matrice)
}
```

```{r}
Test_Fisher=function(entropy_cluster,markers_cluster,marker_top){
level=levels(as.factor(entropy_cluster))
p_value=rep(0,length(level))
odds_ratio=rep(0,length(level))
set_1=marker_top
for (i in 1:length(level)){
  set_2=markers_cluster[names(markers_cluster)==level[i]]
set_all=markers_cluster
show=Function_fisher(set_1,set_2,set_all)

p_value[i]=as.numeric(show$p.value)
names(p_value)[i]=level[i]
odds_ratio[i]=as.numeric(show$estimate)
names(odds_ratio)[i]=level[i]
}
return(list(p_value,odds_ratio))
}
test=Test_Fisher(cluster_race_id,marker_all_day_24_first,entropy_3)
test[[2]][test[[1]]<0.05&test[[2]]>2]
```

```{r }
test=Test_Fisher(cluster_race_id,marker_all_day_24_first,entropy_5)
test[[2]][test[[1]]<0.05&test[[2]]>2]
```


## End RaceID


## Big 10x_dataset 68 cells server hpc

## Pre processing

```{r }
#/home/ies/gabriele.lubatti/entropy_mixing/blood_10x_data/GiniClust2/Proj/10X_full/data






wd<-"/home/ies/gabriele.lubatti/entropy_mixing/blood_10x_data/GiniClust2/Proj/10X_full/data/single-cell-3prime-paper/pbmc68k_analysis/"
source(paste(wd,file.path('util.R'),sep=""))

pbmc_68k <- readRDS('/home/ies/gabriele.lubatti/entropy_mixing/blood_10x_data/GiniClust2/Proj/10X_full/data/pbmc68k_data.rds')
all_data <- pbmc_68k$all_data
pure_11 <- readRDS('/home/ies/gabriele.lubatti/entropy_mixing/blood_10x_data/GiniClust2/Proj/10X_full/data/all_pure_select_11types.rds')

purified_ref_11 <- load_purified_pbmc_types(pure_11,pbmc_68k$ens_genes)

m<-all_data[[1]]$hg19$mat
genes=all_data[[1]]$hg19$genes
barcode=all_data[[1]]$hg19$barcodes
genes_name=all_data[[1]]$hg19$gene_symbols


#l<-.normalize_by_umi(m)  
#ExprM.RawCounts.original<-t(l$m)
#m_n<-l$m
#df<-.get_variable_gene(m_n) 
#disp_cut_off<-sort(df$dispersion_norm,decreasing=T)[1000]
#df$used<-df$dispersion_norm >= disp_cut_off

#geneIndices<-l$use_genes
#genes<-pbmc_68k$ens_genes[geneIndices]

#load cluster labels
clusters<-read.table(paste0(wd,"68k_pbmc_barcodes_annotation.tsv"),sep="\t",header=T)

#clusters=read.table("/home/ies/gabriele.lubatti/entropy_mixing/blood_10x_data/GiniClust2/Proj/10X_ful#l/single-cell-3prime-paper-master/pbmc68k_analysis/PureClusters.csv")
m=t(m)
ExprM.RawCounts.original=m
#clusters<-read.table(paste(wd,"PureClusters.csv",sep=""),sep=",",header=F) #obtained from 10X code
#clusters<-clusters$V1
clusters_name=clusters$celltype
#colnames(ExprM.RawCounts.original)<-clusters
colnames(ExprM.RawCounts.original)<-barcode
rownames(ExprM.RawCounts.original)=genes_name

save(ExprM.RawCounts.original,file="ExprM.RawCounts.original.Rda")
save(clusters_name,file="clusters_name.Rda")
#save(ExprM.RawCounts, file=paste("results/",exprimentID, "_ExprM.RData",sep=""))
#save(ExprM.RawCounts.filter, file=paste("results/", exprimentID, "_ExprM.filter.RData", sep=""))
raw_blood_new_small=raw_blood_new[entropy_genes,]
```

#For seurat with hvgs as features
```{r }
library(EntropyMixingR)
Cluster_analysis_integrate_rare=function (raw_counts, batch_name_1, resolution, neighbors, max_dimension, 
    feature_genes = NULL) 
{
    batch_1 <- CreateSeuratObject(counts = raw_counts, project = batch_name_1)
    batch_1 <- subset(batch_1, subset = nFeature_RNA > 0)
    batch_1 <- NormalizeData(batch_1, verbose = FALSE)
    batch_1 <- FindVariableFeatures(batch_1, selection.method = "vst", 
        nfeatures = 3000)
    batch_combined = batch_1
    batch_combined <- ScaleData(batch_combined, verbose = FALSE)
    batch_combined <- RunPCA(batch_combined, npcs = max_dimension, 
        verbose = FALSE, features = feature_genes)
    batch_combined = RunUMAP(batch_combined, reduction = "pca", 
        dims = 1:20, set.seed = 42)
    batch_combined <- FindNeighbors(batch_combined, reduction = "pca", 
        dims = 1:max_dimension, k.param = neighbors)
    batch_combined <- FindClusters(batch_combined, resolution = resolution)
    return(batch_combined)
}
```


#For seurat with entropy of mixing (is not working we obtain more than 500 clusters)
```{r }
library(EntropyMixingR)
Cluster_analysis_integrate_rare=function (raw_counts, batch_name_1, resolution, neighbors, max_dimension, 
    feature_genes = NULL) 
{
    batch_1 <- CreateSeuratObject(counts = raw_counts, project = batch_name_1)
    batch_1 <- subset(batch_1, subset = nFeature_RNA > 0)
    batch_1 <- NormalizeData(batch_1, verbose = FALSE)
    #batch_1 <- FindVariableFeatures(batch_1, selection.method = "vst", 
       #nfeatures = 3000)
    batch_combined = batch_1
    batch_combined <- ScaleData(batch_combined, verbose = FALSE)
    #batch_combined@assays$RNA@var.features=common_genes
    batch_combined <- RunPCA(batch_combined, npcs = max_dimension, 
        verbose = FALSE, features = feature_genes)
   # batch_combined = RunUMAP(batch_combined, reduction = "pca", 
        #dims = 1:20, set.seed = 42)
    batch_combined <- FindNeighbors(batch_combined, reduction = "pca", 
        dims = 1:max_dimension, k.param = neighbors)
    batch_combined <- FindClusters(batch_combined, resolution = resolution)
    return(batch_combined)
}
```




```{r }
setwd("/home/ies/gabriele.lubatti/entropy_mixing/blood_10x_data/GiniClust2/Proj/10X_full/data")
blood_10x=Cluster_analysis_integrate_rare(ExprM.RawCounts.original,"Elmir_data",0.1,5,30)
save(blood_10x,file="blood_10x.Rda")

```

```{r }

setwd("/home/ies/gabriele.lubatti/entropy_mixing/blood_10x_data/GiniClust2/Proj/10X_full/data")
norm_blood=(GetAssayData(blood_10x, slot = "data",assay="RNA"))


general_cluster_blood=blood_10x$RNA_snn_res.0.1


knn_matrix_blood=(blood_10x@graphs$RNA_nn)

cordinate_umap=as.data.frame(Embeddings(blood_10x, reduction = "umap")[, 1:2])

save(norm_blood,file="norm_blood.Rda")
save(general_cluster_blood,file="general_cluster_blood.Rda")
save(knn_matrix_blood,file="knn_matrix_blood.Rda")

save(cordinate_umap,file="cordinate_umap.Rda")
#setwd("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/elmir_data")
#save(raw_elmir_5_30,file="raw_elmir_5_30.Rda")
#save(norm_elmir_5_30,file="norm_elmir_5_30.Rda")
#save(knn_matrix_elmir_5_30,file="knn_matrix_elmir_5_30.Rda")
#save(general_cluster_version_elmir,file="general_cluster_version_elmir.Rda")

as_matrix <- function(mat){

  tmp <- matrix(data=0L, nrow = mat@Dim[1], ncol = mat@Dim[2])
  
  row_pos <- mat@i+1
  col_pos <- findInterval(seq(mat@x)-1,mat@p[-1])+1
  val <- mat@x
    
  for (i in seq_along(val)){
      tmp[row_pos[i],col_pos[i]] <- val[i]
  }
    
  row.names(tmp) <- mat@Dimnames[[1]]
  colnames(tmp) <- mat@Dimnames[[2]]
  return(tmp)}

norm_blood_new=as_matrix(norm_blood)
knn_matrix_blood_new=as_matrix(knn_matrix_blood)

save(norm_blood_new,file="norm_blood_new.Rda")
save(knn_matrix_blood_new,file="knn_matrix_blood_new.Rda")
```



## Start with Entropy of Mixing

#Rscript --vanilla /home/ies/gabriele.lubatti/entropy_mixing/blood_10x_data/GiniClust2/Proj/10X_full/data/Entropy_hyper_server_blood_new.R -n norm_blood_new -k knn_matrix_blood_new -o blood_hyper_giusto.Rda -r 1 -l 3 -h 100 -t 1 -p 0.001
#Time difference of 4.383946 mins


#Rscript --vanilla /home/ies/gabriele.lubatti/entropy_mixing/blood_10x_data/GiniClust2/Proj/10X_full/d#ata/Entropy_hyper_server_blood_new.R -n norm_blood_new -k knn_matrix_blood_new -o blood_hyper_giusto.Rda -r 1 -l 3 -h 30 -t 1 -p 0.001

#Time difference of 1.056231 mins


##New
```{r }
##LAST VERSION (setting srun -p normal_q -c 10  --mem=500G -t 1:00:00 --pty bash)
 #cd /home/ies/gabriele.lubatti/entropy_mixing/blood_10x_data/GiniClust2/Proj/10X_full/data


#Fisica203!95

#Rscript --vanilla /home/ies/gabriele.lubatti/entropy_mixing/published_data/Entropy_server_december_2021_easy.R -n norm_blood_new -k knn_matrix_blood_new -o blood_december_2021.Rda -r 1 -l 3 -h 100 -t 1 -p 0.001 -c 10 -a TRUE

#Time difference of 3.81506 mins


#Rscript --vanilla /home/ies/gabriele.lubatti/entropy_mixing/published_data/Entropy_server_december_2021_easy.R -n norm_blood_new -k knn_matrix_blood_new -o blood_december_2021_all.Rda -r 1 -l 3 -h 100 -t 1 -p 0.001 -c 10 -a FALSE


```

##Old
```{r }
##LAST VERSION (setting srun -p normal_q -c 20  --mem=100G -t 1:00:00 --pty bash)

#srun -p normal_q -c 20  --mem=500G -t 1:00:00 --pty bash


#Rscript --vanilla /home/ies/gabriele.lubatti/entropy_mixing/published_data/Entropy_server_december_2021_new.R -n norm_blood_new -k knn_matrix_blood_new -o blood_december_2021.Rda -r 1 -l 3 -h 100 -t 1 -p 0.001 -c 10 -a TRUE


#Time difference of 4.784287 mins


```

#Below I used entropy of mixing with option h equal to 100
##Start here
```{r }
setwd("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/published_data")
load("norm_blood_new.Rda")
load("clusters_name.Rda")
load("raw_blood_new.Rda")
setwd("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/published_data")
load("tsne_10x.Rda")
load("clusters.Rda")
clusters=clusters$V1
load("cluster_blood_10x.Rda")
load("knn_matrix_blood_new.Rda")

setwd("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/published_data")
load("blood_hyper_giusto.Rda")
load("blood_hyper_giusto_2.Rda")
load("clusters_name.Rda")
load("blood_hyper_giusto.Rda")
load("cordinate_umap.Rda")
```

```{r }
entropy_genes=row.names(result)[result[,1]<1]
geni_top=row.names(result)[result[,1]==0&result[,2]<0.001]

geni_top_top=row.names(result)[order(unlist(result[,2]))]

geni_top_2=row.names(result)[order(as.numeric(result[,2]))]
```



```{r }
background=row.names(result)
#background=c(names(entropy_genes),names(entropy_genes_excluded))
```


```{r }
library(ggplot2)
p=list()
for(i in (geni_top)[1:5]){
  q=Plot_gene(norm_blood_new,tsne_10x,i,i)
  p=list(p,q)
}
p
```
```{r }
library(CIARA)
pdf("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/github_january_2022/figures_paper/sup_figure_6/sup_fig_6_c.pdf")
Plot_gene(norm_blood_new,tsne_10x,"PF4","PF4")
dev.off()
pdf("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/github_january_2022/figures_paper/sup_figure_6/sup_fig_6_d.pdf")
Plot_gene(norm_blood_new,tsne_10x,"GATA1","GATA1")
dev.off()
pdf("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/github_january_2022/figures_paper/sup_figure_6/sup_fig_6_e.pdf")
Plot_gene(norm_blood_new,tsne_10x,"HBG2","HBG2")
dev.off()
Plot_gene(norm_blood_new,tsne_10x,"TMEM40","TMEM40")




```




```{r }
#Plot_genes_sum(cordinate_umap,norm_blood_new,entropy_genes,"sum")
```

```{r }

Plot_genes_sum(tsne_10x,norm_blood_new,geni_top,"sum")
```

```{r }
#Plot_umap(cordinate_umap,clusters_name)
#clusters_name_new=rep("Other clusters",length(clusters_name))
#clusters_name_new[clusters_name=="Dendritic"]="Dendritic"
#Plot_umap(cordinate_umap,clusters_name_new)

#clusters_name_new=rep("Other clusters",length(clusters_name))
#clusters_name_new[clusters_name=="CD34+"]="CD34+"
#Plot_umap(cordinate_umap,clusters_name_new)
```
```{r }
#Plot_umap(cordinate_umap,cluster_blood_10x)
```

```{r }
Plot_umap(tsne_10x,clusters_name)
#Plot_umap(tsne_10x,final_cluster_version_sub)
pdf("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/github_january_2022/figures_paper/sup_figure_6/sup_fig_6_b.pdf")
Plot_umap(tsne_10x,clusters)
dev.off()
```

## END 68k dataset


## GiniClust2 dataset



## Pre processing

## Server

```{r }
Cluster_analysis_integrate_rare=function (raw_counts, batch_name_1, resolution, neighbors, max_dimension, 
    feature_genes = NULL) 
{
    batch_1 <- CreateSeuratObject(counts = raw_counts, project = batch_name_1)
    batch_1 <- subset(batch_1, subset = nFeature_RNA > 0)
    batch_1 <- NormalizeData(batch_1, verbose = FALSE)
    batch_1 <- FindVariableFeatures(batch_1, selection.method = "vst", 
        nfeatures = 2000)
    batch_combined = batch_1
    batch_combined <- ScaleData(batch_combined, verbose = FALSE)
    batch_combined <- RunPCA(batch_combined, npcs = max_dimension, 
        verbose = FALSE, features = feature_genes)
    batch_combined = RunUMAP(batch_combined, reduction = "pca", 
        dims = 1:20, set.seed = 42)
    batch_combined <- FindNeighbors(batch_combined, reduction = "pca", 
        dims = 1:max_dimension, k.param = neighbors)
    batch_combined <- FindClusters(batch_combined, resolution = resolution)
    return(batch_combined)
}
```

## Server
```{r }

ExprM.RawCounts <- read.delim("/home/ies/gabriele.lubatti/entropy_mixing/blood_10x_data/GiniClust2/Proj/inDrop_day4/data/GSM1599498_ES_d4_LIFminus.csv", sep=",", head=F)

#raw data

title=c("d4");
name=rep("bo",length(colnames(ExprM.RawCounts)))
for(i in 2:ncol(ExprM.RawCounts)){
  name[i]=c(paste(title,".Cell_",i-1,sep=""))
};
colnames(ExprM.RawCounts)=name
rownames(ExprM.RawCounts)=ExprM.RawCounts[,1]
ExprM.RawCounts= ExprM.RawCounts[,-1]

ExprM.RawCounts_new

giniclust2_data_small=read.csv('/home/ies/gabriele.lubatti/entropy_mixing/blood_10x_data/GiniClust2/Proj/inDrop_day4/results/d4_gene.expression.matrix.RawCounts.filtered.csv')

giniclust2=Cluster_analysis_integrate_rare(giniclust2_data_small,"Elmir_data",0.1,5,30)

norm_gini2=as.matrix(GetAssayData(giniclust2, slot = "data",assay="RNA"))


general_cluster_gini=giniclust2$RNA_snn_res.0.1


knn_matrix_gini=as.matrix(giniclust2@graphs$RNA_nn)

cordinate_umap_gini=as.data.frame(Embeddings(giniclust2, reduction = "umap")[, 1:2])

save(norm_gini2,file="norm_gini2.Rda")
save(general_cluster_gini,file="general_cluster_gini.Rda")
save(knn_matrix_gini,file="knn_matrix_gini.Rda")


load("/home/ies/gabriele.lubatti/entropy_mixing/blood_10x_data/GiniClust2/Proj/inDrop_day4/results/d4_FinalClustering.RData")

```


## Start with Entropy of Mixing

## IN THE FINAL VERSION I USED DEFAULT PARAMETERS FOR ENTROPY OF MIXING

#Rscript --vanilla /home/ies/gabriele.lubatti/entropy_mixing/blood_10x_data/GiniClust2/Proj/10X_full/data/Entropy_hyper_server_all_cells.R -n norm_gini2 -k knn_matrix_gini -o gini_2_old.Rda -r 1 -l 3 -h 30 -t 1 -p 0.001
#n_cells:278
#Time difference of 24.21628 secs


#Rscript --vanilla /home/ies/gabriele.lubatti/entropy_mixing/blood_10x_data/GiniClust2/Proj/10X_full/data/Entropy_hyper_server_all_cells.R -n norm_gini2 -k knn_matrix_gini -o gini_2_old_test.Rda -r 1 -l 3 -h 20 -t 1 -p 0.001
#n_cells:278
#Time difference of  8.870712 secs secs


#Rscript --vanilla /home/ies/gabriele.lubatti/entropy_mixing/blood_10x_data/GiniClust2/Proj/10X_full/data/Entropy_hyper_server_blood_new.R -n norm_gini2 -k knn_matrix_gini -o gini_2_new.Rda -r 1 -l 3 -h 30 -t 1 -p 0.001
#n_cells 278: 
#Time difference of 9.089941 secs

#Rscript --vanilla /home/ies/gabriele.lubatti/entropy_mixing/blood_10x_data/GiniClust2/Proj/10X_full/data/Entropy_hyper_server_blood_new.R -n norm_gini2 -k knn_matrix_gini -o gini_2_new_new.Rda -r 1 -l 3 -h 20 -t 1 -p 0.001
#n_cells 278: 
#Time difference of 2.819211 secs

##New
```{r }
##LAST VERSION (setting srun -p normal_q -c 10  --mem=100G -t 1:00:00 --pty bash)
 #cd /home/ies/gabriele.lubatti/entropy_mixing/blood_10x_data/GiniClust2/Proj/10X_full/data


#Rscript --vanilla /home/ies/gabriele.lubatti/entropy_mixing/published_data/Entropy_server_december_2021_easy.R -n norm_gini2 -k knn_matrix_gini -o gini2_december_2021_new.Rda -r 1 -l 3 -h 30 -t 1 -p 0.001 -c 10 -a FALSE
#Time difference of 24.37659 secs 

#Rscript --vanilla /home/ies/gabriele.lubatti/entropy_mixing/published_data/Entropy_server_december_2021_easy.R -n norm_gini2 -k knn_matrix_gini -o gini2_december_2021_new_a.Rda -r 1 -l 3 -h 30 -t 1 -p 0.001 -c 10 -a TRUE
#Time difference of 5.914052 secs


#Execution Time CIARA Approx: 2.1s
#Execution Time CIARA Full: 8.09s

```

##Old
```{r }
##LAST VERSION (setting srun -p normal_q -c 20  --mem=500G -t 1:00:00 --pty bash)




#Rscript --vanilla /home/ies/gabriele.lubatti/entropy_mixing/published_data/Entropy_server_december_2021_new.R -n norm_gini2 -k knn_matrix_gini -o gini2_december_2021_new.Rda -r 1 -l 3 -h 30 -t 1 -p 0.001 -c 10 -a FALSE


#Time difference of 28.94114 secs 


#Rscript --vanilla /home/ies/gabriele.lubatti/entropy_mixing/published_data/Entropy_server_december_2021_new.R -n norm_gini2 -k knn_matrix_gini -o gini2_december_2021_a.Rda -r 1 -l 3 -h 30 -t 1 -p 0.001 -c 10 -a TRUE


#Time difference of 6.929575 secs
```


```{r }
setwd("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/published_data")
load("gini_2_old_test.Rda")
vedo_1=result
load("finalCluster.Rda")
load("giniclust2.Rda")

norm_gini2=(GetAssayData(giniclust2, slot = "data",assay="RNA"))
raw_gini2=(GetAssayData(giniclust2, slot = "data",assay="RNA"))
setwd("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/published_data")
save(norm_gini2,file="norm_gini2.Rda")

general_cluster_gini=giniclust2$RNA_snn_res.0.1


knn_matrix_gini=(giniclust2@graphs$RNA_nn)

setwd("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/published_data")
knn_matrix_gini2=knn_matrix_gini
save(knn_matrix_gini2,file="knn_matrix_gini2.Rda")

cordinate_umap_gini=as.data.frame(Embeddings(giniclust2, reduction = "umap")[, 1:2])

```

```{r }
setwd("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/published_data")
load("gini_2_old.Rda")
result_old=result
```






```{r }
entropy_genes_old=row.names(result_old)[result_old[,2]<1]
#geni_top_old=row.names(result_old)[result_old[,1]<0.00001&result_old[,2]==0]
```




```{r }
setwd("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/published_data")
load("gini_2_new.Rda")
result_new=result
```


```{r }
entropy_genes_new=row.names(result_new)[result_new[,1]<1]
#geni_top_new=row.names(result_new)[result_new[,2]<0.00001&result_new[,1]==0]
```



```{r }
jacardi_entropy=find_jacardi_genes(entropy_genes_old,entropy_genes_new)
#jacardi_entropy=find_jacardi_genes(background_old,background_new)

```


```{r }
#setwd("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/published_data")
#load("gini_2.Rda")
#library(EntropyMixingR)
setwd("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/published_data")
load("gini_2_old_old.Rda")
entropy_genes=row.names(result)[result[,2]<1]
geni_top=row.names(result)[result[,1]<0.00001&result[,2]==0]
background=row.names(result)
#geni_top=row.names(result)[result[,1]==0&result[,2]<0.001]
#background=c(names(entropy_genes),names(entropy_genes_excluded))
```



```{r }
library(ggplot2)
p=list()
for(i in (geni_top)[1:10]){
  q=Plot_gene(norm_gini2,cordinate_umap_gini,i,i)
  p=list(p,q)
}
p
```
```{r }
Plot_genes_sum(cordinate_umap_gini,norm_gini2,entropy_genes,"sum")
```


```{r }
library(ggplot2)
Plot_genes_sum(cordinate_umap_gini,norm_gini2,geni_top,"sum")
```



```{r }
#mayra_seurat_entropy_24=Cluster_analysis_integrate_rare(mayra_dati_raw_24,"24h",0.01,5,30,names(entro#py_genes))
library(IDENTOM)
giniclu2_entropy=Cluster_analysis_integrate_rare(raw_gini2,"24h",0.3,5,30,entropy_genes)

#giniclu2_entropy_2=EntropyMixingR::Cluster_analysis_integrate_rare(raw_gini2,"24h",0.3,3,30,entropy_genes)


library(clustree)
setwd("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library")
#library(EntropyMixingR)
find_resolution(giniclu2_entropy,seq(0.01,1,0.1))

#find_resolution(giniclust2,seq(0.01,1,0.1))

#giniclust2=FindClusters(giniclust2,resolution = 0.2)
#mayra_seurat_entropy_24=FindClusters(mayra_seurat_entropy_24,resolution = 0.1)

```


```{r }
Plot_umap(cordinate_umap_gini,as.vector(giniclu2_entropy$RNA_snn_res.0.3))

cluster_entropy=as.vector(giniclu2_entropy$RNA_snn_res.0.3)

finalCluster_plot=cluster_entropy
finalCluster_plot[cluster_entropy==1]="Rare cell type 2"
finalCluster_plot[cluster_entropy==2]="Rare cell type 1"
finalCluster_plot[cluster_entropy!=2&cluster_entropy!=1]="No rare cell type"
Plot_umap(cordinate_umap_gini,finalCluster_plot)



Plot_umap(cordinate_umap_gini,finalCluster)

finalCluster_plot=finalCluster
finalCluster_plot[finalCluster==1]="Rare cell type 1"
finalCluster_plot[finalCluster==2]="Rare cell type 2"
finalCluster_plot[finalCluster!=2&finalCluster!=1]="No rare cell type"
Plot_umap(cordinate_umap_gini,finalCluster_plot)
```

```{r}
merge_cluster=function (old_cluster, new_cluster, max_number = NULL) 
{
  old_cluster=as.vector(giniclust2$RNA_snn_res.0.2)
  new_cluster=as.vector(giniclu2_entropy$RNA_snn_res.0.3)
  names(old_cluster)=colnames(norm_gini2)
  names(new_cluster)=colnames(norm_gini2)
  max_number=15
    if (is.null(max_number)) {
        cluster_final = old_cluster
        cluster_final[match(names(new_cluster), names(old_cluster))] = new_cluster
        return(cluster_final)
    }
    else {
        index_small = which(table(new_cluster) < max_number)
        cluster_small = names(table(new_cluster))[index_small]
        print(cluster_small)
        cluster_final = old_cluster
        for (i in 1:length(cluster_small)) {
            cluster_final[names(old_cluster) %in% names(new_cluster)[as.vector(new_cluster) == 
                cluster_small[i]]] = paste(cluster_small[i], 
                "step_2", sep = "-")
        }
        return(cluster_final)
    }
}
```


## Gene used in the paper
```{r }
norm_gini_entropy_3=as.matrix(GetAssayData(giniclu2_entropy, slot = "data",assay="RNA"))

geni_marker_1=c("Col4a1","Col4a2","Lama1","Ctsl")

geni_marker_2=c("Rhox6","Rhox9","Sct")
positive_cells_1=find_positive_cells(norm_gini_entropy_3,geni_marker_1,threshold = 0)
positive_cells_2=find_positive_cells(norm_gini_entropy_3,geni_marker_2,threshold = 0)

```

```{r }

names(giniclu2_entropy$RNA_snn_res.0.3)[giniclu2_entropy$RNA_snn_res.0.3==10]





find_jacardi_new(as.vector(giniclu2_entropy$RNA_snn_res.0.3),norm_gini_entropy_3,2,positive_cells_1)

find_jacardi_new(as.vector(giniclu2_entropy$RNA_snn_res.0.3),norm_gini_entropy_3,1,positive_cells_2)


find_jacardi_new(finalCluster,norm_gini_entropy_3,2,positive_cells_2)

find_jacardi_new(finalCluster,norm_gini_entropy_3,1,positive_cells_1)


```

```{r}
#old_cluster=as.vector(giniclust2$RNA_snn_res.0.2)
  #new_cluster=as.vector(giniclu2_entropy$RNA_snn_res.0.3)
  #names(old_cluster)=colnames(norm_gini2)
  #names(new_cluster)=colnames(norm_gini2)
#final_cluster=merge_cluster(old_cluster,new_cluster,15)
```

```{r }
#Plot_umap(cordinate_umap_gini,final_cluster)
```











```{r }


Plot_gene(norm_gini2,cordinate_umap_gini,"Col4a1","Col4a1")
Plot_gene(norm_gini2,cordinate_umap_gini,"Col4a2","Col4a2")
Plot_gene(norm_gini2,cordinate_umap_gini,"Lama1","Lama1")
#Plot_gene(norm_gini2,cordinate_umap_gini,"Lama2","Lama2")
#Plot_gene(norm_gini2,cordinate_umap_gini,"Cstl","Cstl")

Plot_gene(norm_gini2,cordinate_umap_gini,"Rhox9","Rhox9")

Plot_gene(norm_gini2,cordinate_umap_gini,"Rhox6","Rhox6")

Plot_gene(norm_gini2,cordinate_umap_gini,"Sct","Sct")


Plot_gene(norm_gini2,cordinate_umap_gini,"Sct","Sct")



```

## End GiniClust2


##Start SAM dataset
##Start with new dataset (link here for tutorial https://github.com/atarashansky/self-assembling-manifold/blob/hnswlib/tutorial/SAM_Batch/schisto_batch_correction.ipynb)
## link for paper https://elifesciences.org/articles/48994#s4

```{r } 
new_data=read.csv("/Users/gabriele.lubatti/Downloads/self-assembling-manifold-master/example_data/schisto2.5_tpm.csv")
row.names(new_data)=new_data[,1]
new_data=new_data[,-1]
new_data=log(new_data+1)
```

```{r }
Cluster_analysis_integrate_rare_here=function (raw_counts, batch_name_1, resolution, neighbors, max_dimension, 
    feature_genes = NULL) 
{
    batch_1 <- CreateSeuratObject(counts = raw_counts, project = batch_name_1)
    batch_1 <- subset(batch_1, subset = nFeature_RNA > 0)
    #batch_1 <- NormalizeData(batch_1, verbose = FALSE)
    batch_1 <- FindVariableFeatures(batch_1, selection.method = "vst", 
        nfeatures = 2000)
    batch_combined = batch_1
    batch_combined <- ScaleData(batch_combined, verbose = FALSE)
    batch_combined <- RunPCA(batch_combined, npcs = max_dimension, 
        verbose = FALSE, features = feature_genes)
    batch_combined = RunUMAP(batch_combined, reduction = "pca", 
        dims = 1:20, set.seed = 42)
    batch_combined <- FindNeighbors(batch_combined, reduction = "pca", 
        dims = 1:max_dimension, k.param = neighbors)
    batch_combined <- FindClusters(batch_combined, resolution = resolution)
    return(batch_combined)
}
```

```{r }
snn_seurat=Cluster_analysis_integrate_rare_here(new_data,"Elmir data",0.01,5,30)

setwd("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/github_january_2022")
snn_knn=as.matrix(snn_seurat@graphs$RNA_nn)

snn_norm=as.matrix(GetAssayData(snn_seurat, slot = "data",assay="RNA"))

cordinate_umap=as.data.frame(Embeddings(snn_seurat, reduction = "umap")[, 1:2])

save(snn_knn,file="snn_knn.Rda")
save(snn_norm,file="snn_norm.Rda")

```


##LAST VERSION (setting srun -p normal_q -c 10  --mem=10G -t 1:00:00 --pty bash)
#Rscript --vanilla /home/ies/gabriele.lubatti/entropy_mixing/elmir_data/Entropy_server_december_2021_easy.R -n snn_norm -k snn_knn -o snn_december_2021.Rda -r 1 -l 3 -h 170 -t 1 -p 0.001 -c 10 -a FALSE




```{r }
load("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/github_january_2022/snn_december_2021.Rda")


entropy_sort=row.names(result)[order(as.numeric(result[,1]))]
entropy_genes=row.names(result)[result[,1]<1]
p=list()
for(i in (entropy_sort)[1:10]){
  q=Plot_gene(snn_norm,cordinate_umap,i,i)
  p=list(p,q)
}
p



genes_paper=c('Smp-051920','Smp-041540','Smp-142120', 'Smp-087310', 'Smp-005350')

#apply(snn_norm[genes_paper,],1,function(x){length(x[x>1])})
length(entropy_genes)
which(entropy_sort%in%genes_paper)


```

```{r }
genes_paper_small=genes_paper[(genes_paper%in%entropy_genes)]

p=list()
for(i in genes_paper_small){
  q=Plot_gene(snn_norm,cordinate_umap,i,i)
  p=list(p,q)
}
p



```


##Stop SAM dataset
## Start comaprison with runnin time (to update)


```{r}
n_cells=c(278,2484,2545,1580,3984,317,68579,472,1195,766)
n_time_old=c(24,434.8289,648.4758,316.3698,1085.404,45.14149,3600,22.9829,146.3029,90.08904)
n_time_new=c(9,111.8667,107.7418,129.3736,326.7854,18.67952,294.2758,22.38227,63.48432,18.54127)
n_time_new[5]=227.8808
n_time_old[5]=737.0982
n_time_old=n_time_old/60
n_time_new=n_time_new/60
origin=c("GiniClust2_fig4a","GiniClust-fig3b","GiniClust-fig6b","FiRe-fig3b","CellSIUS-fig4d","RaceID-vignette","Zheng-PBMCs","Human gastrula","mouse ESCs")


n_cells_all=rep(n_cells,2)
n_time_all=c(n_time_old,n_time_new)
label_all=c(rep("All local regions",length(n_cells)),rep("Approximation",length(n_cells)))
all_data=data.frame(n_cells_all,n_time_all,label_all)

```

```{r}
ggplot(all_data, aes(x=n_cells_all, y=n_time_all,color=label_all)) +
  geom_point(size=2)+xlab("Number of cells") + ylab("Time (minutes)") +ggtitle("Entropy of Mixing: running time")+labs(col="setting")+geom_line(aes(colour = label_all), size = 1)

```
12.28497
```{r}
n_cells=c(278,2484,2545,1580,3984,317,472,1195,766)
n_time_old=c(24,434.8289,648.4758,316.3698,1085.404,45.14149,22.982,146.3029,90.08904)
n_time_new=c(9,111.8667,107.7418,129.3736,326.7854,18.67952,22.38227,63.48432,18.54127)
n_time_new[5]=227.8808
n_time_old[5]=737.0982

n_time_old=n_time_old/60
n_time_new=n_time_new/60
origin=c("GiniClust2_fig4a","GiniClust-fig3b","GiniClust-fig6b","FiRe-fig3b","CellSIUS-fig4d","RaceID-vignette","GiniClust-fig5b","Human gastrula","mouse ESCs")


n_cells_all=rep(n_cells,2)
n_time_all=c(n_time_old,n_time_new)
label_all=c(rep("All local regions",length(n_cells)),rep("Approximation",length(n_cells)))

#label_all=rep(origin,2)
all_data=data.frame(n_cells_all,n_time_all,label_all)

```

```{r}
ggplot(all_data, aes(x=n_cells_all, y=n_time_all,color=label_all)) +
  geom_point(size=2)+xlab("Number of cells") + ylab("Time (minutes)") +ggtitle("Entropy of Mixing: running time")+labs(col="setting")+geom_line(aes(colour = label_all), size = 1)


```


## Stop running time comparison

##New time running plot
```{r}
n_cells=c(2484,472,1195,665,766,1580,3984,317,68579,278)
n_time_old=c(12.84244,0.4654357,6.323627,2.44859,2.995655,5.809975,21.26795,1.190571,0.4823523)
n_time_new=c(1.565962,0.2397915,1.647605,0.2696713,0.4873552,1.389862,3.951076,0.3308998,4.784287,0.1154929)
n_time_old=n_time_old
n_time_new=n_time_new
origin=c("GiniClust-fig3b","GiniClust-fig6b","Human gastrula","mouse after gastrula","mouse ESCs","FiRe-fig3b","CellSIUS-fig4d","RaceID-vignette","Zheng-PBMCs","GiniClust2_fig4a")



n_cells_all=rep(n_cells,2)
n_cells_all=c(n_cells[origin!="Zheng-PBMCs"],n_cells)
n_time_all=c(n_time_old,n_time_new)
label_all=c(rep("All local regions",length(n_cells[origin!="Zheng-PBMCs"])),rep("Approximation",length(n_cells)))
all_data=data.frame(n_cells_all,n_time_all,label_all)

```




```{r}
ggplot(all_data, aes(x=n_cells_all, y=n_time_all,color=label_all)) +
  geom_point(size=2)+xlab("Number of cells") + ylab("Time (minutes)") +ggtitle("Entropy of Mixing: running time")+labs(col="setting")+geom_line(aes(colour = label_all), size = 1)


```

```{r}
n_cells=c(665,1285,766,1195,2484,472,1580,3984,317,278,68579)
n_time_old=c(1.17694,4.450016,1.506383,3.346655,9.34686,0.4444718,5.016574,17.8528,0.6385703,0.4062765)
n_time_new=c(0.1391201,0.8378267,0.2914613,0.9303223,1.323001,0.233798,1.20237,2.947906,0.1719375,0.09856753,3.81506)



n_time_old=n_time_old
n_time_new=n_time_new


origin=c("mouse endoderm","mouse ESCs","mouse ESCs(24hRA)","human gastrula","GiniClust-fig3b","GiniClust-fig6b","FiRe-fig3b","CellSIUS-fig4d","RaceID-vignette","GiniClust2_fig4a","Zheng-PBMCs")



n_cells_all=rep(n_cells,2)
n_cells_all=c(n_cells[origin!="Zheng-PBMCs"],n_cells)
n_time_all=c(n_time_old,n_time_new)
label_all=c(rep("All local regions",length(n_cells[origin!="Zheng-PBMCs"])),rep("Approximation",length(n_cells)))
all_data=data.frame(n_cells_all,n_time_all,label_all)

```

```{r}
pdf("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/github_january_2022/figures_paper/sup_figure_6/sup_fig_6_a.pdf")
ggplot(all_data, aes(x=n_cells_all, y=n_time_all,color=label_all)) +
  geom_point(size=2)+xlab("Number of cells") + ylab("Time (minutes)") +ggtitle("CIARA: running time - R")+labs(col="setting")+geom_line(aes(colour = label_all), size = 1)+ylim(c(0,20))+scale_y_continuous(breaks=seq(0,20,2),limits = c(0,20))
dev.off()

```





```{r}
##Giniclust fig3b
Execution Time CIARA Approx: 42.54s
Execution Time CIARA Full: 203.07s
##Giniclust fig5b
Execution Time CIARA Approx: 11.96s
Execution Time CIARA Full: 22.26s
##Giniclust2
Execution Time CIARA Approx: 2.1s
Execution Time CIARA Full: 8.09s
##FiRE
Execution Time CIARA Approx: 51.29s
Execution Time CIARA Full: 176.15s
##Cellsius
Execution Time CIARA Approx: 119.78s
Execution Time CIARA Full: 462.7s
##RaceiD
Execution Time CIARA Approx: 4.31s
Execution Time CIARA Full: 15.93s
#Ane24h:
Execution Time CIARA Approx: 16.01s
Execution Time CIARA Full: 82.14s
#Ane0h:
#Execution Time CIARA Approx: 45.63s
#Execution Time CIARA Full: 194.34s

Execution Time CIARA Approx: 33.17s
Execution Time CIARA Full: 137.17s
##Elmir:
#Execution Time CIARA Approx: 45.1s
#Execution Time CIARA Full: 153.31s

Execution Time CIARA Approx: 30.45s
Execution Time CIARA Full: 104.66s
##Endoderm
Execution Time CIARA Approx: 6.17s
Execution Time CIARA Full: 37.38s

```


```{r}
n_cells=c(665,1285,766,1195,2484,472,1580,3984,317,278)

n_time_new=c(6.17,33.17,16.01,30.45,42.54,11.96,51.29,119.78,4.31,2.1)

n_time_old=c(37.38,137.17,82.14,104.66,203.07,22.26,176.15,462.7,15.93,8.09)


n_time_old=n_time_old/60
n_time_new=n_time_new/60


#origin=c("mouse endoderm","mouse ESCs","mouse ESCs(24hRA)","human gastrula","GiniClust-fig3b","GiniClust-fig6b","FiRe-fig3b","CellSIUS-fig4d","RaceID-vignette","GiniClust2_fig4a","Zheng-PBMCs")
origin=c("mouse endoderm","mouse ESCs","mouse ESCs(24hRA)","human gastrula","GiniClust-fig3b","GiniClust-fig6b","FiRe-fig3b","CellSIUS-fig4d","RaceID-vignette","GiniClust2_fig4a")



n_cells_all=rep(n_cells,2)
#n_cells_all=c(n_cells[origin!="Zheng-PBMCs"],n_cells)
n_time_all=c(n_time_old,n_time_new)
label_all=c(rep("All local regions",length(n_cells)),rep("Approximation",length(n_cells)))
all_data=data.frame(n_cells_all,n_time_all,label_all)



```

```{r}
#pdf("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/github_january_2022/figures_paper/sup_figure_6/sup_fig_6_a_python.pdf")
ggplot(all_data, aes(x=n_cells_all, y=n_time_all,color=label_all)) +
  geom_point(size=2)+xlab("Number of cells") + ylab("Time (minutes)") +ggtitle("CIARA: running time - python")+labs(col="setting")+geom_line(aes(colour = label_all), size = 1)+ylim(c(0,20))+scale_y_continuous(breaks=seq(0,20,2),limits = c(0,20))
#dev.off()

```
```{r}
n_cells=c(665,1285,766,1195,2484,472,1580,3984,317,278)
n_time_old=c(1.17694,4.450016,1.506383,3.346655,9.34686,0.4444718,5.016574,17.8528,0.6385703,0.4062765)
n_time_new=c(0.1391201,0.8378267,0.2914613,0.9303223,1.323001,0.233798,1.20237,2.947906,0.1719375,0.09856753)



n_time_old=n_time_old
n_time_new=n_time_new


origin=c("mouse endoderm","mouse ESCs","mouse ESCs(24hRA)","human gastrula","GiniClust-fig3b","GiniClust-fig6b","FiRe-fig3b","CellSIUS-fig4d","RaceID-vignette","GiniClust2_fig4a")



n_cells_all=rep(n_cells,2)
n_cells_all=c(n_cells,n_cells)
n_time_all=c(n_time_old,n_time_new)
label_all=c(rep("All local regions",length(n_cells)),rep("Approximation",length(n_cells)))
all_data=data.frame(n_cells_all,n_time_all,label_all)

```

```{r}
pdf("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/github_january_2022/figures_paper/sup_figure_6/sup_fig_6_a.pdf")
ggplot(all_data, aes(x=n_cells_all, y=n_time_all,color=label_all)) +
  geom_point(size=2)+xlab("Number of cells") + ylab("Time (minutes)") +ggtitle("CIARA: running time - R")+labs(col="setting")+geom_line(aes(colour = label_all), size = 1)+ylim(c(0,20))+scale_y_continuous(breaks=seq(0,20,2),limits = c(0,20))
dev.off()

```
##Command for Marco for generating csv file for published dataset

```{r}
setwd("/home/ies/gabriele.lubatti/ciara_marco")

vector_rpkm=list.files("/home/ies/gabriele.lubatti/ciara_marco")
setwd("/home/ies/gabriele.lubatti/ciara_marco/endoderm")
vector_rpkm=vector_rpkm[17:21]
name_file=rep(0,length(vector_rpkm))
for (i in 1:length(vector_rpkm)){
name_file[i]=load(vector_rpkm[i])
setwd("/home/ies/gabriele.lubatti/ciara_marco")
write.csv(name_file[i],file=paste0(name_file[i],".csv"))
}



/home/ies/gabriele.lubatti/.linuxbrew/Cellar/python@3.7/3.7.9_2/bin/marco/bin/python3 wrapper.py --filepath_norm "/home/ies/gabriele.lubatti/ciara_marco/norm_cellsius.csv" --filepath_knn "/home/ies/gabriele.lubatti/ciara_marco/knn_matrix_cellsius.csv"


/home/ies/gabriele.lubatti/.linuxbrew/Cellar/python@3.7/3.7.9_2/bin/marco/bin/python3 wrapper.py --filepath_norm "/home/ies/gabriele.lubatti/ciara_marco/giniclust_fig_3b/norm_gini.csv" --filepath_knn "/home/ies/gabriele.lubatti/ciara_marco/giniclust_fig_3b/knn_matrix_gini_5_30.csv"

```


```{r}
/home/ies/gabriele.lubatti/.linuxbrew/Cellar/python@3.7/3.7.9_2/bin/marco/bin/python3 wrapper.py --filepath_norm "/home/ies/gabriele.lubatti/ciara_marco/68k_cells/norm_blood_new.csv" --filepath_knn "/home/ies/gabriele.lubatti/ciara_marco/knn_matrix_blood_new.csv"
```

```{r}
/home/ies/gabriele.lubatti/.linuxbrew/Cellar/python@3.7/3.7.9_2/bin/marco/bin/python3 wrapper.py --filepath_norm "/home/ies/gabriele.lubatti/ciara_marco/elmir/norm_elmir_5_30.csv" --filepath_knn "/home/ies/gabriele.lubatti/ciara_marco/elmir/knn_matrix_elmir_5_30.csv"
```



#Time Time difference of 3.346655 mins 
#55.81934 secs

#Time difference of 2.830453 mins

##LAST VERSION (setting srun -p normal_q -c 10  --mem=10G -t 1:00:00 --pty bash)
#Rscript --vanilla /home/ies/gabriele.lubatti/entropy_server_december_2021_easy.R -n norm_elmir_5_30 -k knn_matrix_elmir_5_30 -o elmir_december_2021_update.Rda -r 1 -l 3 -h 20 -t 1 -p 0.001 -c 10 -a FALSE
#Time difference of 6.323627 mins 1.647605 mins

#Rscript --vanilla /home/ies/gabriele.lubatti/Entropy_server_december_2021_easy.R -n norm_elmir_5_30 -k knn_matrix_elmir_5_30 -o elmir_december_2021_update.Rda -r 1 -l 3 -h 20 -t 1 -p 0.001 -c 10 -a TRUE
#Time difference of 1.647605 mins

```{r}
load("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/elmir_data/knn_matrix_elmir_5_30.Rda")

load("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/elmir_data/norm_elmir_5_30.Rda")

load("/Users/gabriele.lubatti/Downloads/elmir_december_2021_update.Rda")
result_update=result
load(system.file("extdata", "elmir_december_2021.Rda", package ="CIARA"))
load("/Users/gabriele.lubatti/Downloads/elmir_december_2021_update_a.Rda")
result_a=result

find_jacardi_genes(row.names(result_a)[result_a[,1]<1],row.names(result_update)[result_update[,1]<1])

```

