---
title: "Mouse_after_gastrula"
author: "Gabriele Lubatti"
date: "10/21/2021"
output:
  html_document: default
  pdf_document: default
---


```{r } 
gc()
rm(list=ls())
#library(devtools)
#install_github("ScialdoneLab/CIARA",auth_token="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",ref="master")
```

```{r } 
#library(Seurat)
#library(clustree)
#library(plotly)
#library(ggplot2)
#library(CIARA)
```


```{r } 
library(CIARA)
library(UpSetR)
library(Seurat)
```





```{r }
detect_localized_genes=function(knn_matrix_ane_24_update,norm_counts_ane_24_update,geni_top_top,top_number){
list_intersect=rep(list(0),length(colnames(knn_matrix_ane_24_update)))
rank_intersect=rep(0,length(colnames(knn_matrix_ane_24_update)))
for ( i in 1:length(colnames(knn_matrix_ane_24_update))){
  knn_name=names(which(knn_matrix_ane_24_update[i,]==1))
  geni_select=geni_top_top[1:top_number]
  sum_knn=apply(norm_counts_ane_24_update[geni_select,knn_name],2,function(x){
    somma=geni_select[which(x>1)]
    return(somma)
  })
    #rank_intersect[i]=sum(duplicated(as.vector(unlist(sum_knn))))
    list_intersect[[i]]=unique(as.vector(unlist(sum_knn))[duplicated(as.vector(unlist(sum_knn)))])
    rank_intersect[i]=length(list_intersect[[i]])

}
return(list(list_intersect,rank_intersect))
}
```

##old implementation
```{r }
detect_localized_genes_old=function(knn_matrix_ane_24_update,norm_counts_ane_24_update,geni_top_top,top_number){
list_intersect=rep(list(0),length(colnames(knn_matrix_ane_24_update)))
rank_intersect=rep(0,length(colnames(knn_matrix_ane_24_update)))
for ( i in 1:length(colnames(knn_matrix_ane_24_update))){
  knn_name=names(which(knn_matrix_ane_24_update[i,]==1))
  geni_select=geni_top_top[1:top_number]
  sum_knn=apply(norm_counts_ane_24_update[geni_select,knn_name],2,function(x){
    somma=geni_select[which(x>1)]
    return(somma)
  })
    rank_intersect[i]=sum(duplicated(as.vector(unlist(sum_knn))))
    list_intersect[[i]]=unique(as.vector(unlist(sum_knn))[duplicated(as.vector(unlist(sum_knn)))])

}
return(list(list_intersect,rank_intersect))
}
```

```{r }

selection_localized_genes_here=function (list_intersect, geni_top_top, 
    max_number_genes = 5){ 
text=rep(0,length(list_intersect))

for (i in 1:length(list_intersect)){
  if (length(list_intersect[[i]])==0){
    localized_genes="No specific genes"
    text[i] <- "No specific genes"
   
  }
  else {
  localized_genes=geni_top_top[which(geni_top_top %in% list_intersect[[i]])]
  
  max_genes=min(max_number_genes,length(localized_genes))
  if (max_genes<max_number_genes){
    warning("max_number bigger than length of localized_genes. Set max_number_genes = length(localized_genes) ")
  }
  localized_genes=localized_genes[1:max_genes]
  text[i]=paste(localized_genes, collapse = "-")
  
  
  }
}
return(text)
}


```



```{r }
plot_interactive_here=function (coordinate_umap, norm_counts,color, text,ramp_list, min_x = NULL, max_x = NULL, 
    min_y = NULL, max_y = NULL)
  {
  
  
  index_sort <- order(color)
  text=text[index_sort]
    row.names(coordinate_umap) <- colnames(norm_counts)
    coordinate_umap <- coordinate_umap[colnames(norm_counts)[index_sort], 
        ]
    color <- sort(color)
    

    if (!requireNamespace("plotly", quietly = TRUE)) {
        stop("Package plotly needed for interactive == TRUE. Please install it: install.packages('plotly')")
    }
    colnames(coordinate_umap) <- c("UMAP_1", "UMAP_2")
    fig <- plotly::plot_ly(data = coordinate_umap, x = ~UMAP_1, 
        y = ~UMAP_2, color = ~color, type = "scatter", mode = "markers", 
        marker = list(size = 5, width = 2,line = list(color = "black",
                                         width = 0.5)), text = ~text, hoverinfo = "text",colors=ramp_list)
   #fig <- fig %>% 
#layout(plot_bgcolor='white')
    
    if (!is.null(min_x)) {
        fig <- fig %>% plotly::layout(xaxis = list(range = c(min_x, 
            max_x)), yaxis = list(range = c(min_y, max_y))) %>% 
layout(plot_bgcolor='rgb(254, 247, 234)') %>% 
layout(paper_bgcolor='rgb(254, 247, 234)')
    }
    return(fig)
    }
    

#line = list(color = 'rgba(152, 0, 0, .8)',
                                         #width = 2)
#'#e5ecf6'
```

```{r }
plot_genes_sum_here=function (coordinate_umap, norm_counts, color_cell, name_title,hm.colors) 
{
    index_sort <- order(color_cell)
    row.names(coordinate_umap) <- colnames(norm_counts)
    coordinate_umap <- coordinate_umap[colnames(norm_counts)[index_sort], 
        ]
    color_cell <- sort(color_cell)
    
    umap_plot <- ggplot2::ggplot(coordinate_umap, ggplot2::aes(coordinate_umap[, 
        1], coordinate_umap[, 2])) + ggplot2::geom_point(ggplot2::aes(colour = color_cell),size = 1) +
        ggplot2::xlab("UMAP 1") + ggplot2::ylab("UMAP 2") + ggplot2::labs(col = "Number of localized genes") + 
        #ggplot2::scale_colour_gradient(low = "white",high = "blue4") + 
      scale_colour_gradientn(colours = hm.colors)+
        ggplot2::ggtitle(name_title)
    return((umap_plot))
}

#"#FFFFFF",ramp.list[index_color])

#gradient2 = colorpanel( sum( breaks[-1] > as.numeric(quantile(breaks,0.25)) ), #ramp.list[index_color], "#00008B" )
#hm.colors
#geom_point(size = 1.5, alpha = 0.5, na.rm = T, shape = 21, colour = "black",stroke=0.1)

```

```{r }
geni_ortogoli_unique=function(rabbit_id,name_mouse){

#setwd("/Users/gabriele.lubatti/Desktop/Phd/Early_embryonic_like_cells/Rabbit/10x_data/Processed_data/Radius_of_gyration")

name_mouse=name_mouse[isUnique(as.vector(name_mouse$initial_ensg))&isUnique(as.vector(name_mouse$ortholog_name)),]
marker_plot_all_1=rabbit_id[rabbit_id%in%name_mouse$initial_ensg]
name_rabbit=as.vector(name_mouse$ortholog_name)
name_rabbit[name_rabbit=="N/A"]=as.vector(name_mouse$initial_ensg[name_rabbit=="N/A"])

#setwd(paste0(base_path,"10x_data/Output_script"))
#name_mouse=read.csv(file=geni_ortogoli_g_profiler)
#name_mouse_no=name_mouse[!(isUnique(as.vector(name_mouse$initial_ensg))&isUnique(as.vector(name_mouse$ortholog_name))),]
#marker_plot_all_2=rabbit_id[rabbit_id%in%name_mouse_no$initial_ensg]
#name_rabbit_no=unique(as.vector(name_mouse_no$initial_ensg))
marker_plot_all_2=rabbit_id[!rabbit_id%in%name_mouse$initial_ensg]
name_rabbit_no=rabbit_id[!rabbit_id%in%name_mouse$initial_ensg]
ortogoli_name_final=c(name_rabbit,name_rabbit_no)
rabbit_id_final=c(marker_plot_all_1,marker_plot_all_2)
return(list(ortogoli_name_final,rabbit_id_final))}
```


```{r }
Test_Fisher=function(marker_all_cluster,marker_specific,background,cluster){
level=levels(as.factor(cluster))
p_value=rep(0,length(level))
odds_ratio=rep(0,length(level))
for (i in 1:length(level)){
show=build_fisher_test(marker_all_cluster[names(marker_all_cluster)==level[i]],marker_specific,background)
p_value[i]=as.numeric(show$p.value)
names(p_value)[i]=level[i]
odds_ratio[i]=as.numeric(show$estimate)
names(odds_ratio)[i]=level[i]
}
return(list(p_value,odds_ratio))
}

```



```{r }

setwd("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/github_january_2022/CIARA_archive/CIARA_old/inst/extdata")
load("knn_matrix_heart_new.Rda")
load("norm_data_heart.Rda")
load("raw_data_heart.Rda")
load("cluster_heart_small.Rda")
load("umap_cordinate_heart_small.Rda")


#load(system.file("extdata", "knn_matrix_heart_new.Rda", package ="CIARA"))
#load(system.file("extdata", "norm_data_heart.Rda", package ="CIARA"))
#load(system.file("extdata", "raw_data_heart.Rda", package ="CIARA"))
#load(system.file("extdata", "cluster_heart_small.Rda", package ="CIARA"))
#load(system.file("extdata", "umap_cordinate_heart_small.Rda", package ="CIARA"))


```

```{r }
#setwd("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/published_data")
#load("heart_hyper_new.Rda")
#result_old=result
```


```{r }
#load(system.file("extdata", "heart_december_2021.Rda", package ="CIARA"))
#result_old=result
#ciara_genes_old=row.names(result_old)[result_old[,1]<1]
#geni_top=row.names(result_old)[result_old[,2]<0.0001&result_old[,1]==0]
#geni_top_top=row.names(result_old)[order(as.numeric(result_old[,2]))]
```

## Load 

Rscript --vanilla /home/ies/gabriele.lubatti/ciara_marco/Entropy_server_december_2021_easy.R -n norm_data_heart -k knn_matrix_heart_new -o endoderm_a.Rda -r 1 -l 3 -h 30 -t 1 -p 0.001 -c 10 -a TRUE

Rscript --vanilla /home/ies/gabriele.lubatti/ciara_marco/Entropy_server_december_2021_easy.R -n norm_data_heart -k knn_matrix_heart_new -o endoderm.Rda -r 1 -l 3 -h 30 -t 1 -p 0.001 -c 10 -a FALSE

```{r }
#setwd("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/github_january_2022/CIARA_archive/CIARA_old/inst/extdata")
#load("heart_december_2021.Rda")

```

```{r }
setwd("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/github_january_2022/vignettes")
load("endoderm.Rda")
```


```{r }
ciara_genes_old=row.names(result)[result[,1]<1]
geni_top_top_old=row.names(result)[order(as.numeric(result[,1]))]
background_old=row.names(result)
```


```{r }
ciara_genes = ciara_genes_old
geni_top_top = geni_top_top_old
background = background_old
```


```{r }
setwd("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/github_january_2022/vignettes")
load("endoderm_a.Rda")
```


```{r }
ciara_genes=row.names(result)[result[,1]<1]
geni_top_top=row.names(result)[order(as.numeric(result[,1]))]
background=row.names(result)
```


```{r}
#setwd('/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/elmir_data/heart_data')
gene_id=row.names(raw_data_heart)
#setwd("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/elmir_data/heart_data")
#write.table(gene_id,file="geni_heart.txt",row.names = F,quote=F)
#setwd("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/elmir_data/heart_data")

setwd("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/github_january_2022/CIARA_archive/CIARA_old/inst/extdata")
name_mouse=read.csv("human_mouse_heart.csv")
#name_mouse=read.csv(system.file("extdata", "human_mouse_heart.csv", package ="CIARA"))


human_gene_name=geni_ortogoli_unique(gene_id,name_mouse)
human_name_all=human_gene_name[[1]]
human_id_all=human_gene_name[[2]]

data_all_human=data.frame(human_id_all,human_name_all)
row.names(data_all_human)=human_id_all
data_all_human_new=data_all_human[gene_id,]
human_name_final=as.vector(data_all_human_new[,2])

row.names(norm_data_heart)=human_name_final

row.names(raw_data_heart)=human_name_final

ciara_genes=data_all_human[ciara_genes,2]

background=data_all_human[background,2]

#geni_top=data_all_human[geni_top,2]
geni_top_top=data_all_human[geni_top_top,2]

```

```{r }


p=list()
for(i in (geni_top_top)[1:10]){
  q=plot_gene(norm_data_heart,umap_cordinate_heart_small,i,i)
  p=list(p,q)
}
p
```





```{r }
plot_genes_sum(umap_cordinate_heart_small,norm_data_heart,(ciara_genes),"Sum from top CIARA genes")

```





```{r }

norm_counts_small = apply(norm_data_heart[(ciara_genes), ], 
        1, function(x) {
            y = x/sum(x)
            return(y)
        })
    gene_sum = apply(norm_counts_small, 1, sum)
    
    
genes_name_text=selection_localized_genes(norm_data_heart,ciara_genes,min_number_cells=4,max_number_genes=4)
plot.elmir=umap_cordinate_heart_small
colnames(plot.elmir)=c("UMAP_1","UMAP_2")
plot_interactive(plot.elmir,gene_sum,genes_name_text,min_x=NULL,max_x=NULL,min_y=NULL,max_y=NULL)
```


##new part interactive plot

```{r }




localized_genes_mouse=detect_localized_genes(knn_matrix_heart_new,norm_data_heart,geni_top_top,100) 
rank_intersect=localized_genes_mouse[[2]]
list_intersect=localized_genes_mouse[[1]]

ramp <- colorRamp(c("white", "blue4"))




ramp <- colorRamp(c("white", "blue4"))
 

rank_intersect=localized_genes_mouse[[2]]
list_intersect=localized_genes_mouse[[1]]


ramp.list <- rgb( ramp(seq(0, 1, length = length(unique(rank_intersect)))), max = 255)

index_color=round(length(ramp.list)/2,0)


breaks = seq(0,max(rank_intersect),length.out=1000)
gradient1 = colorpanel( sum( breaks[-1]<= as.numeric(quantile(breaks,0.25))), "#FFFFFF",ramp.list[index_color])

gradient2 = colorpanel( sum( breaks[-1] > as.numeric(quantile(breaks,0.25)) ), ramp.list[index_color], "#00008B" )


hm.colors = c(gradient1,gradient2)
 





genes_name_text=selection_localized_genes_here(list_intersect ,geni_top_top,max_number_genes=5)

setwd("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/github_january_2022/figures_paper/interactive_plot")
p=plot_interactive_here(umap_cordinate_heart_small,norm_data_heart,rank_intersect,genes_name_text,hm.colors,min_x=NULL,max_x=NULL,min_y=NULL,max_y=NULL)
htmlwidgets::saveWidget(p, file = "mouse_after_gastrula.html")

pdf("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/github_january_2022/figures_paper/interactive_plot/mouse_after_gastrula.pdf",width = 8,height = 6)
p=plot_genes_sum_here(umap_cordinate_heart_small,norm_data_heart,rank_intersect,"Top 100 genes CIARA",hm.colors)
p+theme_bw()+theme(panel.border = element_blank(), panel.grid.major = element_blank(),
panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"),axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),axis.text.y=element_blank(),
        axis.ticks.y=element_blank())+theme(text = element_text(size=12))+geom_point( alpha = 0.3, na.rm = T, shape = 21, colour = "black",stroke=0.1)
dev.off()






```




## Cluster analysis using ciara of mixing (step 2)
```{r }


heart_seurat_ciara=cluster_analysis_integrate_rare(raw_data_heart,"Elmir data",0.1,5,30,ciara_genes)

#heart_seurat_ciara_2=Cluster_analysis_integrate_rare(raw_data_heart,"Elmir data",0.1,3,30,ciara_genes_old)

#heart_seurat_ciara=heart_seurat_ciara_2
heart_seurat_ciara=FindClusters(heart_seurat_ciara,resolution = 0.2)

```

```{r }


find_resolution(heart_seurat_ciara,seq(0.1,1,0.1))
```

```{r }
heart_seurat_ciara=FindClusters(heart_seurat_ciara,resolution = 0.2)
```


```{r }
plot_umap(umap_cordinate_heart_small,heart_seurat_ciara$RNA_snn_res.0.2)
```


```{r}
names(cluster_heart_small)=colnames(raw_data_heart)
cluster_ciara_heart=as.vector(heart_seurat_ciara$RNA_snn_res.0.2)
names(cluster_ciara_heart)=colnames(raw_data_heart)
final_cluster_heart=merge_cluster(cluster_heart_small,cluster_ciara_heart,30)
```





```{r }
pdf("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/github_january_2022/figures_paper/figure_4/figure_4_e.pdf")
plot_umap(umap_cordinate_heart_small,final_cluster_heart)
dev.off()

```

```{r }

final_cluster_heart_new=final_cluster_heart
final_cluster_heart_new[final_cluster_heart=="2-step_2"]="Rare Endoderm"
final_cluster_heart_new[final_cluster_heart=="En1"]="Endoderm1"
final_cluster_heart_new[final_cluster_heart=="En2"]="Endoderm2"

pdf("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/github_january_2022/figures_paper/figures_conference/umap_endoderm_conference.pdf",width = 8, height = 6)
p=plot_umap(umap_cordinate_heart_small,final_cluster_heart_new)
p[[1]]+theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"),axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),axis.text.y=element_blank(),
        axis.ticks.y=element_blank())+theme(text = element_text(size=18))
dev.off()

```

```{r }
table(final_cluster_heart)
```



```{r}
DefaultAssay(heart_seurat_ciara) <- "RNA"
markers_first_elmir=markers_cluster_seurat(heart_seurat_ciara,final_cluster_heart,names(heart_seurat_ciara$RNA_snn_res.0.1),50)


marker_heart_top=markers_first_elmir[[1]]
marker_heart=markers_first_elmir[[3]]
```

```{r}

top_heart=white_black_markers(final_cluster_heart,"2-step_2",norm_data_heart,marker_heart,0)
top_heart=names(top_heart)[top_heart]

mean_top_heart=apply(norm_data_heart[top_heart,final_cluster_heart=="2-step_2"],1,mean)
mean_top_heart=sort(mean_top_heart,decreasing = T)

top_heart=names(mean_top_heart)
names(top_heart)=rep("2-step_2",length(top_heart))
```

```{r}

plot_balloon_marker(norm_data_heart,final_cluster_heart,marker_heart_top,10,max_size=2,text_size=9)
pdf("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/github_january_2022/figures_paper/figure_4/figure_4_f.pdf")

plot_balloon_marker(norm_data_heart,final_cluster_heart,top_heart,40,max_size=2,text_size=9)
dev.off()
```









```{r }
setwd("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/github_january_2022/CIARA_archive/CIARA_old/inst/extdata")
load("markers_first_elmir_final.Rda")
#load(system.file("extdata", "markers_first_elmir_final.Rda", package ="CIARA"))
#setwd("/Users/gabriele.lubatti/Desktop/Phd/ciara_of_mixing_library/elmir_data")
#load(file="markers_first_elmir_final.Rda")
marker_elmir_all=markers_first_elmir_final[[3]]
marker_endoderm_2=marker_elmir_all[names(marker_elmir_all)=="Endoderm_2"]





```





```{r }
result=Test_Fisher(marker_heart,marker_endoderm_2,as.vector(marker_heart),as.vector(final_cluster_heart))




names(result[[1]])[result[[1]]<0.05&result[[2]]>2]

geni_comuni=intersect(marker_heart[names(marker_heart)=="2-step_2"],marker_endoderm_2)
```

```{r }
plot_balloon_marker_here=function (norm_counts, cluster, marker_complete, max_number, 
    max_size = 5, text_size = 7) 
{
    livelli <- levels(factor(cluster))
    all_markers_plot <- rep(list(0), length(livelli))
    all_markers <- rep(list(0), length(livelli))
    for (i in 1:length(livelli)) {
        genes_4_0 <- marker_complete[(names(marker_complete) == 
            livelli[i])]
        if (length(genes_4_0) > 0) {
            genes_4_0 <- genes_4_0[1:max_number]
            all_markers[[i]] <- genes_4_0
            #genes_4_0_plot <- paste(livelli[i], genes_4_0, sep = "_")
            genes_4_0_plot <- genes_4_0
            all_markers_plot[[i]] <- genes_4_0_plot
        }
    }
    all_markers <- lapply(all_markers, function(x) {
        x[x != 0]
    })
    all_markers_plot <- lapply(all_markers_plot, function(x) {
        x[x != 0]
    })
    all_markers_values <- rep(list(0), length(livelli))
    for (i in 1:length(livelli)) {
        gene_values_0 <- apply(norm_counts[unlist(all_markers, 
            use.names = FALSE), cluster == livelli[i]], 1, mean)
        all_markers_values[[i]] <- gene_values_0
    }
    values <- rep(list(0), length(livelli))
    for (i in 1:length(livelli)) {
        valore <- apply(norm_counts[unlist(all_markers, use.names = FALSE), 
            cluster == livelli[i]], 1, function(x) {
            sum(x > 1)/length(x)
        })
        values[[i]] <- valore
    }
    cluster_label_all <- rep(list(0), length(livelli))
    for (i in 1:length(livelli)) {
        cluster_label_all[[i]] <- rep(livelli[i], length(unlist(all_markers)))
        #cluster_label_all[[i]] <- rep(livelli[i], length(unlist(all_markers)))
    }
    all_markers_plot <- unlist(all_markers_plot)
    cluster_label_all <- unlist(cluster_label_all)
    values <- unlist(values)
    all_markers_values <- unlist(all_markers_values)
    balloon_melted <- data.frame(all_markers_plot, cluster_label_all, 
        values)
    colnames(balloon_melted) <- c("all_markers_plot", "cluster_label_all", 
        "values")
    p <- ggplot2::ggplot(balloon_melted, ggplot2::aes(x = cluster_label_all, 
        y = all_markers_plot))
    p + ggplot2::geom_point(ggplot2::aes(size = values, colour = as.numeric(all_markers_values))) + 
        ggplot2::theme(panel.background = ggplot2::element_blank(), 
            panel.border = ggplot2::element_rect(colour = "blue", 
                fill = NA, size = 3)) + ggplot2::scale_size_area(max_size = max_size) + 
        ggplot2::scale_colour_gradient(low = "white", high = "midnight blue", 
            na.value = NA) + ggplot2::labs(colour = "Mean", 
        size = "Value") + 
        ggplot2::xlab("Cluster") + ggplot2::ylab("Markers") + 
        ggplot2::theme(text = ggplot2::element_text(size = text_size), 
            axis.text.x = ggplot2::element_text(angle = 45, vjust = 1, 
                hjust = 1,size=14), axis.title.x = ggplot2::element_blank())
}
```


```{r }
names(geni_comuni)=rep("Rare Endoderm",length(geni_comuni))
pdf("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/github_january_2022/figures_paper/figures_conference/marker_endoderm_conference.pdf",width = 8, height = 6)
plot_balloon_marker_here(norm_data_heart,final_cluster_heart_new,geni_comuni,10,max_size=8,text_size=18)
dev.off()
```


```{r }

listInput <- list(mouse_rare_endoderm=marker_heart[names(marker_heart)=="2-step_2"],human_rare_endoderm=marker_endoderm_2)
pdf("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/github_january_2022/figures_paper/figure_4/figure_4_g.pdf")
upset(fromList(listInput), order.by = "freq")
dev.off()

#listInput <- list(ciara_mixing_rare_2=marker_2_small,giniclust2_rare_2=marker_2_original)

#upset(fromList(listInput), order.by = "freq")









```




```{r }
p=list()
for(i in (geni_comuni)[1:10]){
  q=plot_gene(norm_data_heart,umap_cordinate_heart_small,i,i)
  p=list(p,q)
}
p
```


