---
title: "mouse_esc"
author: "Gabriele Lubatti"
date: "10/22/2021"
output: html_document
---


```{r } 
gc()
rm(list=ls())
#library(devtools)
#install_github("ScialdoneLab/CIARA",auth_token="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",ref="master")
```


```{r } 
#library(Biobase)
#library(Seurat)
#library(clustree)
#library(plotly)
#library(ggplot2)
library(CIARA)
library(UpSetR)
library(Seurat)
library(clustree)
```

```{r }
detect_localized_genes=function(knn_matrix_ane_24_update,norm_counts_ane_24_update,geni_top_top,top_number){
list_intersect=rep(list(0),length(colnames(knn_matrix_ane_24_update)))
rank_intersect=rep(0,length(colnames(knn_matrix_ane_24_update)))
for ( i in 1:length(colnames(knn_matrix_ane_24_update))){
  knn_name=names(which(knn_matrix_ane_24_update[i,]==1))
  geni_select=geni_top_top[1:top_number]
  sum_knn=apply(norm_counts_ane_24_update[geni_select,knn_name],2,function(x){
    somma=geni_select[which(x>1)]
    return(somma)
  })
    rank_intersect[i]=sum(duplicated(as.vector(unlist(sum_knn))))
    list_intersect[[i]]=unique(as.vector(unlist(sum_knn))[duplicated(as.vector(unlist(sum_knn)))])

}
return(list(list_intersect,rank_intersect))
}
```

```{r }

selection_localized_genes_here=function (list_intersect, geni_top_top, 
    max_number_genes = 5){ 
text=rep(0,length(list_intersect))

for (i in 1:length(list_intersect)){
  if (length(list_intersect[[i]])==0){
    localized_genes="No specific genes"
    text[i] <- "No specific genes"
   
  }
  else {
  localized_genes=geni_top_top[which(geni_top_top %in% list_intersect[[i]])]
  
  max_genes=min(max_number_genes,length(localized_genes))
  if (max_genes<max_number_genes){
    warning("max_number bigger than length of localized_genes. Set max_number_genes = length(localized_genes) ")
  }
  localized_genes=localized_genes[1:max_genes]
  text[i]=paste(localized_genes, collapse = "-")
  
  
  }
}
return(text)
}


```

```{r }
plot_interactive_here=function (coordinate_umap, norm_counts,color, text,ramp_list, min_x = NULL, max_x = NULL, 
    min_y = NULL, max_y = NULL)
  {
  
  
  index_sort <- order(color)
  text=text[index_sort]
    row.names(coordinate_umap) <- colnames(norm_counts)
    coordinate_umap <- coordinate_umap[colnames(norm_counts)[index_sort], 
        ]
    color <- sort(color)
    

    if (!requireNamespace("plotly", quietly = TRUE)) {
        stop("Package plotly needed for interactive == TRUE. Please install it: install.packages('plotly')")
    }
    colnames(coordinate_umap) <- c("UMAP_1", "UMAP_2")
    fig <- plotly::plot_ly(data = coordinate_umap, x = ~UMAP_1, 
        y = ~UMAP_2, color = ~color, type = "scatter", mode = "markers", 
        marker = list(size = 5, width = 2), text = ~text, hoverinfo = "text",colors=ramp_list)
   fig <- fig %>% 
layout(plot_bgcolor='#e5ecf6')
    
    if (!is.null(min_x)) {
        fig <- fig %>% plotly::layout(xaxis = list(range = c(min_x, 
            max_x)), yaxis = list(range = c(min_y, max_y))) %>% 
layout(plot_bgcolor='rgb(254, 247, 234)') %>% 
layout(paper_bgcolor='rgb(254, 247, 234)')
    }
    return(fig)
    }
    
```

```{r }
plot_genes_sum_here=function (coordinate_umap, norm_counts, color_cell, name_title) 
{
    index_sort <- order(color_cell)
    row.names(coordinate_umap) <- colnames(norm_counts)
    coordinate_umap <- coordinate_umap[colnames(norm_counts)[index_sort], 
        ]
    color_cell <- sort(color_cell)
    
    umap_plot <- ggplot2::ggplot(coordinate_umap, ggplot2::aes(coordinate_umap[, 
        1], coordinate_umap[, 2])) + ggplot2::geom_point(ggplot2::aes(colour = color_cell)) + 
        ggplot2::xlab("UMAP 1") + ggplot2::ylab("UMAP 2") + ggplot2::labs(col = "Number of localized genes") + 
        ggplot2::scale_colour_gradient(low = "white",high = "blue4") + 
        ggplot2::ggtitle(name_title)
    return((umap_plot))
}



```

```{r }

Plot_qc_cluster=function(seurat_object,cluster,mit_prefix,rib_prefix_1,rib_prefix_2){

raw_counts<- (GetAssayData(seurat_object, slot = "counts",assay="RNA"))
sum_umi=apply(raw_counts,2,sum)
sum_geni=apply(raw_counts,2,function(x){
  x=x[x!=0]
  return(length(x))
})


frac_mito=apply(raw_counts,2,function(x){
  #return(sum(x[grep("MT-",row.names(raw_counts))])/sum(x))
  return(sum(x[grep(mit_prefix,row.names(raw_counts))])/sum(x))
  
  
  
})


frac_ribo=apply(raw_counts,2,function(x){
  #frac_ribo_s=x[grep("RPS",row.names(raw_counts))]
  #frac_ribo_l=x[grep("RPL",row.names(raw_counts))]
  frac_ribo_s=x[grep(rib_prefix_1,row.names(raw_counts))]
  frac_ribo_l=x[grep(rib_prefix_2,row.names(raw_counts))]
  frac_ribo_all=c(frac_ribo_s,frac_ribo_l)
  return(sum(frac_ribo_all)/sum(x))



})


Cluster=as.factor(cluster)

data_plot=data.frame(as.numeric(frac_mito),as.numeric(sum_geni),as.numeric(sum_umi),as.numeric(frac_ribo),Cluster)

library(ggplot2)

plot_1=ggplot(data_plot, aes(x=Cluster, y=frac_mito,fill=Cluster)) + 
  geom_boxplot()+xlab("clusters")+ylab("frac mitochindrial reads")+ggtitle("Frac mito reads")


plot_2=ggplot(data_plot, aes(x=Cluster, y=sum_umi,fill=Cluster)) + 
  geom_boxplot()+xlab("clusters")+ylab("UMI counts")+ggtitle("UMI counts")

plot_3=ggplot(data_plot, aes(x=Cluster, y=sum_geni,fill=Cluster)) + 
  geom_boxplot()+xlab("clusters")+ylab("Number of genes")+labs(color="Cluster")+ggtitle("Number of genes")

plot_4=ggplot(data_plot, aes(x=Cluster, y=frac_ribo,fill=Cluster)) + 
  geom_boxplot()+xlab("clusters")+ylab("Frac ribosomial reads")+labs(color="Cluster")+ggtitle("Frac ribosomial reads")
return (list(plot_1,plot_2,plot_3,plot_4))
}

```

```{r}
Cluster_analysis_integrate_rare_update=function (raw_counts, batch_name_1, resolution, neighbors, max_dimension, 
    feature_genes = NULL) 
{
    batch_1 <- CreateSeuratObject(counts = raw_counts, project = batch_name_1)
    batch_1 <- subset(batch_1, subset = nFeature_RNA > 0)
    batch_1 <- NormalizeData(batch_1, verbose = FALSE)
    batch_1 <- FindVariableFeatures(batch_1, selection.method = "vst", 
        nfeatures = 2000)
    batch_combined = batch_1
    #batch_combined <- ScaleData(batch_combined, verbose = FALSE)
    batch_combined$frac_mito=frac_mito
batch_combined$frac_ribo=frac_ribo
batch_combined <- ScaleData(batch_combined, verbose = FALSE,vars.to.regress = c("nCount_RNA","frac_mito","frac_ribo"))
    batch_combined <- RunPCA(batch_combined, npcs = max_dimension, 
        verbose = FALSE, features = feature_genes)
    batch_combined = RunUMAP(batch_combined, reduction = "pca", 
        dims = 1:20, set.seed = 42)
    batch_combined <- FindNeighbors(batch_combined, reduction = "pca", 
        dims = 1:max_dimension, k.param = neighbors)
    batch_combined <- FindClusters(batch_combined, resolution = resolution)
    return(batch_combined)
}
```


##Start with dataset at 24 h



```{r }
##LAST VERSION condition srun -p normal_q -c 10 --mem=100G -t 1:00:00 --pty bash
#Rscript --vanilla /home/ies/gabriele.lubatti/entropy_mixing/ane_data/entropy_server_december_2021_easy.R -n norm_counts_ane_24_update -k knn_matrix_ane_24_update -o ane_december_2021.Rda -r 1 -l 3 -h 20 -t 1 -p 0.001 -c 10 -a FALSE

#  Time difference of Time difference of 1.506383 mins


#Rscript --vanilla /home/ies/gabriele.lubatti/entropy_mixing/ane_data/entropy_server_december_2021_easy.R -n norm_counts_ane_24_update -k knn_matrix_ane_24_update -o ane_december_2021_a.Rda -r 1 -l 3 -h 20 -t 1 -p 0.001 -c 10 -a TRUE

# Time difference of 17.48768 secs

```




```{r }

setwd("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/github_january_2022/CIARA_archive/CIARA_old/inst/extdata")
load("raw_counts_ane_24_update.Rda")
#load(system.file("extdata", "raw_counts_ane_24_update.Rda", package ="CIARA"))


```


```{r }
load("/Users/gabriele.lubatti/Desktop/Phd/Early_embryonic_like_cells/Mayra_project/final_cluster_version_sub.Rda")
```


```{r }
mayra_seurat_24=cluster_analysis_integrate_rare(raw_counts_ane_24_update,"Mayra_data_24",0.1,5,30)
#a=Plot_qc_cluster(mayra_seurat_24,as.vector(mayra_seurat_24$RNA_snn_res.0.1),"mt-","Rps","Rpl")
pdf("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/github_january_2022/figures_paper/sup_figure_3/sup_fig_3_c.pdf")
a[[1]]
dev.off()
pdf("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/github_january_2022/figures_paper/sup_figure_3/sup_fig_3_b.pdf")
a[[2]]
dev.off()
pdf("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/github_january_2022/figures_paper/sup_figure_3/sup_fig_3_d.pdf")
a[[3]]
dev.off()

```

```{r }
#norm_counts_ane_24=as.matrix(GetAssayData(mayra_seurat_24, slot = "data",assay="RNA"))

cordinate_umap_24=as.data.frame(Embeddings(mayra_seurat_24, reduction = "umap")[, 1:2])
pdf("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/github_january_2022/figures_paper/sup_figure_3/sup_fig_3_a.pdf")
plot_umap(cordinate_umap_24,mayra_seurat_24$RNA_snn_res.0.1)
dev.off()
```


```{r }

frac_mito=apply(raw_counts_ane_24_update,2,function(x){
  #return(sum(x[grep("MT-",row.names(raw_counts))])/sum(x))
  return(sum(x[grep("mt-",row.names(raw_counts_ane_24_update))])/sum(x))
  
  
  
})


frac_ribo=apply(raw_counts_ane_24_update,2,function(x){
  #frac_ribo_s=x[grep("RPS",row.names(raw_counts))]
  #frac_ribo_l=x[grep("RPL",row.names(raw_counts))]
  frac_ribo_s=x[grep("Rps",row.names(raw_counts_ane_24_update))]
  frac_ribo_l=x[grep("Rpl",row.names(raw_counts_ane_24_update))]
  frac_ribo_all=c(frac_ribo_s,frac_ribo_l)
  return(sum(frac_ribo_all)/sum(x))



})
```


```{r }
mayra_seurat_24_update=Cluster_analysis_integrate_rare_update(raw_counts_ane_24_update,"Mayra_data_24",0.1,5,30)
norm_counts_ane_24_update=as.matrix(GetAssayData(mayra_seurat_24_update, slot = "data",assay="RNA"))
knn_matrix_ane_24_update=as.matrix(mayra_seurat_24_update@graphs$RNA_nn)
cordinate_umap_24_update=as.data.frame(Embeddings(mayra_seurat_24_update, reduction = "umap")[, 1:2])
```

```{r }
find_resolution(mayra_seurat_24_update,seq(0.1,0.5,0.1))
```

```{r }
FindClusters(mayra_seurat_24_update, resolution = 0.2)
```

```{r}

initial_cluster_24=as.vector(mayra_seurat_24_update$RNA_snn_res.0.1)

#initial_cluster_24_2=as.vector(mayra_seurat_24_update$RNA_snn_res.0.2)

```
```{r}
pdf("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/github_january_2022/figures_paper/figure_3/fig_3_a.pdf")
plot_umap(cordinate_umap_24_update,initial_cluster_24)
dev.off()

mayra_seurat_24_update=FindClusters(mayra_seurat_24_update,resolution = 0.2)
plot_umap(cordinate_umap_24_update,mayra_seurat_24_update$RNA_snn_res.0.2)
```

## Start with approx = TRUE and approx = FALSE
Rscript --vanilla /home/ies/gabriele.lubatti/ciara_marco/Entropy_server_december_2021_easy.R -n norm_counts_ane_24_update -k knn_matrix_ane_24_update -o ane_24_a.Rda -r 1 -l 3 -h 20 -t 1 -p 0.001 -c 10 -a TRUE

Rscript --vanilla /home/ies/gabriele.lubatti/ciara_marco/Entropy_server_december_2021_easy.R -n norm_counts_ane_24_update -k knn_matrix_ane_24_update -o ane_24.Rda -r 1 -l 3 -h 20 -t 1 -p 0.001 -c 10 -a FALSE


```{r }
#setwd("/Users/gabriele.lubatti/Desktop/Phd/Early_embryonic_like_cells/Mayra_project")
#load("ane_december_2021.Rda")

```

```{r }
setwd("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/github_january_2022/vignettes")
load("ane_24_a.Rda")
```

```{r }
ciara_genes=row.names(result)[result[,1]<1]
#geni_top=row.names(result)[result[,1]==0&result[,2]<0.0001]
geni_top_top=row.names(result)[order(as.numeric(result[,1]))]
```

```{r }
background=row.names(result)
```


```{r }
setwd("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/github_january_2022/vignettes")
load("ane_24.Rda")
```

```{r }
#load("/Users/gabriele.lubatti/Desktop/Phd/ciara_of_mixing_library/github_january_2022/ane_dec#ember_2021.Rda")
```

```{r }
ciara_genes_old=row.names(result)[result[,1]<1]
#geni_top=row.names(result)[result[,1]==0&result[,2]<0.0001]
geni_top_top_old=row.names(result)[order(as.numeric(result[,1]))]
```

```{r }
background_old=row.names(result)
```

```{r }
background=background_old
ciara_genes=ciara_genes_old
geni_top_top=geni_top_top_old

```

```{r }
library(ggplot2)
p=list()
for(i in (geni_top_top)[1:10]){
  q=plot_gene(norm_counts_ane_24_update,cordinate_umap_24_update,i,i)
  p=list(p,q)
}
p
```

```{r }
plot_gene(norm_counts_ane_24_update,cordinate_umap_24_update,geni_top_top[length(geni_top_top)],geni_top_top[length(geni_top_top)])
```


```{r }
plot_gene(norm_counts_ane_24_update,cordinate_umap_24_update,"Gata6","Gata6")
```


```{r }

plot_genes_sum(cordinate_umap_24_update,norm_counts_ane_24_update,(ciara_genes),"Sum from top ciara genes")

```










```{r }
localized_genes_mouse=detect_localized_genes(knn_matrix_ane_24_update,norm_counts_ane_24_update,geni_top_top,100)  

ramp <- colorRamp(c("white", "blue4"))
 

rank_intersect=localized_genes_mouse[[2]]
list_intersect=localized_genes_mouse[[1]]

#ramp.list <- rgb( ramp(seq(0, 1, length = length(rank_intersect)), max = 255))

ramp.list <- rgb( ramp(seq(0, 1, length = length(unique(rank_intersect)))), max = 255)

index_color=round(length(ramp.list)/2,0)
#index_color=index_color*9


breaks = seq(0,max(rank_intersect),length.out=1000)
gradient1 = colorpanel( sum( breaks[-1]<= as.numeric(quantile(breaks,0.25))), "#FFFFFF",ramp.list[index_color])

gradient2 = colorpanel( sum( breaks[-1] > as.numeric(quantile(breaks,0.25)) ), ramp.list[index_color], "#00008B" )


hm.colors = c(gradient1,gradient2)




genes_name_text=selection_localized_genes_here(list_intersect ,geni_top_top,max_number_genes=5)

plot_interactive_here(cordinate_umap_24_update,norm_counts_ane_24_update,rank_intersect,genes_name_text,hm.colors,min_x=NULL,max_x=NULL,min_y=NULL,max_y=NULL)

plot_genes_sum_here(cordinate_umap_24_update,norm_counts_ane_24_update,rank_intersect,"Top genes CIARA")

```

## Old
```{r }

plot_genes_sum(cordinate_umap_24_update,norm_counts_ane_24_update,(ciara_genes),"Sum from top ciara genes")

```










## Cluster analysis using CIARA (step 2)

```{r }
#mayra_seurat_ciara_24=Cluster_analysis_integrate_rare(mayra_dati_raw_24,"24h",0.01,5,30,names(entro#py_genes))
mayra_seurat_ciara_24=Cluster_analysis_integrate_rare_update(raw_counts_ane_24_update,"24h",0.1,3,30,(ciara_genes))


#mayra_seurat_ciara_24_2=Cluster_analysis_integrate_rare_update(raw_counts_ane_24_update,"24h",0.1,5,30,(ciara_genes))

#mayra_seurat_ciara_24=FindClusters(mayra_seurat_ciara_24,resolution = 0.1)

```


```{r }
find_resolution(mayra_seurat_ciara_24,seq(0.01,0.5,0.1))
```



```{r }

#mayra_seurat_ciara_24=mayra_seurat_ciara_24_2
#mayra_seurat_ciara_24=FindClusters(mayra_seurat_ciara_24,resolution = 0.2)

norm_mayra_ciara_24=as.matrix(GetAssayData(mayra_seurat_ciara_24, slot = "data",assay="RNA"))



cordinate_umap_24_ciara=as.data.frame(Embeddings(mayra_seurat_ciara_24, reduction = "umap")[, 1:2])


```


```{r }
plot_umap(cordinate_umap_24_ciara,mayra_seurat_ciara_24$RNA_snn_res.0.1)
```

```{r}
table(mayra_seurat_ciara_24$RNA_snn_res.0.1)
```

```{r }
plot_umap(cordinate_umap_24_update,mayra_seurat_ciara_24$RNA_snn_res.0.1)
```

```{r }
cluster_original=as.vector(mayra_seurat_24_update$RNA_snn_res.0.1)
names(cluster_original)=names(mayra_seurat_24_update$RNA_snn_res.0.1)
cluster_sub=as.vector(mayra_seurat_ciara_24$RNA_snn_res.0.1)
names(cluster_sub)=names(mayra_seurat_ciara_24$RNA_snn_res.0.1)
final_cluster_version_sub=merge_cluster(cluster_original,cluster_sub,25)
```



```{r }
pdf("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/github_january_2022/figures_paper/figure_3/fig_3_b.pdf")
plot_umap(cordinate_umap_24_update,final_cluster_version_sub)
dev.off()
```

```{r }
pdf("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/github_january_2022/figures_paper/figures_conference/umap_esc_24.pdf",width=8,height = 6)
p=plot_umap(cordinate_umap_24_update,final_cluster_version_sub)
p[[1]]+theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"),axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),axis.text.y=element_blank(),
        axis.ticks.y=element_blank())+theme(text = element_text(size=18))
dev.off()
```


```{r}
DefaultAssay(mayra_seurat_24_update) <- "RNA"
markers_first_elmir=markers_cluster_seurat(mayra_seurat_24_update,final_cluster_version_sub,names(mayra_seurat_24_update$RNA_snn_res.0.1),5)
```



```{r}
marker_plot_day_24_first=markers_first_elmir[[1]]

marker_plot_plot_day_24_first=markers_first_elmir[[2]]
marker_all_day_24_first=markers_first_elmir[[3]]

```




```{r}
 markers_mega <- Seurat::FindMarkers(mayra_seurat_24_update, ident.1 = names(mayra_seurat_24_update$RNA_snn_res.0.1)[final_cluster_version_sub == 
            0], ident.2 = names(mayra_seurat_24_update$RNA_snn_res.0.1)[final_cluster_version_sub != 0], 
            only.pos = T)
markers_mega <- markers_mega[markers_mega$p_val <= 0.05, 
            ]
        markers_mega <- markers_mega[order(markers_mega$p_val_adj), 
            ]
        markers_mega <- row.names(markers_mega)

markers_mega=markers_mega[!markers_mega%in%marker_all_day_24_first]
names(markers_mega)=rep("0",length(markers_mega))
marker_all_day_24_first_all=c(marker_all_day_24_first,markers_mega)
```


```{r }
top_0=white_black_markers(final_cluster_version_sub,0,norm_counts_ane_24_update,marker_all_day_24_first_all,0)

top_0=names(top_0)[top_0]

mean_top_0=apply(norm_counts_ane_24_update[top_0,final_cluster_version_sub==0],1,mean)
mean_top_0=sort(mean_top_0,decreasing = T)

top_0=names(mean_top_0)
names(top_0)=rep(0,length(top_0))
```

```{r }
top_1=white_black_markers(final_cluster_version_sub,"1-step_2",norm_counts_ane_24_update,marker_all_day_24_first,0)

top_1=names(top_1)[top_1]

mean_top_1=apply(norm_counts_ane_24_update[top_1,final_cluster_version_sub=="1-step_2"],1,mean)
mean_top_1=sort(mean_top_1,decreasing = T)

top_1=names(mean_top_1)
names(top_1)=rep("1-step_2",length(top_1))
```

```{r }
top_2=white_black_markers(final_cluster_version_sub,"2-step_2",norm_counts_ane_24_update,marker_all_day_24_first,0)
top_2=names(top_2)[top_2]

mean_top_2=apply(norm_counts_ane_24_update[top_2,final_cluster_version_sub=="2-step_2"],1,mean)
mean_top_2=sort(mean_top_2,decreasing = T)

top_2=names(mean_top_2)
names(top_2)=rep("2-step_2",length(top_2))
```

```{r }
plot_balloon_marker(norm_counts_ane_24_update,final_cluster_version_sub,c(top_0[1:5],c(top_1[1:4],top_1[8]),top_2[1:5]),5,max_size=5,text_size=10)
```

```{r }
pdf("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/github_january_2022/figures_paper/figures_conference/marker_esc_24.pdf",width=8,height = 6)
plot_balloon_marker(norm_counts_ane_24_update,final_cluster_version_sub,c(top_0[1:5],c(top_1[1:4],top_1[8]),top_2[1:5]),5,max_size=5,text_size=10)
dev.off()
```


```{r }
top_3=white_black_markers(final_cluster_version_sub,"3-step_2",norm_counts_ane_24_update,marker_all_day_24_first,0)
top_3=names(top_3)[top_3]

mean_top_3=apply(norm_counts_ane_24_update[top_3,final_cluster_version_sub=="3-step_2"],1,mean)
mean_top_3=sort(mean_top_3,decreasing = T)

top_3=names(mean_top_3)
names(top_3)=rep("3-step_2",length(top_3))
```

```{r }
pdf("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/github_january_2022/figures_paper/figure_3/fig_3_c.pdf")
plot_balloon_marker(norm_counts_ane_24_update,final_cluster_version_sub,top_1,30,max_size=5,text_size=10)
dev.off()
#Plot_balons(norm_counts_ane_24_update,final_cluster_version_sub,marker_all_day_24_first[names(marker_all_day_24_first)=="1-step_2"],30,max_size=5,text_size=10)

pdf("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/github_january_2022/figures_paper/figure_3/fig_3_d.pdf")
plot_balloon_marker(norm_counts_ane_24_update,final_cluster_version_sub,top_2,30,max_size=5,text_size=10)
dev.off()

plot_balloon_marker(norm_counts_ane_24_update,final_cluster_version_sub,top_3,30,max_size=5,text_size=10)



```
```{r}


white_black_7=white_black_markers(final_cluster_version_sub,"1-step_2",norm_counts_ane_24_update,marker_all_day_24_first,0)
sum(white_black_7)

white_black_7=white_black_markers(final_cluster_version_sub,"2-step_2",norm_counts_ane_24_update,marker_all_day_24_first,0)
sum(white_black_7)

```


## Marker at 48h


```{r}
setwd("/Users/gabriele.lubatti/Desktop/Phd/Early_embryonic_like_cells/Mayra_project")
cluster_mayra=read.csv('LIFcells_leidenClustering.csv')
cluster_mayra_48=cluster_mayra[as.vector(cluster_mayra$batch)=='48h',]
#row.names(cluster_mayra_48)=as.vector(cluster_mayra_48$cell)
#cell_name=as.vector(cluster_mayra_48$cell)
cluster_number_48=(cluster_mayra_48$cluster)
cluster_number_48=as.character(cluster_number_48)
#cell_name_new=substring(cell_name,1,nchar(cell_name)-5)

```



```{r}
#setwd("/Users/gabriele.lubatti/Desktop/Phd/Early_embryonic_like_cells/Mayra_project")
#load(file='mayra_dati_raw_48.Rda')
setwd("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/github_january_2022/CIARA_archive/CIARA_old/inst/extdata")
#load(system.file("extdata", "mayra_dati_raw_48.Rda", package ="CIARA"))
load("mayra_dati_raw_48.Rda")
mayra_seurat_48=cluster_analysis_integrate_rare(mayra_dati_raw_48,"Mayra_data_48",0.1,5,30)
norm_counts_ane_48=as.matrix(GetAssayData(mayra_seurat_48, slot = "data",assay="RNA"))
knn_matrix_ane_48=as.matrix(mayra_seurat_48@graphs$RNA_nn)
```


```{r}
#setwd("/Users/gabriele.lubatti/Desktop/Phd/Early_embryonic_like_cells/Mayra_project")
#load(file='mayra_dati_raw_48.Rda')

cordinate_umap_48=as.data.frame(Embeddings(mayra_seurat_48, reduction = "umap")[, 1:2])
```




```{r}
DefaultAssay(mayra_seurat_48) <- "RNA"
markers_first_elmir_48=markers_cluster_seurat(mayra_seurat_48,cluster_number_48,names(mayra_seurat_48$RNA_snn_res.0.1),20)
```



```{r}

marker_all_day_48_first=markers_first_elmir_48[[3]]

marker_1_small=marker_all_day_48_first[names(marker_all_day_48_first)==5]

marker_all_day_24_first_2=marker_all_day_24_first[names(marker_all_day_24_first)=="2-step_2"]

```


```{r }
Function_fisher=function(a,b,c){
matrice=matrix(c(sum(c[c%in%a]%in%c[c%in%b]),sum(c[c%in%a]%in%c[!c%in%b]),sum(c[!c%in%a]%in%c[c%in%b]),sum(c[!c%in%a]%in%c[!c%in%b])),nrow=2,ncol=2,byrow = T)
#Fisher test
fisher.test(matrice)
}
```

```{r}
Test_Fisher=function(ciara_cluster,markers_cluster,marker_top){
level=levels(as.factor(ciara_cluster))
p_value=rep(0,length(level))
odds_ratio=rep(0,length(level))
set_1=marker_top
for (i in 1:length(level)){
  set_2=markers_cluster[names(markers_cluster)==level[i]]
set_all=markers_cluster
show=Function_fisher(set_1,set_2,set_all)

p_value[i]=as.numeric(show$p.value)
names(p_value)[i]=level[i]
odds_ratio[i]=as.numeric(show$estimate)
names(odds_ratio)[i]=level[i]
}
return(list(p_value,odds_ratio))
}
test=Test_Fisher(final_cluster_version_sub,marker_all_day_24_first,marker_1_small)
test[[2]][test[[1]]<0.05&test[[2]]>2]
```

```{r }

listInput <- list(mouse_48h_differentiation=marker_1_small,mouse_24h_differentiation=marker_all_day_24_first_2)

pdf("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/github_january_2022/figures_paper/figure_3/fig_3_e.pdf")
upset(fromList(listInput), order.by = "freq")

dev.off()
```


##SCODA algorithm 
# Start here
# Day 0
```{r}
setwd("/Users/gabriele.lubatti/Desktop/Phd/Early_embryonic_like_cells/Mayra_project")
load(file='mayra_dati_raw_0.Rda')
mayra_seurat_0=cluster_analysis_integrate_rare(mayra_dati_raw_0,"Mayra_data_0",0.1,5,30)
#norm_counts_ane_0=as.matrix(GetAssayData(mayra_seurat_0, slot = "data",assay="RNA"))
#knn_matrix_ane_0=as.matrix(mayra_seurat_0@graphs$RNA_nn)

load("knn_matrix_ane_0.Rda")
load("norm_counts_ane_0.Rda")
setwd("/Users/gabriele.lubatti/Desktop/Phd/Early_embryonic_like_cells/Mayra_project")
vedo=load("entropy_ane_0.Rda")
cordinate_umap_0=as.data.frame(Embeddings(mayra_seurat_0, reduction = "umap")[, 1:2])
```

# Original cluster annotation
```{r}
setwd("/Users/gabriele.lubatti/Desktop/Phd/Early_embryonic_like_cells/Mayra_project")
cluster_mayra=read.csv('LIFcells_leidenClustering.csv')
cluster_mayra_0=cluster_mayra[as.vector(cluster_mayra$batch)=='0h',]
row.names(cluster_mayra_0)=as.vector(cluster_mayra_0$cell)
cell_name=as.vector(cluster_mayra_0$cell)
cluster_number=(cluster_mayra_0$cluster)
cluster_number_0=as.character(cluster_number)

cell_name_new=substring(cell_name,1,nchar(cell_name)-4)

norm_counts_ane_0=norm_counts_ane_0[,cell_name_new]
```



```{r }
plot_umap(cordinate_umap_0,cluster_number_0)
plot_gene(norm_counts_ane_0,cordinate_umap_0,"Gata6","Gata6")
plot_gene(norm_counts_ane_0,cordinate_umap_0,"Gata4","Gata4")

plot_gene(norm_counts_ane_0,cordinate_umap_0,"Zfp42","Zfp42")
plot_gene(norm_counts_ane_0,cordinate_umap_0,"Nanog","Nanog")
```

# Start here 12h
```{r}
setwd("/Users/gabriele.lubatti/Desktop/Phd/Early_embryonic_like_cells/Mayra_project")
load(file='mayra_dati_raw_12.Rda')
mayra_seurat_12=cluster_analysis_integrate_rare(mayra_dati_raw_12,"Mayra_data_12",0.1,5,30)
norm_counts_ane_12=as.matrix(GetAssayData(mayra_seurat_12, slot = "data",assay="RNA"))
knn_matrix_ane_12=as.matrix(mayra_seurat_12@graphs$RNA_nn)
#setwd("/Users/gabriele.lubatti/Desktop/Phd/Early_embryonic_like_cells/Mayra_project")
#save(norm_counts_ane_12,file="norm_counts_ane_12.Rda")
#save(knn_matrix_ane_12,file="knn_matrix_ane_12.Rda")

load("knn_matrix_ane_12.Rda")
load("norm_counts_ane_12.Rda")
setwd("/Users/gabriele.lubatti/Desktop/Phd/Early_embryonic_like_cells/Mayra_project")
load("entropy_ane_12.Rda")
cordinate_umap_12=as.data.frame(Embeddings(mayra_seurat_12, reduction = "umap")[, 1:2])
```


# Original cluster annotation
```{r}
setwd("/Users/gabriele.lubatti/Desktop/Phd/Early_embryonic_like_cells/Mayra_project")
cluster_mayra=read.csv('LIFcells_leidenClustering.csv')
cluster_mayra_12=cluster_mayra[as.vector(cluster_mayra$batch)=='12h',]
row.names(cluster_mayra_12)=as.vector(cluster_mayra_12$cell)
cell_name=as.vector(cluster_mayra_12$cell)
cluster_number=(cluster_mayra_12$cluster)
cluster_number_12=as.character(cluster_number)

cell_name_new=substring(cell_name,1,nchar(cell_name)-5)

norm_counts_ane_12=norm_counts_ane_12[,cell_name_new]
```



```{r }
plot_umap(cordinate_umap_12,cluster_number_12)

plot_gene(norm_counts_ane_12,cordinate_umap_12,"Gata6","Gata6")
plot_gene(norm_counts_ane_12,cordinate_umap_12,"Gata4","Gata4")
plot_gene(norm_counts_ane_12,cordinate_umap_12,"Zscan4c","Zscan4c")
```

# 48 h

```{r}
setwd("/Users/gabriele.lubatti/Desktop/Phd/Early_embryonic_like_cells/Mayra_project")
load(file='mayra_dati_raw_48.Rda')
load("knn_matrix_ane_48.Rda")
load("norm_counts_ane_48.Rda")
load("entropy_ane.Rda")

setwd("/Users/gabriele.lubatti/Desktop/Phd/Early_embryonic_like_cells/Mayra_project")
load(file='mayra_dati_raw_48.Rda')
mayra_seurat_48=cluster_analysis_integrate_rare(mayra_dati_raw_48,"Mayra_data_48",0.1,5,30)
cordinate_umap_48=as.data.frame(Embeddings(mayra_seurat_48, reduction = "umap")[, 1:2])
```

```{r}
setwd("/Users/gabriele.lubatti/Desktop/Phd/Early_embryonic_like_cells/Mayra_project")
cluster_mayra=read.csv('LIFcells_leidenClustering.csv')
cluster_mayra_48=cluster_mayra[as.vector(cluster_mayra$batch)=='48h',]
row.names(cluster_mayra_48)=as.vector(cluster_mayra_48$cell)
cell_name=as.vector(cluster_mayra_48$cell)
cluster_number_48=(cluster_mayra_48$cluster)
cluster_number_48=as.character(cluster_number_48)
cell_name_new=substring(cell_name,1,nchar(cell_name)-5)
```


```{r }
plot_umap(cordinate_umap_48,cluster_number_48)

plot_gene(norm_counts_ane_48,cordinate_umap_48,"Gata6","Gata6")
plot_gene(norm_counts_ane_48,cordinate_umap_48,"Gata4","Gata4")

plot_gene(norm_counts_ane_48,cordinate_umap_48,"Zfp42","Zfp42")
plot_gene(norm_counts_ane_48,cordinate_umap_48,"Nanog","Nanog")
```
```{r }
setwd("/Users/gabriele.lubatti/Desktop/Phd/Early_embryonic_like_cells/Mayra_project")
load(file="final_cluster_version_sub.Rda")
load("norm_counts_ane_24_update.Rda")
load(file="raw_counts_ane_24_update.Rda")
load(file="knn_matrix_ane_24_update.Rda")
load(file="cordinate_umap_24_update.Rda")
load(file="final_cluster_version_sub.Rda")
load("initial_cluster_24_2.Rda")

plot_gene(norm_counts_ane_24_update,cordinate_umap_24_update,"Zfp42","Zfp42")
plot_gene(norm_counts_ane_24_update,cordinate_umap_24_update,"Nanog","Nanog")
plot_gene(norm_counts_ane_24_update,cordinate_umap_24_update,"Sox2","Sox2")
plot_gene(norm_counts_ane_24_update,cordinate_umap_24_update,"Pou5f1","Pou5f1")
```


##Create input data for SCODA
```{r }

#summary_0=table(cluster_number_0)
#summary_12=table(cluster_number_12)
cluster_number_24=as.vector(final_cluster_version_sub)
cluster_number_24[cluster_number_24=="1-step_2"]="4"
cluster_number_24[cluster_number_24=="2-step_2"]="5"

#summary_24=table(final_cluster_version_sub)
#summary_48=table(cluster_number_48)

load("/Users/gabriele.lubatti/Desktop/Phd/Early_embryonic_like_cells/Mayra_project/final_cluster_version_sub.Rda")

full_names=c("0","1","2","3","4","5")
label_48=rep(0,length(full_names))
label_48[full_names%in%names(table(cluster_number_48))]=as.vector(table(cluster_number_48))[full_names%in%names(table(cluster_number_48))]

label_0=rep(0,length(full_names))
label_0[full_names%in%names(table(cluster_number_0))]=as.vector(table(cluster_number_0))

label_12=rep(0,length(full_names))
label_12[full_names%in%names(table(cluster_number_12))]=as.vector(table(cluster_number_12))


label_24=rep(0,length(full_names))
label_24[full_names%in%names(table(cluster_number_24))]=as.vector(table(cluster_number_24))

label_48=rep(0,length(full_names))
label_48[full_names%in%names(table(cluster_number_48))]=as.vector(table(cluster_number_48))


label_all=data.frame(label_0,label_12,label_24,label_48)

label_all=t(label_all)
colnames(label_all)=full_names


label_new_1=label_all[,1]+label_all[,2]+label_all[,3]
label_new_2=label_all[,4]+label_all[,5]
label_new_3=label_all[,6]
label_all_new=data.frame(label_new_1,label_new_2,label_new_3)


label_all_norm=apply(label_all_new,1,function(x){
  x=x/sum(x)
})

label_all_norm=t(label_all_norm)
colnames(label_all_norm)=c("Pluripotent","two_cells","Differentiating")

setwd("/Users/gabriele.lubatti/Desktop/Phd/Early_embryonic_like_cells/Mayra_project")
write.csv(label_all_norm,file="label_all_norm.csv")

label_all_complete=cbind(label_all_new,row.names(label_all_new))
colnames(label_all_complete)=c(colnames(label_all_new),"Time")

setwd("/Users/gabriele.lubatti/Desktop/Phd/Early_embryonic_like_cells/Mayra_project")
write.csv(label_all_complete,file="label_all_complete.csv",row.names = F)
```

```{r }
sum_time=apply(label_all_new,1,sum)
all_time_points=c(rep("Time 0h",sum_time[1]),rep("Time 12h",sum_time[2]),rep("Time 24h",sum_time[3]),rep("Time 48h",sum_time[4]))
all_time_cluster=c(cluster_number_0,cluster_number_12,cluster_number_24,cluster_number_48)
all_time_cluster[all_time_cluster==0|all_time_cluster==1|all_time_cluster==2]="ESC"
all_time_cluster[all_time_cluster==3|all_time_cluster==4]="2CLC"
all_time_cluster[all_time_cluster==5]="Precursor cells"
```

```{r}

barplot_cell_cycle=function(cell_cycle,cluster,title,fraction=TRUE){
dfForPlot <- data.frame(Phases =cell_cycle
                                ,clu =cluster)

if(fraction==TRUE){
pnm <-ggplot(dfForPlot, aes(y=cell_cycle,x=cluster))+geom_bar( aes(fill = cell_cycle,y = (..count..)/sum(..count..)),position="fill")+ ggtitle(title)+ylab("Fraction of cells")+xlab("Cluster")+labs(fill="Cluster annotation")+theme(text = element_text(size=10),
        axis.text.x = element_text(angle=90, hjust=1,size=6)) 
pnm}

if(fraction!=TRUE){
  pnm <-ggplot(dfForPlot, aes(y=cell_cycle,x=cluster))+geom_bar( aes(fill = cell_cycle,y = (..count..),position="fill"))+ ggtitle(title)+ylab("Number of cells")+xlab("Cluster")+labs(fill="Batch annotation")+theme(text = element_text(size=10),
        axis.text.x = element_text(angle=90, hjust=1,size=6)) 
pnm


}
return(list(pnm))}
```


##spatial origin ExE mesoderm
```{r }

library(ggplot2)
all_time_cluster=factor(all_time_cluster,levels=c("ESC","2CLC","Precursor cells"))
p=barplot_cell_cycle(all_time_cluster,all_time_points,"mouse ESCs",fraction = F)
p[[1]]+scale_fill_manual(values=CIARA:::gg_color_hue(3))
p=barplot_cell_cycle(all_time_cluster,all_time_points,"mouse ESCs",fraction = T)
p[[1]]+scale_fill_manual(values=CIARA:::gg_color_hue(3))






```
##Mouse ESC at day 0

```{r }
load("/Users/gabriele.lubatti/Desktop/Phd/Early_embryonic_like_cells/Mayra_project/result_ane_0.Rda")
setwd("/Users/gabriele.lubatti/Desktop/Phd/Early_embryonic_like_cells/Mayra_project")
load(file='mayra_dati_raw_0.Rda')
mayra_seurat_0=cluster_analysis_integrate_rare(mayra_dati_raw_0,"Mayra_data_0",0.1,5,30)
norm_counts_ane_0=as.matrix(GetAssayData(mayra_seurat_0, slot = "data",assay="RNA"))
#knn_matrix_ane_0=as.matrix(mayra_seurat_0@graphs$RNA_nn)

load("knn_matrix_ane_0.Rda")
load("norm_counts_ane_0.Rda")
#setwd("/Users/gabriele.lubatti/Desktop/Phd/Early_embryonic_like_cells/Mayra_project")
#vedo=load("entropy_ane_0.Rda")
cordinate_umap_0=as.data.frame(Embeddings(mayra_seurat_0, reduction = "umap")[, 1:2])
```


```{r }
ciara_genes=row.names(result)[result[,1]<1]
#geni_top=row.names(result)[result[,1]==0&result[,2]<0.0001]
geni_top_top=row.names(result)[order(as.numeric(result[,1]))]
```

```{r }
background=row.names(result)
```


```{r }
library(ggplot2)
p=list()
for(i in (geni_top_top)[1:10]){
  q=plot_gene(norm_counts_ane_0,cordinate_umap_0,i,i)
  p=list(p,q)
}
p
```


```{r }
localized_genes_mouse=detect_localized_genes(knn_matrix_ane_0,norm_counts_ane_0,geni_top_top,100) 
rank_intersect=localized_genes_mouse[[2]]
list_intersect=localized_genes_mouse[[1]]

ramp <- colorRamp(c("white", "blue4"))




ramp <- colorRamp(c("white", "blue4"))
 

rank_intersect=localized_genes_mouse[[2]]
list_intersect=localized_genes_mouse[[1]]


ramp.list <- rgb( ramp(seq(0, 1, length = length(unique(rank_intersect)))), max = 255)

index_color=round(length(ramp.list)/2,0)


breaks = seq(0,max(rank_intersect),length.out=1000)
gradient1 = colorpanel( sum( breaks[-1]<= as.numeric(quantile(breaks,0.25))), "#FFFFFF",ramp.list[index_color])

gradient2 = colorpanel( sum( breaks[-1] > as.numeric(quantile(breaks,0.25)) ), ramp.list[index_color], "#00008B" )


hm.colors = c(gradient1,gradient2)
 





genes_name_text=selection_localized_genes_here(list_intersect ,geni_top_top,max_number_genes=5)

plot_interactive_here(cordinate_umap_0,norm_counts_ane_0,rank_intersect,genes_name_text,hm.colors,min_x=NULL,max_x=NULL,min_y=NULL,max_y=NULL)

plot_genes_sum_here(cordinate_umap_0,norm_counts_ane_0,rank_intersect,"Top 100 genes CIARA")

```







```{r }
mayra_0_ciara=cluster_analysis_integrate_rare(mayra_dati_raw_0,"Mayra data",0.1,5,30,(ciara_genes))

#mayra_0_ciara=cluster_analysis_integrate_rare(mayra_dati_raw_0,"Mayra data",0.1,5,30,NULL)

```

```{r }

find_resolution(mayra_0_ciara,seq(0.01,1,0.1))

```


```{r }
pdf("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/github_january_2022/figures_paper/figures_conference/umap_cluster_ane_day_0.pdf",width = 8,height = 6)
plot_umap(cordinate_umap_0,as.vector(mayra_0_ciara$RNA_snn_res.0.1))
dev.off()

```

```{r}
table(as.vector(mayra_0_ciara$RNA_snn_res.0.1),cluster_number_0)
```

```{r}
pdf("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/github_january_2022/figures_paper/figures_conference/day_0_ane_gata6.pdf",width = 8,height = 6)
plot_gene(norm_counts_ane_0,cordinate_umap_0,"Gata6","Gata6")
dev.off()
cluster_2=as.vector(mayra_0_ciara$RNA_snn_res.0.1)
cluster_2[as.vector(mayra_0_ciara$RNA_snn_res.0.1)!=2]="Other cluster"

ind_1=which(cluster_2=="Other cluster")
ind_2=which(cluster_2==2)
ind_all=c(ind_1,ind_2)
cordinate_umap_0_new=cordinate_umap_0[ind_all,]
cluster_2_new=cluster_2[ind_all]
p=plot_umap(cordinate_umap_0_new,cluster_2_new)
p[[1]]+scale_color_manual(values=c("black", "white"))

```




# Original cluster annotation
```{r}
setwd("/Users/gabriele.lubatti/Desktop/Phd/Early_embryonic_like_cells/Mayra_project")
cluster_mayra=read.csv('LIFcells_leidenClustering.csv')
cluster_mayra_0=cluster_mayra[as.vector(cluster_mayra$batch)=='0h',]
row.names(cluster_mayra_0)=as.vector(cluster_mayra_0$cell)
cell_name=as.vector(cluster_mayra_0$cell)
cluster_number=(cluster_mayra_0$cluster)
cluster_number_0=as.character(cluster_number)

cell_name_new=substring(cell_name,1,nchar(cell_name)-4)

norm_counts_ane_0=norm_counts_ane_0[,cell_name_new]
```



```{r }
pdf("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/github_january_2022/figures_paper/figures_conference/umpa_mouse_esc_0_day.pdf",width = 8,height = 6)
cluster_number_0_letter=cluster_number_0
cluster_number_0_letter[cluster_number_0==0]="A"
cluster_number_0_letter[cluster_number_0==1]="B"
cluster_number_0_letter[cluster_number_0==2]="C"
cluster_number_0_letter[cluster_number_0==3]="D"
cluster_number_0_letter[cluster_number_0==4]="E"
cluster_number_0_letter[cluster_number_0==5]="F"
cc=rep(0,length(cluster_number_0))
cl=cluster_number_0_letter
cc[cl=="A"]="#E69F00"
cc[cl=="B"]="#56B4E9"
cc[cl=="C"]="#D62728"
cc[cl=="D"]="#009E73"
cc[cl=="E"]="#2ECC71"
cc[cl=="F"]="#9467BD"

p=plot_umap(cordinate_umap_0,cl)
p[[1]]+ scale_color_manual(values=c("#E69F00", "#56B4E9", "#D62728","#009E73","#2ECC71","#9467BD"))+theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"),axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),axis.text.y=element_blank(),
        axis.ticks.y=element_blank())+theme(text = element_text(size=18))
#plot_gene(norm_counts_ane_0,cordinate_umap_0,"Gata6","Gata6")
dev.off()
```




## Dendogram 24h + 48 h
```{r}
setwd("/Users/gabriele.lubatti/Desktop/Phd/Early_embryonic_like_cells/Mayra_project")
load(file='mayra_dati_raw_48.Rda')
load("knn_matrix_ane_48.Rda")
load("norm_counts_ane_48.Rda")
load("entropy_ane.Rda")

setwd("/Users/gabriele.lubatti/Desktop/Phd/Early_embryonic_like_cells/Mayra_project")
load(file='mayra_dati_raw_48.Rda')
mayra_seurat_48=cluster_analysis_integrate_rare(mayra_dati_raw_48,"Mayra_data_48",0.1,5,30)
cordinate_umap_48=as.data.frame(Embeddings(mayra_seurat_48, reduction = "umap")[, 1:2])
```



```{r}
setwd("/Users/gabriele.lubatti/Desktop/Phd/Early_embryonic_like_cells/Mayra_project")
cluster_mayra=read.csv('LIFcells_leidenClustering.csv')
cluster_mayra_48=cluster_mayra[as.vector(cluster_mayra$batch)=='48h',]
row.names(cluster_mayra_48)=as.vector(cluster_mayra_48$cell)
cell_name=as.vector(cluster_mayra_48$cell)
cluster_number_48=(cluster_mayra_48$cluster)
cluster_number_48=as.character(cluster_number_48)
cell_name_new=substring(cell_name,1,nchar(cell_name)-5)
```



```{r }
cluster_number_48_lett=cluster_number_48
cluster_number_48_lett[cluster_number_48==0]="A"
cluster_number_48_lett[cluster_number_48==1]="B"
cluster_number_48_lett[cluster_number_48==2]="C"
cluster_number_48_lett[cluster_number_48==3]="D"
cluster_number_48_lett[cluster_number_48==4]="E"
cluster_number_48_lett[cluster_number_48==5]="F"
mayra_seurat_48_small=cluster_analysis_integrate_rare(mayra_dati_raw_48[,cluster_number_48_lett=="C"|cluster_number_48_lett=="D"|cluster_number_48_lett=="F"],"Mayra_data_48",0.1,5,30)
```


```{r }
#setwd("/Users/gabriele.lubatti/Desktop/Phd/Early_embryonic_like_cells/Mouse/10x_data/Processed_data")
#load("ips_combined_old.Rda")
#load("es_combined_old.Rda")

#mayra_seurat_48
#mayra_seurat_24_update
livelli_24=levels(factor(final_cluster_version_sub))
livelli_48=levels(factor(cluster_number_48_lett[cluster_number_48_lett=="C"|cluster_number_48_lett=="D"|cluster_number_48_lett=="F"]))
livelli_24[livelli_24==0]="Pluri cells"
livelli_24[livelli_24=="1-step_2"]="2CLC"
livelli_24[livelli_24=="2-step_2"]="Precursor cells"

livelli_48[livelli_48=="C"]="Pluripotent cells"
livelli_48[livelli_48=="D"]="2CLC"
livelli_48[livelli_48=="F"]="Precursor cells"


norm_es=as.matrix(GetAssayData(mayra_seurat_48_small, slot = "data",assay="RNA"))
norm_ips=as.matrix(GetAssayData(mayra_seurat_24_update, slot = "data",assay="RNA"))
hvg_es=mayra_seurat_48_small@assays$RNA@var.features
hvg_ips=mayra_seurat_24_update@assays$RNA@var.features
hvg_all=c(hvg_es,hvg_ips)
hvg_all=unique(hvg_all)
find_median_cluster=function(cluster,norm,genes){
number_cluster=levels(factor(cluster))
median_es=rep(list(0),length(number_cluster))
for (i in 1:length(number_cluster)){
median_es[[i]]=apply(norm[genes,cluster==number_cluster[i]],1,'mean')

}
return(median_es)}

median_es=find_median_cluster(cluster_number_48_lett[cluster_number_48_lett=="C"|cluster_number_48_lett=="D"|cluster_number_48_lett=="F"],norm_es,hvg_all)
median_ips=find_median_cluster(final_cluster_version_sub,norm_ips,hvg_all)


median_es_matrix <- matrix(unlist(median_es), ncol = length(median_es), byrow = FALSE)
median_ips_matrix <- matrix(unlist(median_ips), ncol = length(median_ips), byrow = FALSE)

median_all_matrix=cbind(median_es_matrix,median_ips_matrix)
names_es=rep(0,length(median_es))
for ( i in 1:length(median_es)){
  j=i-1
names_es[i]=paste0("48h-",livelli_48[i])}

names_ips=rep(0,length(median_ips))
for ( i in 1:length(median_ips)){
  j=i-1
names_ips[i]=paste0("24h-",livelli_24[i])}

colnames(median_all_matrix)=c(names_es,names_ips)



median_all_matrix<-as.matrix(median_all_matrix)
dissimilarity <- sqrt((1-cor((median_all_matrix), method = "spearman"))/2)

my.dist <- as.dist(dissimilarity) 

#dist_mouse=dist(t(median_all_matrix),method="euclidean",upper=T)
#dist_mouse=as.matrix(dist_mouse)
dist_mouse=as.matrix(my.dist)
#dist_mouse=dist_mouse[grepl("48h-",row.names(dist_mouse)),grepl("24h-",row.names(dist_mouse))]


#heatmap(dist_mouse, scale = "none")




library("gplots")
heatmap.2(dist_mouse, scale = "none", col = bluered(100), 
          trace = "none", density.info = "none",cexRow = 0.5,cexCol = 0.5,notecex=5)


#https://r-graph-gallery.com/31-custom-colors-in-dendrogram.html see here for customization

pdf("/Users/gabriele.lubatti/Desktop/Phd/entropy_of_mixing_library/github_january_2022/figures_paper/figures_conference/dendogram_24_48.pdf",width=8,height = 6)
hc <- hclust(my.dist)


plot(hc)
dev.off()
```
```{r}

```


